### 2020-03-18 document.Rmd
first commit

---
title: "PIK3CA mutation prediction"
output: "html_document"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE)
```

```{r}
knitr::read_chunk('data.R')
knitr::read_chunk('prediction.R')
knitr::read_chunk('logisticRegression.R')
```

### Purpose

PIK3CA mutation prediction by gene expression data

### Introduction

Breast cancer with PIK3CA mutation has been approved to use PIK3CA inhibitor in hormone receptor positive HER2 negative subtype. Prediction of PIK3CA mutation was done by gene expression data of TCGA.

### Dataset

PIK3CA mutation data was get using cgdsr rpackage. Gene expression data was get from GDAC firehose using RTCGAToolbox R package. ER immunostain postive HER2 immunostain negative and/or SISH negative breast cancers are included. Data of invasive ductal carcinomas were used for training set and data of invasive lobular carcinoma were used for test set. Number of observations were 530 in training set and 188 in test set.

```{r load}
<<load>>
```

```{r}
<<modeling>>
```

```{r}
<<Finalize_model>>

```

Table 1. Frequency of PIK3CA mutation in training set and test set

```{r}
library(tidyverse)
train_target <- tibble::tibble(summary(trainSetDiff$PIK3CA_T)) 
test_target <- tibble::tibble(summary(testSetDiff$PIK3CA_T))
target <- bind_cols(train_target, test_target) %>%
  rename(TrainSet = "summary(trainSetDiff$PIK3CA_T)", TestSet = "summary(testSetDiff$PIK3CA_T)")

knitr::kable(target)

```

### Selecting variable for modeling

To narrow down potential predictors, Genes with a large variance (more than median) and differently expressed by PIK3CA mutation status (wilcox test, *P* $<10^{-8}$) were selected. 111 out of 20502 genes were included in the modeling process.

### Preprocessing

BoxCox transformation was done to correct skewness. Centering and scaling were done. All preprocessing was done using recipe r package.

### Modeling

Penalized logistic regression was applied to prediction modeling. 10-fold cross-validation with targe variable stratification was done over the hyperparameter grid: $\lambda$ {$10^{-5}$, $10^{-4}$,$10^{-3}$,$10^{-2}$,$10^{-1}$, $10^{0}$}, $\alpha$ {0.0, 0.25, 0.5, 0.75}. Model performance was evealuated with the receiver operating characteristic (AUROC).

The model showed best performance at lambda = 0.01 and alpha = 1.0 (Ridge regression).

```{r}
<<trainset_prediction>>

knitr::kable(confusion_matrix$table)
```

### 2020-03-18 all_histology.R

@@ -1,0 +1,166 @@
+## @knitr all_histology_diffset
+
+load('datasetHistology.RData')
+library(purrr)
+library(tidyverse)
+library(ggplot2)
+library(tidymodels)
+
+init_split <- initial_split(dataset, strata = HISTOLOGICAL_DIAGNOSIS)
+trainset_ah <- training(init_split)
+testset_ah <- testing(init_split)
+  
+variance <- trainset_ah %>%
+  select(-PIK3CA_T, -ID, -HISTOLOGICAL_DIAGNOSIS) %>%
+  map_dbl(var)
+
+sum(is.na(variance))
+summary(variance)
+summary(variance[variance > median(variance)])
+
+trainset_ahLargeVariance <- trainset_ah[,-1:-3][,variance > median(variance)]
+trainset_ahLargeVariance$PIK3CA_T <- trainset_ah$PIK3CA_T
+trainset_ahLargeVariance$HISTOLOGICAL_DIAGNOSIS <- trainset_ah$HISTOLOGICAL_DIAGNOSIS
+
+testset_ahLargeVariance <- testset_ah[,-1:-3][,variance > median(variance)]
+testset_ahLargeVariance$PIK3CA_T <- testset_ah$PIK3CA_T
+testset_ahLargeVariance$HISTOLOGICAL_DIAGNOSIS <- testset_ah$HISTOLOGICAL_DIAGNOSIS
+
+stasticsWilcox <- vector(mode = "list", length = dim(trainset_ahLargeVariance)[2])
+pvalueWilcox <- vector(mode = "list", length = dim(trainset_ahLargeVariance)[2])
+
+for (i in 1:(dim(trainset_ahLargeVariance)[2]-1)) {
+  a <- wilcox.test(trainset_ahLargeVariance[,i] ~ trainset_ahLargeVariance$PIK3CA_T)
+  stasticsWilcox[[i]] <- a$statistic
+  pvalueWilcox[[i]] <- a$p.value
+}
+
+colnames(trainset_ahLargeVariance)[log(unlist(pvalueWilcox), base = 10) < -8]
+
+trainset_ahDiff <- trainset_ahLargeVariance[,log(unlist(pvalueWilcox), base = 10) < -8]
+
+trainset_ahDiff$PIK3CA_T<- trainset_ah$PIK3CA_T
+
+trainset_ahDiff$HISTOLOGICAL_DIAGNOSIS <- trainset_ah$HISTOLOGICAL_DIAGNOSIS
+
[…]

### 2020-03-18 data.R

@@ -1,0 +1,117 @@
+## @knitr cbioportal
+
+library(cgdsr)
+library(tidyverse)
+library(RTCGAToolbox)
+library(data.table)
+library(readr)
+# Create CGDS object
+mycgds = CGDS("https://www.cbioportal.org/")
+
+test(mycgds)
+
+# Get list of cancer studies at server
+cancerStudy <- getCancerStudies(mycgds)
+
+breastPanCancer <- cancerStudy[48,1]
+# Get available case lists (collection of samples) for a given cancer study
+caselist = getCaseLists(mycgds,breastPanCancer)[1,1]
+
+# Get available genetic profiles
+geneticprofile = getGeneticProfiles(mycgds,breastPanCancer)
+
+mutation <- geneticprofile[5,1]
+
+# Get data slices for a specified list of genes, genetic profile and case list
+pik3ca <- getProfileData(mycgds,'PIK3CA', mutation, caselist)
+pik3caDataTable <- data.table(pik3ca)
+pik3caDataTable$ID <- gsub('\\.', '-',rownames(pik3ca))
+pik3caDataTable$PIK3CA_T <- is.na(pik3caDataTable$PIK3CA)
+pik3caDataTable$PIK3CA_T <- factor(pik3caDataTable$PIK3CA_T,
+                                   levels = c(TRUE, FALSE),
+                                   labels = c("Wild", "Mutant"))
+pik3caDataTable <- pik3caDataTable[,-1]
+
+# Get clinical data for the case list
+clinicaldata = getClinicalData(mycgds,caselist)
+
+clinicaldata <- clinicaldata %>%
+  mutate(ID = rownames(clinicaldata))
+
+
+
+breastFirehoseCancer <- cancerStudy[46,1]
+# Get available case lists (collection of samples) for a given cancer study
+caselistFirehose = getCaseLists(mycgds,breastFirehoseCancer)[1,1]
+
+# Get clinical data for the case list
+
+clinicaldataFirehose = getClinicalData(mycgds,caselistFirehose)
+clinicaldataFirehose <-  clinicaldataFirehose %>%
+  mutate(ID = rownames(clinicaldataFirehose))
+
[…]

### 2020-03-18 decisionTree.R

@@ -1,0 +1,65 @@
+load("trainset.RData")
+load("cart_tune.RData")
+
+library(tidymodels)
+library(rpart.plot)
+library(partykit)
+
+set.seed(930093)
+
+cv_splits <- rsample::vfold_cv(trainSetDiff, strata = PIK3CA_T)
+
+mod <- decision_tree(mode = "classification",
+                     cost_complexity = tune(),
+                     tree_depth = tune(),
+                     min_n = tune()) %>%
+    set_engine("rpart")
+
+rec <- recipe(PIK3CA_T ~ ., data = trainSetDiff)
+
+wfl <- 
+  workflow() %>%
+  add_recipe(rec) %>%
+  add_model(mod)
+
+cart_set <- parameters(cost_complexity(range = c(-10, -1)),
+                       tree_depth(range = c(1L, 5L)),
+                       min_n(range = c(5L, 20L)))
+
+cart_grid <- 
+  grid_regular(cart_set, levels = c(5, 5, 10))
+
+
+ctrl <- control_grid(save_pred = TRUE, verbose = TRUE)
+
+cart_tune <- 
+  tune_grid(wfl,
+            resamples = cv_splits,
+            grid = cart_grid,
+            metrics = metric_set(roc_auc),
+            control = ctrl)
+
+show_best(cart_tune)
+
+bestTune <- select_best(cart_tune)
+
+cart_rec_final <- prep(rec)
+
+cart_mod_final <- finalize_model(mod, bestTune)
+
+cart_fit <- 
+  cart_mod_final %>% 
+  fit(PIK3CA_T ~ ., data = juice(cart_rec_final))
+
+cart_fit$fit$call
+
+cart_fit$fit$call <- rlang::call2("rpart", PIK3CA_T ~ .,
+                                  data = expr(trainSetDiff),
+                                  cp = ~1e-10, maxdepth = ~5L, 
+                                  minsplit = ~15L)
+
+as.party()
+plot( as.party(cart_fit$fit) )
+rpart.plot(cart_fit$fit)
+
+save(cart_tune, file = "cart_tuen.RData")

### 2020-03-18 diffset.R

@@ -1,0 +1,45 @@
+load('dataset.RData')
+library(purrr)
+library(tidyverse)
+library(ggplot2)
+
+variance <- trainSet %>%
+  select(-PIK3CA_T) %>%
+  map_dbl(var)
+
+sum(is.na(variance))
+summary(variance)
+summary(variance[variance > median(variance)])
+
+trainsetLargeVariance <- trainSet[,-1][,variance > median(variance)]
+trainsetLargeVariance$PIK3CA_T <- trainSet$PIK3CA_T
+
+testsetLargeVariance <- testSet[,-1][,variance > median(variance)]
+testsetLargeVariance$PIK3CA_T <- testSet$PIK3CA_T
+
+stasticsWilcox <- vector(mode = "list", length = dim(trainsetLargeVariance)[2])
+pvalueWilcox <- vector(mode = "list", length = dim(trainsetLargeVariance)[2])
+
+for (i in 1:(dim(trainsetLargeVariance)[2]-1)) {
+  a <- wilcox.test(trainsetLargeVariance[,i] ~ trainsetLargeVariance$PIK3CA_T)
+  stasticsWilcox[[i]] <- a$statistic
+  pvalueWilcox[[i]] <- a$p.value
+}
+
+colnames(trainsetLargeVariance)[log(unlist(pvalueWilcox), base = 10) < -8]
+
+trainSetDiff <- trainsetLargeVariance[,log(unlist(pvalueWilcox), base = 10) < -8]
+
+trainSetDiff$PIK3CA_T<- trainSet$PIK3CA_T
+
+testSetDiff <- testsetLargeVariance[,log(unlist(pvalueWilcox), base = 10) < -8]
+
+testSetDiff$PIK3CA_T<- testSet$PIK3CA_T
+
+heatmap(cor(trainSetDiff))
+
+
+summary(trainSet$PIK3CA_T)
+summary(testSet$PIK3CA_T)
+
+save(trainSetDiff, testSetDiff,  file = "datasetDiff.RData")

### 2020-03-18 ductalPrediction.R

@@ -1,0 +1,157 @@
+## @knitr ductal_diffset
+
+load('ductalset.RData')
+library(purrr)
+library(tidyverse)
+library(ggplot2)
+library(tidymodels)
+
+init_split <- initial_split(ductalSet, strata = PIK3CA_T)
+trainset_ah <- training(init_split)
+testset_ah <- testing(init_split)
+
+variance <- trainset_ah %>%
+  select(-PIK3CA_T) %>%
+  map_dbl(var)
+
+sum(is.na(variance))
+summary(variance)
+summary(variance[variance > median(variance)])
+
+trainset_ahLargeVariance <- trainset_ah[,-1][,variance > median(variance)]
+trainset_ahLargeVariance$PIK3CA_T <- trainset_ah$PIK3CA_T
+
+testset_ahLargeVariance <- testset_ah[,-1][,variance > median(variance)]
+testset_ahLargeVariance$PIK3CA_T <- testset_ah$PIK3CA_T
+
+lobularsetLargeVariance <- lobularSet[,-1][,variance > median(variance)]
+lobularsetLargeVariance$PIK3CA_T <- lobularSet$PIK3CA_T
+
+stasticsWilcox <- vector(mode = "list", length = dim(trainset_ahLargeVariance)[2])
+pvalueWilcox <- vector(mode = "list", length = dim(trainset_ahLargeVariance)[2])
+
+for (i in 1:(dim(trainset_ahLargeVariance)[2]-1)) {
+  a <- wilcox.test(trainset_ahLargeVariance[,i] ~ trainset_ahLargeVariance$PIK3CA_T)
+  stasticsWilcox[[i]] <- a$statistic
+  pvalueWilcox[[i]] <- a$p.value
+}
+
+colnames(trainset_ahLargeVariance)[log(unlist(pvalueWilcox), base = 10) < -8]
+
+trainset_ahDiff <- trainset_ahLargeVariance[,log(unlist(pvalueWilcox), base = 10) < -8]
+
+trainset_ahDiff$PIK3CA_T<- trainset_ah$PIK3CA_T
+
+testset_ahDiff <- testset_ahLargeVariance[,log(unlist(pvalueWilcox), base = 10) < -8]
+
+testset_ahDiff$PIK3CA_T<- testset_ah$PIK3CA_T
+
+lobularset_Diff <- lobularsetLargeVariance[,log(unlist(pvalueWilcox), base = 10) < -8]
+
[…]

### 2020-03-18 logisticRegression.R

@@ -1,0 +1,48 @@
+## @knitr modeling
+
+load("datasetDiff.RData")
+load("glmn_tune.RData")
+
+library(tidymodels)
+library(glmnet)
+library(skimr)
+
+# skim(trainSetDiff)
+
+set.seed(930093)
+
+cv_splits <- rsample::vfold_cv(trainSetDiff, strata = PIK3CA_T, repeats = 5)
+
+mod <- logistic_reg(penalty = tune(),
+                    mixture = tune()) %>%
+  set_engine("glmnet")
+
+rec <- recipe(PIK3CA_T ~ ., data = trainSetDiff) %>%
+  step_BoxCox(all_predictors()) %>%
+  step_center(all_predictors()) %>%
+  step_scale(all_predictors()) %>%
+  step_downsample(PIK3CA_T)
+
+wfl <- 
+  workflow() %>%
+  add_recipe(rec) %>%
+  add_model(mod)
+
+glmn_set <- parameters(penalty(range = c(-5,3), trans = log10_trans()),
+                       mixture())
+
+glmn_grid <- 
+  grid_regular(glmn_set, levels = c(9, 5))
+
+ctrl <- control_grid(save_pred = TRUE, verbose = TRUE)
+
+## @knitr fitting
+
+glmn_tune <- 
+  tune_grid(wfl,
+            resamples = cv_splits,
+            grid = glmn_grid,
+            metrics = metric_set(roc_auc),
+            control = ctrl)
+
+

### 2020-03-18 prediction.R

@@ -1,0 +1,45 @@
+## @knitr load
+
+library(tidymodels)
+library(glmnet)
+
+load("datasetDiff.RData")
+load("glmn_tune.RData")
+
+show_best(glmn_tune)
+
+lasso_glmn <- show_best(glmn_tune)[4,1:2]
+best_glmn <- select_best(glmn_tune)
+mix <- show_best(glmn_tune)[3,1:2]
+
+## @knitr Finalize_model
+
+wfl_final <- 
+  wfl %>%
+  finalize_workflow(best_glmn) %>%
+  fit(data = trainSetDiff)
+
+## @knitr trainset_prediction
+
+train_probs <- 
+  predict(wfl_final, type = "prob", new_data = trainSetDiff) %>%
+  bind_cols(obs = trainSetDiff$PIK3CA_T) %>%
+  bind_cols(predict(wfl_final, new_data = trainSetDiff))
+
+confusion_matrix <- conf_mat(train_probs, obs, .pred_class)
+
+roc_curve_train <- autoplot(roc_curve(train_probs, obs, .pred_Mutant))
+
+roc_auc_train <- roc_auc(train_probs, obs, .pred_Mutant)
+
+## @knitr testset_prediction
+
+test_probs <- 
+  predict(wfl_final, type = "prob", new_data = testSetDiff) %>%
+  bind_cols(obs = testSetDiff$PIK3CA_T) %>%
+  bind_cols(predict(wfl_final, new_data = testSetDiff))
+
+conf_mat(test_probs, obs, .pred_class)
+autoplot(roc_curve(test_probs, obs, .pred_Mutant))
+
+roc_auc(test_probs, obs, .pred_Mutant)

### 2020-03-18 protein.R

@@ -1,0 +1,7 @@
+library(readr)
+library(tidyverse)
+
+protein <- read_delim("D:/Studies/TheHumanProteinAtlas/pathology.tsv/pathology.tsv", delim = "\t")
+
+protein %>% 
+  filter(Cancer == "breast cancer" & `Gene name` == " WBP1L")

### 2020-03-18 all_histology.R
coefficients

@@ -6,7 +6,7 @@ library(tidyverse)
 library(ggplot2)
 library(tidymodels)
 
-init_split <- initial_split(dataset, strata = HISTOLOGICAL_DIAGNOSIS)
+init_split <- initial_split(dataset, strata = PIK3CA_T)
 trainset_ah <- training(init_split)
 testset_ah <- testing(init_split)
   
@@ -20,14 +20,14 @@ summary(variance[variance > median(variance)])
 
 trainset_ahLargeVariance <- trainset_ah[,-1:-3][,variance > median(variance)]
 trainset_ahLargeVariance$PIK3CA_T <- trainset_ah$PIK3CA_T
-trainset_ahLargeVariance$HISTOLOGICAL_DIAGNOSIS <- trainset_ah$HISTOLOGICAL_DIAGNOSIS
+#trainset_ahLargeVariance$HISTOLOGICAL_DIAGNOSIS <- trainset_ah$HISTOLOGICAL_DIAGNOSIS
 
 testset_ahLargeVariance <- testset_ah[,-1:-3][,variance > median(variance)]
 testset_ahLargeVariance$PIK3CA_T <- testset_ah$PIK3CA_T
 testset_ahLargeVariance$HISTOLOGICAL_DIAGNOSIS <- testset_ah$HISTOLOGICAL_DIAGNOSIS
 
-stasticsWilcox <- vector(mode = "list", length = dim(trainset_ahLargeVariance)[2])
-pvalueWilcox <- vector(mode = "list", length = dim(trainset_ahLargeVariance)[2])
+stasticsWilcox <- vector(mode = "list", length = dim(trainset_ahLargeVariance)[2]-1)
+pvalueWilcox <- vector(mode = "list", length = dim(trainset_ahLargeVariance)[2]-1)
 
 for (i in 1:(dim(trainset_ahLargeVariance)[2]-1)) {
   a <- wilcox.test(trainset_ahLargeVariance[,i] ~ trainset_ahLargeVariance$PIK3CA_T)
@@ -39,7 +39,7 @@ colnames(trainset_ahLargeVariance)[log(unlist(pvalueWilcox), base = 10) < -8]
 
 trainset_ahDiff <- trainset_ahLargeVariance[,log(unlist(pvalueWilcox), base = 10) < -8]
 
[…]

### 2020-03-18 data.R

@@ -104,7 +104,9 @@ histology$ID <- gsub('\\.','-', histology$ID)
 
 dataset <- inner_join(histology, dataset, by = "ID")
 
-dataset[1:5,1:5]
+dataset <- dataset %>%
+  filter(HISTOLOGICAL_DIAGNOSIS == "Infiltrating Ductal Carcinoma" |
+           HISTOLOGICAL_DIAGNOSIS == "Infiltrating Lobular Carcinoma")
 
 ductalSet <- dataset %>%
   filter(HISTOLOGICAL_DIAGNOSIS == "Infiltrating Ductal Carcinoma") %>%

### 2020-03-20 pancancer_data.R
rna

@@ -1,0 +1,20 @@
+library(tidyverse)
+library(stringr)
+
+rna <- read_delim("D:/rproject/XCIpancancer/EBPlusPlusAdjustPANCAN_IlluminaHiSeq_RNASeqV2-v2.geneExp.tsv",
+                  delim = '\t')
+
+gene <- stringr::str_split_fixed(rna$gene_id, "\\|", n =2)[,1]
+sample_ID <- colnames(rna)[-1]
+
+data_rna <- data.table::data.table(t(rna[,-1]))
+colnames(data_rna) <- gene
+data_rna$ID <- sample_ID
+data_rna[1:5,1:5]
+data_rna$ID
+
+ID <- gsub("[A-Z]{1}-[0-9]{2}[A-Z]{1}-[0-9A-Z]{4}-[0-9]{2}", '',
+           data_rna$ID)
+data_rna_unique <- data_rna[duplicated(ID) == FALSE,]
+
+duplicated(ID)

### 2020-03-20 pancancer_data.R
f

@@ -18,3 +18,5 @@ ID <- gsub("[A-Z]{1}-[0-9]{2}[A-Z]{1}-[0-9A-Z]{4}-[0-9]{2}", '',
 data_rna_unique <- data_rna[duplicated(ID) == FALSE,]
 
 duplicated(ID)
+
+## Mutation

### 2020-03-24 pancancer_data.R
glmn

@@ -12,11 +12,142 @@ colnames(data_rna) <- gene
 data_rna$ID <- sample_ID
 data_rna[1:5,1:5]
 data_rna$ID
+data_rna$ID <- gsub("[A-Z]{1}-[0-9]{2}[A-Z]{1}-[0-9A-Z]{4}-[0-9]{2}", '', data_rna$ID)
 
-ID <- gsub("[A-Z]{1}-[0-9]{2}[A-Z]{1}-[0-9A-Z]{4}-[0-9]{2}", '',
-           data_rna$ID)
-data_rna_unique <- data_rna[duplicated(ID) == FALSE,]
+data_rna_unique <- data_rna[!duplicated(data_rna$ID), c(-1:-29)]
+data_rna_unique <- data_rna_unique[, c(-16273, -16274)]
+data_rna_unique$ID  <- gsub("-[0-9]{2}$", "", data_rna_unique$ID)
+sum(colnames(data_rna_unique) == `SLC35E2`)
 
-duplicated(ID)
+data_rna_unique <- data_rna_unique %>%
+  select_(as.name("SLC35E2"))
 
+grep(as.name("SLC35E2"), colnames(data_rna_unique))
+
+save(data_rna_unique, file = "panrna.RData")
 ## Mutation
+
+mut <- read_delim("mutations.txt", delim = '\t')
+summary(factor(mut$PIK3CA))
+
+## https://www.accessdata.fda.gov/scripts/cdrh/cfdocs/cfpma/pma.cfm?id=P190001
+
+variants <- c('C420R', 'E542K', 'E545A', 'E545D', 'E545G', 'E545K', 'Q546E', 'Q546R', 'H1047L', 'H1047R', 'H1047Y')
+
+mut %>%
+  filter(PIK3CA == variants)
+mut$PIK3CA %in% variants
+
+mut$PIK3CA
+
+variant_dual <- stringr::str_split_fixed(mut$PIK3CA, " ", n =2)
+variant_first <- mut$PIK3CA %in% variant_dual[,1]
+variant_second <- mut$PIK3CA %in% variant_dual[,2]
+
+mut$variant <- variant_first | variant_second
+mut <- mut %>%
+  mutate(ID = gsub("-[0-9]{2}$", "", SAMPLE_ID)) %>%
+  select(ID, variant)
+
+data <- inner_join(mut, data_rna_unique, by = "ID")
+save(data, file = "panrna.RData")
+
+load('panrna.RData')
+
[…]

### 2020-03-24 pancancer.Rmd
bibliography

--- 2020-03-18 document.Rmd
@@
 output: "html_document"
+bibliography: bibliography.bib
 ---
@@
 
-Breast cancer with PIK3CA mutation has been approved to use PIK3CA inhibitor in hormone receptor positive HER2 negative subtype. Prediction of PIK3CA mutation was done by gene expression data of TCGA.
+- PIK3CA mutation in selecting drug
+
+Breast cancer with PIK3CA mutation has been approved to use PIK3CA inhibitor in hormone receptor positive HER2 negative subtype.
+Prediction of PIK3CA mutation was done by gene expression data of TCGA.[]
+
+- Predicting mutation by gene expression data
+  * What for?  
+  * Feasible?  
+  * Cost?
+
+- Varing frquency of PIK3CA mutation across cancer types  
+  * High in endometrial breast  
+  * not common other cancer type
+
+Universial prediction using gene expression data across cancer type for certain kind of mutation.
+It's clinically useless now, but we want to explore the possibilities of the PIK3CA mutation prediction.
 

### 2020-03-24 pancancer_data.R

@@ -64,34 +64,15 @@ valueMad <- data %>%
 
 summary(valueMad)
 
-variance <- data %>%
-  select(-variant, -ID) %>%
-  map_dbl(var, na.rm = TRUE)
-
-sum(is.na(variance))
-summary(variance)
-summary(variance[variance > median(variance)])
-
 dataLV <- data[,-1:-2][,valueMad > median(valueMad)]
 
 dataLV$variant <- data$variant
 dataLV$ID <- data$ID
 
-stasticsWilcox <- vector(mode = "list", length = dim(dataLV)[2])
-pvalueWilcox <- vector(mode = "list", length = dim(dataLV)[2])
-
-for (i in 1:(dim(dataLV)[2]-1)) {
-  a <- wilcox.test(pull(dataLV[,i]) ~ dataLV$variant)
-  stasticsWilcox[[i]] <- a$statistic
-  pvalueWilcox[[i]] <- a$p.value
-}
 
 clinical <- read_delim("clinical.txt", delim = '\t')
 
 ## @knitr modeling
-
-load("datasetDiff.RData")
-load("glmn_tune.RData")
 
 library(tidymodels)
 library(glmnet)

### 2020-03-26 pancancer_data.R
success run without parallel processing

@@ -62,9 +62,9 @@ valueMad <- data %>%
   select(-variant, -ID) %>%
   map_dbl(mad, na.rm = TRUE)
 
-summary(valueMad)
+quantile(valueMad)[4]
 
-dataLV <- data[,-1:-2][,valueMad > median(valueMad)]
+dataLV <- data[,-1:-2][,valueMad > quantile(valueMad)[4]]
 
 dataLV$variant <- data$variant
 dataLV$ID <- data$ID
@@ -80,23 +80,26 @@ library(skimr)
 
 
 
-skim(clinical)
-summary(factor(clinical$type))
-head(clinical$bcr_patient_barcode)
+#skim(clinical)
+#summary(factor(clinical$type))
+#head(clinical$bcr_patient_barcode)
 
 hist <- clinical %>%
   select(bcr_patient_barcode, type) %>%
   rename(ID = "bcr_patient_barcode")
 
 dataset <- inner_join(hist, dataLV, by = "ID" )
+dataset$variant <- factor(dataset$variant)
+dataset$type <- factor(dataset$type)
 
+sum(is.na(dataset))
 set.seed(930093)
 
 initSplit <- initial_split(dataset, strata = variant)
 trainset <- training(initSplit)
 testset <- testing(initSplit)
 
-cv_splits <- rsample::vfold_cv(trainset, strata = variant , repeats = 5)
+cv_splits <- rsample::vfold_cv(trainset, strata = variant)
 
 mod <- logistic_reg(penalty = tune(),
                     mixture = tune()) %>%
@@ -104,6 +107,7 @@ mod <- logistic_reg(penalty = tune(),
 
 rec <- recipe(variant ~ ., data = trainset) %>%
   update_role(ID, new_role = "id variable") %>%
+  step_knnimpute(all_numeric()) %>%
   step_BoxCox(all_numeric()) %>%
   step_center(all_numeric()) %>%
   step_scale(all_numeric()) %>%
@@ -115,13 +119,13 @@ wfl <-
   add_recipe(rec) %>%
   add_model(mod)
 
-glmn_set <- parameters(penalty(range = c(-5,3), trans = log10_trans()),
+glmn_set <- parameters(penalty(range = c(-5,2), trans = log10_trans()),
                        mixture())
 
 glmn_grid <- 
-  grid_regular(glmn_set, levels = c(9, 5))
[…]

### 2020-03-26 recipe_reprex.R

@@ -1,0 +1,51 @@
+library(tidymodels)
+set.seed(930093)
+
+dataset <- iris %>% 
+  filter (Species != "virginica") %>%
+  mutate(Species = factor(Species))
+
+initSplit <- initial_split(dataset)
+trainset <- training(initSplit)
+testset <- testing(initSplit)
+
+cv_splits <- rsample::vfold_cv(trainset)
+
+mod <- logistic_reg(penalty = tune(),
+                    mixture = tune()) %>%
+  set_engine("glmnet")
+
+rec <- recipe(Species ~ ., data = trainset) %>%
+  step_center(all_numeric()) %>%
+  step_scale(all_numeric())
+
+
+wfl <- 
+  workflow() %>%
+  add_recipe(rec) %>%
+  add_model(mod)
+
+glmn_set <- parameters(penalty(range = c(-5,2), trans = log10_trans()),
+                       mixture())
+
+glmn_grid <- 
+  grid_regular(glmn_set, levels = c(8, 5))
+
+ctrl <- control_grid(save_pred = FALSE, verbose = TRUE)
+
+## @knitr fitting
+library(doParallel)
+cl <- makeCluster(2)
+registerDoParallel(cl)
+
+
+glmn_tune <- 
+  tune_grid(wfl,
+            resamples = cv_splits,
+            grid = glmn_grid,
+            metrics = metric_set(roc_auc),
+            control = ctrl)
+
+stopCluster(cl)
+
+glmn_tune$.notes[1]

### 2020-03-26 pancancer_data.R
subtype

@@ -173,9 +173,13 @@ roc_auc_train <- roc_auc(train_probs, obs, .pred_TRUE)
 test_probs <- 
   predict(wfl_final, type = "prob", new_data = testset) %>%
   bind_cols(obs = testset$variant) %>%
-  bind_cols(predict(wfl_final, new_data = testset))
+  bind_cols(predict(wfl_final, new_data = testset)) %>%
+  bind_cols(type = testset$type)
 
-conf_mat(test_probs, obs, .pred_class)
+test_probs %>%
+  filter(type == "BLCA") %>%
+  conf_mat(obs, .pred_class)
+
 autoplot(roc_curve(test_probs, obs, .pred_TRUE))
 
 roc_auc(test_probs, obs, .pred_TRUE)

### 2020-03-26 geo.R
first geo

@@ -1,0 +1,9 @@
+library(GEOquery)
+
+
+## https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE14860
+
+gse <- getGEO("GSE14860", GSEMatrix = TRUE)
+show(gse)
+
+save(gse, file = "gse.RData")

### 2020-03-26 pancancer_data.R

@@ -142,7 +142,7 @@ load("glmn_tune_pancancer.RData")
 glmn_tune_tidy <- glmn_tune %>%
   select(-splits)
 
-show_best(glmn_tune)
+show_best(glmn_tune_tidy)
 
 lasso_glmn <- show_best(glmn_tune)[4,1:2]
 best_glmn <- select_best(glmn_tune)

### 2020-03-31 geo.R
geo test

@@ -1,9 +1,65 @@
 library(GEOquery)
-
+library(tidyverse)
+library(tidymodels)
 
 ## https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE14860
 
-gse <- getGEO("GSE14860", GSEMatrix = TRUE)
+gse <- getGEO("GSE60788", GSEMatrix = TRUE)
 show(gse)
 
 save(gse, file = "gse.RData")
+load("gse.RData")
+
+summary(gse)
+
+dim(pData(gse[[1]]))
+pData(gse[[1]])
+
+# df1 <- getGSEDataTables("GSE60788")
+
+gunzip("GSE60788_UCSC_Human_hg19_knownGenes_20120910_addInfo.txt.gz")
+
+gds <- getGEO(filename=system.file("extdata/GSE22035.soft.gz",package="GEOquery"))
+
+gseRNA <- read_delim("GSE60788_rnaseq_gex_normalized.txt", delim = '\t')
+
+testsetGSE <- data.table::data.table(t(gseRNA[,-1]))
+colnames(testsetGSE) <- gseRNA$`Gene Symbol`
+
+## @knitr testsetGEO
+
+glmn_tune_tidy <- glmn_tune %>%
+  select(-splits)
+
+show_best(glmn_tune)
+
+lasso_glmn <- show_best(glmn_tune)[4,1:2]
+best_glmn <- select_best(glmn_tune)
+mix <- show_best(glmn_tune)[3,1:2]
+
+best_glmn$mixture <- 1
+best_glmn$penalty <- .1
+## @knitr Finalize_model
+
+wfl_final <- 
+  wfl %>%
+  finalize_workflow(best_glmn) %>%
+  fit(data = trainset)
+
+predict(wfl_final, type = "prob", new_data = testsetGSE)
+
+sum(colnames(trainset) %in% colnames(testsetGSE))
+
+test_probs <- 
+  predict(wfl_final, type = "prob", new_data = testsetGSE) %>%
+  bind_cols(obs = testset$variant) %>%
+  bind_cols(predict(wfl_final, new_data = testset)) %>%
+  bind_cols(type = testset$type)
+
+test_probs %>%
+  filter(type == "BLCA") %>%
+  conf_mat(obs, .pred_class)
+
+autoplot(roc_curve(test_probs, obs, .pred_TRUE))
+
+roc_auc(test_probs, obs, .pred_TRUE)

### 2020-03-31 pancancer_data.R

@@ -1,5 +1,9 @@
+## @knitr TCGA
+
 library(tidyverse)
+library(tidymodels)
 library(stringr)
+library(ggplot2)
 
 rna <- read_delim("D:/rproject/XCIpancancer/EBPlusPlusAdjustPANCAN_IlluminaHiSeq_RNASeqV2-v2.geneExp.tsv",
                   delim = '\t')
@@ -34,6 +38,8 @@ summary(factor(mut$PIK3CA))
 
 variants <- c('C420R', 'E542K', 'E545A', 'E545D', 'E545G', 'E545K', 'Q546E', 'Q546R', 'H1047L', 'H1047R', 'H1047Y')
 
+na_variant <- c("N345K", "E726K")
+
 mut %>%
   filter(PIK3CA == variants)
 mut$PIK3CA %in% variants
@@ -49,40 +55,101 @@ mut <- mut %>%
   mutate(ID = gsub("-[0-9]{2}$", "", SAMPLE_ID)) %>%
   select(ID, variant)
 
+## @knitr saveTCGA
+
 data <- inner_join(mut, data_rna_unique, by = "ID")
 save(data, file = "panrna.RData")
 
-load('panrna.RData')
+## @knitr GEO
 
-library(purrr)
-library(tidyverse)
-library(ggplot2)
+library(GEOquery)
 
-valueMad <- data %>%
-  select(-variant, -ID) %>%
+## https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE60788
+
+gse <- getGEO("GSE60788", GSEMatrix = TRUE)
+show(gse)
+
+save(gse, file = "gse.RData")
+load("gse.RData")
+
+summary(gse)
+
+dim(pData(gse[[1]]))
+phenotype <- pData(gse[[1]])
+skim(phenotype)
+
+summary(factor(phenotype$`pik3ca mutations:ch1`))
+
+phenotype$`pik3ca mutations:ch1`[grep("N345K", phenotype$`pik3ca mutations:ch1`)] <- NA
+
+variantGES <- phenotype %>%
+  mutate(variant = factor(is.na(`pik3ca mutations:ch1`) == FALSE)) %>%
+  select(title, variant) %>%
+  rename(ID = "title")
+
+tail(colnames(testsetGSE))
+ 
+# df1 <- getGSEDataTables("GSE60788")
+
+gunzip("GSE60788_UCSC_Human_hg19_knownGenes_20120910_addInfo.txt.gz")
+
[…]

### 2020-04-01 geo_HN.R
plot by type

@@ -1,0 +1,58 @@
+library(GEOquery)
+library(tidyverse)
+library(tidymodels)
+
+## https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE74927
+
+gse <- getGEO("GSE74927", GSEMatrix = TRUE)
+show(gse)
+
+summary(gse)
+
+dim(pData(gse[[1]]))
+pData(gse[[1]])
+
+gunzip("GSE74927_all_sample_raw_count.txt.gz")
+
+gseRNA <- read_delim("GSE74927_all_sample_raw_count.txt", delim = '\t')
+
+testsetGSE <- data.table::data.table(t(gseRNA[,-1]))
+colnames(testsetGSE) <- gseRNA$`Gene Symbol`
+
+## @knitr testsetGEO
+
+glmn_tune_tidy <- glmn_tune %>%
+  select(-splits)
+
+show_best(glmn_tune)
+
+lasso_glmn <- show_best(glmn_tune)[4,1:2]
+best_glmn <- select_best(glmn_tune)
+mix <- show_best(glmn_tune)[3,1:2]
+
+best_glmn$mixture <- 1
+best_glmn$penalty <- .1
+## @knitr Finalize_model
+
+wfl_final <- 
+  wfl %>%
+  finalize_workflow(best_glmn) %>%
+  fit(data = trainset)
+
+predict(wfl_final, type = "prob", new_data = testsetGSE)
+
+sum(colnames(trainset) %in% colnames(testsetGSE))
+
+test_probs <- 
+  predict(wfl_final, type = "prob", new_data = testsetGSE) %>%
+  bind_cols(obs = testset$variant) %>%
+  bind_cols(predict(wfl_final, new_data = testset)) %>%
+  bind_cols(type = testset$type)
+
+test_probs %>%
+  filter(type == "BLCA") %>%
+  conf_mat(obs, .pred_class)
+
+autoplot(roc_curve(test_probs, obs, .pred_TRUE))
+
+roc_auc(test_probs, obs, .pred_TRUE)

### 2020-04-01 pancancer_data.R

@@ -121,6 +121,8 @@ colnames(testsetGSE) <- gseRNA$`Gene Symbol`
 
 ## @knitr select matched genes
 load("panrna.RData")
+load("testsetGSE.RData")
+
 colnames(data)[1:5]
 data_temp <- data[, colnames(data) %in% colnames(testsetGSE)]
 #data_temp$ID <- data$ID
@@ -161,8 +163,11 @@ dataset$type <- factor(dataset$type)
 
 sum(is.na(dataset))
 
+save(dataset, testsetGSE, file = "dataset.RData")
+
 ## @knitr modeling
 
+load("dataset.RData")
 set.seed(930093)
 
 initSplit <- initial_split(dataset, strata = variant)
@@ -177,12 +182,12 @@ mod <- logistic_reg(penalty = tune(),
 
 rec <- recipe(variant ~ ., data = trainset) %>%
   update_role(ID, new_role = "id variable") %>%
-  step_knnimpute(all_numeric()) %>%
+  #step_knnimpute(all_numeric()) %>%
   step_YeoJohnson(all_numeric()) %>%
   step_center(all_numeric()) %>%
   step_scale(all_numeric()) %>%
-  step_dummy(type)%>%
-  step_downsample(variant)
+  step_dummy(type)#%>%
+  #step_downsample(variant)
 
 wfl <- 
   workflow() %>%
@@ -230,13 +235,60 @@ wfl_final <-
 train_probs <- 
   predict(wfl_final, type = "prob", new_data = trainset) %>%
   bind_cols(obs = trainset$variant) %>%
-  bind_cols(predict(wfl_final, new_data = trainset))
+  bind_cols(predict(wfl_final, new_data = trainset)) %>%
+  bind_cols(type = trainset$type)
 
-confusion_matrix <- conf_mat(train_probs, obs, .pred_class)
+confusion_matrix <- train_probs %>%
+  #group_by(type) %>%
+  conf_mat(obs, .pred_class)
+
+confusion_matrix
+
+confusion_matrix_type <- train_probs %>%
+  group_by(type) %>%
+  conf_mat(obs, .pred_class)
+
+confusion_matrix[1,2][[1]]
 
 roc_curve_train <- autoplot(roc_curve(train_probs, obs, .pred_TRUE))
 
-roc_auc_train <- roc_auc(train_probs, obs, .pred_TRUE)
[…]

### 2020-04-01 pancancer_data.R
remove geo

@@ -60,87 +60,19 @@ mut <- mut %>%
 data <- inner_join(mut, data_rna_unique, by = "ID")
 save(data, file = "panrna.RData")
 
-## @knitr GEO
-
-library(GEOquery)
-
-## https://www.ncbi.nlm.nih.gov/geo/query/acc.cgi?acc=GSE60788
-
-gse <- getGEO("GSE60788", GSEMatrix = TRUE)
-show(gse)
-
-save(gse, file = "gse.RData")
-load("gse.RData")
-
-summary(gse)
-
-dim(pData(gse[[1]]))
-phenotype <- pData(gse[[1]])
-skim(phenotype)
-
-summary(factor(phenotype$`pik3ca mutations:ch1`))
-
-phenotype$`pik3ca mutations:ch1`[grep("N345K", phenotype$`pik3ca mutations:ch1`)] <- NA
-
-variantGES <- phenotype %>%
-  mutate(variant = factor(is.na(`pik3ca mutations:ch1`) == FALSE)) %>%
-  select(title, variant) %>%
-  rename(ID = "title")
-
-tail(colnames(testsetGSE))
- 
-# df1 <- getGSEDataTables("GSE60788")
-
-gunzip("GSE60788_UCSC_Human_hg19_knownGenes_20120910_addInfo.txt.gz")
-
-gds <- getGEO(filename=system.file("extdata/GSE22035.soft.gz",package="GEOquery"))
-
-gseRNA <- read_delim("GSE60788_rnaseq_gex_normalized.txt", delim = '\t')
-
-gesClinical <- read_delim("GSE60788_UCSC_Human_hg19_knownGenes_20120910_addInfo.txt", delim = '\t')
-
-skim(gesClinical)
-
-colnames(gseRNA)[-1]
-
-testsetGSE <- data.table::data.table(t(gseRNA[,-1]))
-colnames(testsetGSE) <- gseRNA$`Gene Symbol`
-testsetGSE$type <- rep("BRCA", dim(testsetGSE)[1])
-
-testsetGSE$ID <- colnames(gseRNA)[-1]
-
-testsetGSE <- inner_join(variantGES, testsetGSE, by = "ID")
-
-testsetGSE$variant <- factor(testsetGSE$variant)
-
-## @knitr saveGSEtestset
-
-save(testsetGSE, file = "testsetGSE.RData")
-
-colnames(testsetGSE) <- gseRNA$`Gene Symbol`
 
[…]

### 2020-04-01 all_histology.R
delete files

@@ -1,179 +1,0 @@
-## @knitr all_histology_diffset
-
-load('datasetHistology.RData')
-library(purrr)
-library(tidyverse)
-library(ggplot2)
-library(tidymodels)
-
-init_split <- initial_split(dataset, strata = PIK3CA_T)
-trainset_ah <- training(init_split)
-testset_ah <- testing(init_split)
-  
-variance <- trainset_ah %>%
-  select(-PIK3CA_T, -ID, -HISTOLOGICAL_DIAGNOSIS) %>%
-  map_dbl(var)
-
-sum(is.na(variance))
-summary(variance)
-summary(variance[variance > median(variance)])
-
-trainset_ahLargeVariance <- trainset_ah[,-1:-3][,variance > median(variance)]
-trainset_ahLargeVariance$PIK3CA_T <- trainset_ah$PIK3CA_T
-#trainset_ahLargeVariance$HISTOLOGICAL_DIAGNOSIS <- trainset_ah$HISTOLOGICAL_DIAGNOSIS
-
-testset_ahLargeVariance <- testset_ah[,-1:-3][,variance > median(variance)]
-testset_ahLargeVariance$PIK3CA_T <- testset_ah$PIK3CA_T
-testset_ahLargeVariance$HISTOLOGICAL_DIAGNOSIS <- testset_ah$HISTOLOGICAL_DIAGNOSIS
-
-stasticsWilcox <- vector(mode = "list", length = dim(trainset_ahLargeVariance)[2]-1)
-pvalueWilcox <- vector(mode = "list", length = dim(trainset_ahLargeVariance)[2]-1)
-
-for (i in 1:(dim(trainset_ahLargeVariance)[2]-1)) {
-  a <- wilcox.test(trainset_ahLargeVariance[,i] ~ trainset_ahLargeVariance$PIK3CA_T)
-  stasticsWilcox[[i]] <- a$statistic
-  pvalueWilcox[[i]] <- a$p.value
-}
-
-colnames(trainset_ahLargeVariance)[log(unlist(pvalueWilcox), base = 10) < -8]
-
-trainset_ahDiff <- trainset_ahLargeVariance[,log(unlist(pvalueWilcox), base = 10) < -8]
-
-trainset_ahDiff$PIK3CA_T <- trainset_ah$PIK3CA_T
-
-trainset_ahDiff$HISTOLOGICAL_DIAGNOSIS <- trainset_ah$HISTOLOGICAL_DIAGNOSIS
-
[…]

### 2020-04-01 decisionTree.R

@@ -1,65 +1,0 @@
-load("trainset.RData")
-load("cart_tune.RData")
-
-library(tidymodels)
-library(rpart.plot)
-library(partykit)
-
-set.seed(930093)
-
-cv_splits <- rsample::vfold_cv(trainSetDiff, strata = PIK3CA_T)
-
-mod <- decision_tree(mode = "classification",
-                     cost_complexity = tune(),
-                     tree_depth = tune(),
-                     min_n = tune()) %>%
-    set_engine("rpart")
-
-rec <- recipe(PIK3CA_T ~ ., data = trainSetDiff)
-
-wfl <- 
-  workflow() %>%
-  add_recipe(rec) %>%
-  add_model(mod)
-
-cart_set <- parameters(cost_complexity(range = c(-10, -1)),
-                       tree_depth(range = c(1L, 5L)),
-                       min_n(range = c(5L, 20L)))
-
-cart_grid <- 
-  grid_regular(cart_set, levels = c(5, 5, 10))
-
-
-ctrl <- control_grid(save_pred = TRUE, verbose = TRUE)
-
-cart_tune <- 
-  tune_grid(wfl,
-            resamples = cv_splits,
-            grid = cart_grid,
-            metrics = metric_set(roc_auc),
-            control = ctrl)
-
-show_best(cart_tune)
-
-bestTune <- select_best(cart_tune)
-
-cart_rec_final <- prep(rec)
-
-cart_mod_final <- finalize_model(mod, bestTune)
-
-cart_fit <- 
-  cart_mod_final %>% 
-  fit(PIK3CA_T ~ ., data = juice(cart_rec_final))
-
-cart_fit$fit$call
-
-cart_fit$fit$call <- rlang::call2("rpart", PIK3CA_T ~ .,
-                                  data = expr(trainSetDiff),
-                                  cp = ~1e-10, maxdepth = ~5L, 
-                                  minsplit = ~15L)
-
-as.party()
-plot( as.party(cart_fit$fit) )
-rpart.plot(cart_fit$fit)
-
-save(cart_tune, file = "cart_tuen.RData")

### 2020-04-01 diffset.R

@@ -1,45 +1,0 @@
-load('dataset.RData')
-library(purrr)
-library(tidyverse)
-library(ggplot2)
-
-variance <- trainSet %>%
-  select(-PIK3CA_T) %>%
-  map_dbl(var)
-
-sum(is.na(variance))
-summary(variance)
-summary(variance[variance > median(variance)])
-
-trainsetLargeVariance <- trainSet[,-1][,variance > median(variance)]
-trainsetLargeVariance$PIK3CA_T <- trainSet$PIK3CA_T
-
-testsetLargeVariance <- testSet[,-1][,variance > median(variance)]
-testsetLargeVariance$PIK3CA_T <- testSet$PIK3CA_T
-
-stasticsWilcox <- vector(mode = "list", length = dim(trainsetLargeVariance)[2])
-pvalueWilcox <- vector(mode = "list", length = dim(trainsetLargeVariance)[2])
-
-for (i in 1:(dim(trainsetLargeVariance)[2]-1)) {
-  a <- wilcox.test(trainsetLargeVariance[,i] ~ trainsetLargeVariance$PIK3CA_T)
-  stasticsWilcox[[i]] <- a$statistic
-  pvalueWilcox[[i]] <- a$p.value
-}
-
-colnames(trainsetLargeVariance)[log(unlist(pvalueWilcox), base = 10) < -8]
-
-trainSetDiff <- trainsetLargeVariance[,log(unlist(pvalueWilcox), base = 10) < -8]
-
-trainSetDiff$PIK3CA_T<- trainSet$PIK3CA_T
-
-testSetDiff <- testsetLargeVariance[,log(unlist(pvalueWilcox), base = 10) < -8]
-
-testSetDiff$PIK3CA_T<- testSet$PIK3CA_T
-
-heatmap(cor(trainSetDiff))
-
-
-summary(trainSet$PIK3CA_T)
-summary(testSet$PIK3CA_T)
-
-save(trainSetDiff, testSetDiff,  file = "datasetDiff.RData")

### 2020-04-01 ductalPrediction.R

@@ -1,157 +1,0 @@
-## @knitr ductal_diffset
-
-load('ductalset.RData')
-library(purrr)
-library(tidyverse)
-library(ggplot2)
-library(tidymodels)
-
-init_split <- initial_split(ductalSet, strata = PIK3CA_T)
-trainset_ah <- training(init_split)
-testset_ah <- testing(init_split)
-
-variance <- trainset_ah %>%
-  select(-PIK3CA_T) %>%
-  map_dbl(var)
-
-sum(is.na(variance))
-summary(variance)
-summary(variance[variance > median(variance)])
-
-trainset_ahLargeVariance <- trainset_ah[,-1][,variance > median(variance)]
-trainset_ahLargeVariance$PIK3CA_T <- trainset_ah$PIK3CA_T
-
-testset_ahLargeVariance <- testset_ah[,-1][,variance > median(variance)]
-testset_ahLargeVariance$PIK3CA_T <- testset_ah$PIK3CA_T
-
-lobularsetLargeVariance <- lobularSet[,-1][,variance > median(variance)]
-lobularsetLargeVariance$PIK3CA_T <- lobularSet$PIK3CA_T
-
-stasticsWilcox <- vector(mode = "list", length = dim(trainset_ahLargeVariance)[2])
-pvalueWilcox <- vector(mode = "list", length = dim(trainset_ahLargeVariance)[2])
-
-for (i in 1:(dim(trainset_ahLargeVariance)[2]-1)) {
-  a <- wilcox.test(trainset_ahLargeVariance[,i] ~ trainset_ahLargeVariance$PIK3CA_T)
-  stasticsWilcox[[i]] <- a$statistic
-  pvalueWilcox[[i]] <- a$p.value
-}
-
-colnames(trainset_ahLargeVariance)[log(unlist(pvalueWilcox), base = 10) < -8]
-
-trainset_ahDiff <- trainset_ahLargeVariance[,log(unlist(pvalueWilcox), base = 10) < -8]
-
-trainset_ahDiff$PIK3CA_T<- trainset_ah$PIK3CA_T
-
-testset_ahDiff <- testset_ahLargeVariance[,log(unlist(pvalueWilcox), base = 10) < -8]
-
-testset_ahDiff$PIK3CA_T<- testset_ah$PIK3CA_T
-
-lobularset_Diff <- lobularsetLargeVariance[,log(unlist(pvalueWilcox), base = 10) < -8]
-
[…]

### 2020-04-02 pancancer.Rmd
success run without parallel processing
coeff

@@
 title: "PIK3CA mutation prediction"
-output: "html_document"
-bibliography: bibliography.bib
+output: "github_document"
+bibliography: references.bib
 ---
@@
 
-Breast cancer with PIK3CA mutation has been approved to use PIK3CA inhibitor in hormone receptor positive HER2 negative subtype.
-Prediction of PIK3CA mutation was done by gene expression data of TCGA.[]
+Breast cancer with PIK3CA mutation has been approved to use PIK3CA inhibitor in hormone receptor positive HER2 negative subtype. Prediction of PIK3CA mutation was done by gene expression data of TCGA.
 
@@
 
+RNAseq can be widely used. The mutation status directly predictio
 ### Dataset
@@
 10-fold cross-validation with targe variable stratification was done over the hyperparameter grid: $\lambda$ {$10^{-5}$, $10^{-4}$,$10^{-3}$,$10^{-2}$,$10^{-1}$, $10^{0}$}, $\alpha$ {0.0, 0.25, 0.5, 0.75}.
-Model performance was evealuated with the receiver operating characteristic (AUROC).
-
-The model showed best performance at lambda = 0.01 and alpha = 1.0 (Ridge regression).
+Model performance was evealuated with the area under receiver operating characteristic (ROC) and area under the ROC (AUROC) and precision recall (PR) and area under curve (AUPR).
 
@@
 ```
+
+## Results
+
+### selecting model
+Cross-validation 
+The model showed best performance at lambda = 0.01 and alpha = 1.0 (Ridge regression).
+
+### Prediction performance
+
+#### performance of each cancer type
+
+#### Important predictors

### 2020-04-02 pancancer_data.R

@@ -149,10 +149,11 @@ load("glmn_tune_pancancer.RData")
 glmn_tune_tidy <- glmn_tune %>%
   select(-splits)
 
-show_best(glmn_tune)
+show_best(glmn_tune, metric = "roc_auc")
+show_best(glmn_tune, metric = "pr_auc")
 
 lasso_glmn <- show_best(glmn_tune)[4,1:2]
-best_glmn <- select_best(glmn_tune)
+best_glmn <- select_best(glmn_tune, metric = "roc_auc")
 mix <- show_best(glmn_tune)[3,1:2]
 
 ## @knitr Finalize_model
@@ -162,128 +163,5 @@ wfl_final <-
   finalize_workflow(best_glmn) %>%
   fit(data = trainset)
 
-## @knitr trainset_prediction
+#save(wfl_final, file = "wflFinal.RData")
 
-train_probs <- 
-  predict(wfl_final, type = "prob", new_data = trainset) %>%
-  bind_cols(obs = trainset$variant) %>%
-  bind_cols(predict(wfl_final, new_data = trainset)) %>%
-  bind_cols(type = trainset$type)
-
-confusion_matrix <- train_probs %>%
-  #group_by(type) %>%
-  conf_mat(obs, .pred_class)
-
-confusion_matrix
-
-confusion_matrix_type <- train_probs %>%
-  group_by(type) %>%
-  conf_mat(obs, .pred_class)
-
-confusion_matrix[1,2][[1]]
-
-roc_curve_train <- autoplot(roc_curve(train_probs, obs, .pred_TRUE))
-
-roc_auc_train <- train_probs %>%
-  roc_auc(obs, .pred_TRUE)
-
-roc_auc_train$type <- "ALL"
-
-roc_auc_train_type <- train_probs %>%
-  group_by(type) %>%
-  roc_auc(obs, .pred_TRUE) %>%
-  mutate(type = as.character(type))
-
-roc_plot_type <- roc_auc_train_type %>%
-  bind_rows(roc_auc_train) %>%
-  ggplot(aes(x=reorder(type, -.estimate), y=.estimate)) +
-  geom_bar(stat="identity") + 
-  coord_flip()
-
-roc_auc_train <- train_probs %>%
-  roc_auc(obs, .pred_TRUE)
-
-roc_auc_train
-
-## @knitr correlation
-
-mutationRate <- trainset %>%
-  select(type, variant) %>%
-  group_by(type) %>%
[…]

### 2020-04-02 prediction.R

@@ -1,15 +1,18 @@
-## @knitr load
-
+## @knitr exploreCVResult
+library(tidyverse)
 library(tidymodels)
-library(glmnet)
+load(file = "wflFinal.RData")
+load(file = "glmn_tune_pancancer.RData")
 
-load("datasetDiff.RData")
-load("glmn_tune.RData")
 
-show_best(glmn_tune)
+glmn_tune_tidy <- glmn_tune %>%
+  select(-splits)
+
+show_best(glmn_tune, metric = "roc_auc")
+show_best(glmn_tune, metric = "pr_auc")
 
 lasso_glmn <- show_best(glmn_tune)[4,1:2]
-best_glmn <- select_best(glmn_tune)
+best_glmn <- select_best(glmn_tune, metric = "roc_auc")
 mix <- show_best(glmn_tune)[3,1:2]
 
 ## @knitr Finalize_model
@@ -17,29 +20,162 @@ mix <- show_best(glmn_tune)[3,1:2]
 wfl_final <- 
   wfl %>%
   finalize_workflow(best_glmn) %>%
-  fit(data = trainSetDiff)
+  fit(data = trainset)
+
+roc_vals <-
+  collect_metrics(glmn_tune) %>%
+  filter(.metric == "roc_auc")
+
+roc_vals %>%
+  mutate(mixture = format(mixture)) %>%
+  ggplot(aes(x = penalty, y = mean, col = mixture)) +
+  geom_line() +
+  geom_point() +
+  scale_x_log10()
+
+## @knitr coefficiency
+
+coeff <- tidy(extract_model(wfl_final)) %>%
+  filter(lambda > 0.09 & lambda < 0.104) %>%
+  arrange(-abs(estimate)) %>%
+  filter(!grepl("type_", term))
+
+head(coeff, n=30) %>%
+  filter(term != "(Intercept)") %>%
+  ggplot(aes(x=reorder(term, -estimate), y=estimate)) +
+    geom_bar(stat = "identity") +
+    coord_flip()
+
+tidy(extract_model(wfl_final)) %>%
+  filter(lambda > 0.09 & lambda < 0.104) %>%
+  arrange(-abs(estimate)) %>%
+  filter(grepl("type_", term)) %>%
+  ggplot(aes(x=reorder(term, -estimate), y=estimate)) +
+  geom_bar(stat = "identity") +
+  coord_flip()
 
 ## @knitr trainset_prediction
 
 train_probs <- 
[…]

### 2020-04-02 recipe_reprex.R

@@ -34,10 +34,6 @@ glmn_grid <-
 ctrl <- control_grid(save_pred = FALSE, verbose = TRUE)
 
 ## @knitr fitting
-library(doParallel)
-cl <- makeCluster(2)
-registerDoParallel(cl)
-
 
 glmn_tune <- 
   tune_grid(wfl,
@@ -46,6 +42,21 @@ glmn_tune <-
             metrics = metric_set(roc_auc),
             control = ctrl)
 
-stopCluster(cl)
+## finalize model
 
-glmn_tune$.notes[1]
+best_glmn <- select_best(glmn_tune, metric = "roc_auc")
+
+wfl_final <- 
+  wfl %>%
+  finalize_workflow(best_glmn) %>%
+  fit(data = trainset)
+
+## @knitr save
+
+save(wfl_final, file="wfl_final.RData")
+
+## @knitr load
+
+load("wfl_final.RData")
+
+tidy(extract_model(wfl_final))

### 2020-04-02 plosOne/plosOne.Rmd
plos template

--- 2020-04-02 pancancer.Rmd
@@
 ---
-title: "PIK3CA mutation prediction"
-output: "github_document"
-bibliography: references.bib
+title: Prediction of PIK3CA mutation with gene expression
+author:
+  - name: Jun Kang
+    email: jkang.alien@gmail.com
+    affiliation: Department of Hospital Pathology, Seoul St. Mary’s Hospital, College of Medicine, The Catholic University of Korea, Seoul, South Korea
+    corresponding: jkang.alien@gmail.com
+
+abstract: |
+  Lorem ipsum dolor sit amet, consectetur adipiscing elit.
+Curabitur eget porta erat.
+Morbi consectetur est vel gravida pretium.
+Suspendisse ut dui eu ante cursus gravida non sed sem.
+Nullam sapien tellus, commodo id velit id, eleifend volutpat quam.
+Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
+Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
+Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
+Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
+
+bibliography: mybibfile.bib
+output: rticles::plos_article
+csl: plos.csl
 ---
@@
 ```{r}
-knitr::read_chunk('data.R')
-knitr::read_chunk('prediction.R')
-knitr::read_chunk('logisticRegression.R')
+knitr::read_chunk('../data.R')
+knitr::read_chunk('../prediction.R')
+knitr::read_chunk('../logisticRegression.R')
 ```
 
-### Purpose
+# Introduction
 
-PIK3CA mutation prediction by gene expression data
+A list
 
-### Introduction
+- Brevity
+- Logic and clarity
+- Clean typing
+
+The problem
 
@@
 
-Breast cancer with PIK3CA mutation has been approved to use PIK3CA inhibitor in hormone receptor positive HER2 negative subtype. Prediction of PIK3CA mutation was done by gene expression data of TCGA.
+Breast cancer with PIK3CA mutation has been approved to use PIK3CA inhibitor in hormone receptor positive HER2 negative subtype.
+@andre_2019_alpelisib_nengljmeda Prediction of PIK3CA mutation was done by gene expression data of TCGA.
 
@@
 RNAseq can be widely used. The mutation status directly predictio
-### Dataset
+
+Previous study @way_2018_machine_cellreports
+
+The proposed solution
+
+# Materials and Methods
+
+## Dataset
 
@@
 Number of observations were 530 in training set and 188 in test set.
+
+## Selecting variable for modeling
+
+To narrow down potential predictors, Genes with a large variance (more than median) and differently expressed by PIK3CA mutation status (wilcox test, *P* $<10^{-8}$) were selected.
+111 out of 20502 genes were included in the modeling process.
+
+## Preprocessing
+
+BoxCox transformation was done to correct skewness. Centering and scaling were done. All preprocessing was done using recipe r package.
+
+## Modeling
+
+Penalized logistic regression was applied to prediction modeling.
+10-fold cross-validation with targe variable stratification was done over the hyperparameter grid: $\lambda$ {$10^{-5}$, $10^{-4}$,$10^{-3}$,$10^{-2}$,$10^{-1}$, $10^{0}$}, $\alpha$ {0.0, 0.25, 0.5, 0.75}.
+Model performance was evealuated with the area under receiver operating characteristic (ROC) and area under the ROC (AUROC) and precision recall (PR) and area under curve (AUPR).
+
+# Results
 
@@
   rename(TrainSet = "summary(trainSetDiff$PIK3CA_T)", TestSet = "summary(testSetDiff$PIK3CA_T)")
-
 knitr::kable(target)
@@
 ```
-
-### Selecting variable for modeling
-
-To narrow down potential predictors, Genes with a large variance (more than median) and differently expressed by PIK3CA mutation status (wilcox test, *P* $<10^{-8}$) were selected.
-111 out of 20502 genes were included in the modeling process.
-
-### Preprocessing
-
-BoxCox transformation was done to correct skewness. Centering and scaling were done. All preprocessing was done using recipe r package.
-
-### Modeling
-
-Penalized logistic regression was applied to prediction modeling.
-10-fold cross-validation with targe variable stratification was done over the hyperparameter grid: $\lambda$ {$10^{-5}$, $10^{-4}$,$10^{-3}$,$10^{-2}$,$10^{-1}$, $10^{0}$}, $\alpha$ {0.0, 0.25, 0.5, 0.75}.
-Model performance was evealuated with the area under receiver operating characteristic (ROC) and area under the ROC (AUROC) and precision recall (PR) and area under curve (AUPR).
 
@@
 
-## Results
-
-### selecting model
+## selecting model
 Cross-validation 
@@
 
-### Prediction performance
+## Prediction performance
 
-#### performance of each cancer type
+## performance of each cancer type
 
-#### Important predictors
+## Important predictors
+
+# Discussion
+
+Main message answers the question and main supporting evidence
+
+Critical assessment opinions on 
+- any shortcomings in study design  
+- limitations in methods
+- flaws in analysis
+- validity of assumption
+
+Comparison with other studies where inconsistencies are discussed
+
+Conclusions comments on possible biological or clinical implications and suggestions for further research.  
+Evaluate the results - not the authors
+
+# References {#references .unnumbered}

### 2020-04-02 plosOne/plosOne.tex
plos template

% Template for PLoS
% Version 3.5 March 2018
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended
% to minimize problems and delays during our production
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% Once your paper is accepted for publication,
% PLEASE REMOVE ALL TRACKED CHANGES in this file
% and leave only the final text of your manuscript.
% PLOS recommends the use of latexdiff to track changes during review, as this will help to maintain a clean tex file.
% Visit https://www.ctan.org/pkg/latexdiff?lang=en for info or contact us at latex@plos.org.
%
%
% There are no restrictions on package use within the LaTeX files except that
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% The manuscript LaTeX source should be contained within a single file (do not use \input, \externaldocument, or similar commands).
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file.
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission.
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig" instead of "Figure".
% See http://journals.plos.org/plosone/s/figures for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - spacing/line breaks within cells to alter layout or alignment
% - do not nest tabular environments (no tabular environments within tabular environments)
% - no graphics or colored text (cell background color/shading OK)
% See http://journals.plos.org/plosone/s/tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://journals.plos.org/plosone/s/latex
%
% For inline equations, please be sure to include all portions of an equation in the math environment.
%
% Do not include text that is not math in the math environment.
%
% Please add line breaks to long display equations when possible in order to fit size of the column.
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% When adding superscript or subscripts outside of brackets/braces, please group using {}.
%
% Do not use \cal for caligraphic font.  Instead, use \mathcal{}
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8x]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% cite package, to clean up citations in the main text. Do not remove.
% \usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref,hyperref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% color can be used to apply background shading to table cells only
\usepackage[table]{xcolor}

% array package and thick rules for tables
\usepackage{array}

% create "+" rule type for thick vertical lines
\newcolumntype{+}{!{\vrule width 2pt}}

% create \thickcline for thick horizontal lines of variable length
\newlength\savedwidth
\newcommand\thickcline[1]{%
  \noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
  \cline{#1}%
  \noalign{\vskip\arrayrulewidth}%
  \noalign{\global\arrayrulewidth\savedwidth}%
}

% \thickhline command for thick horizontal lines that span the table
\newcommand\thickhline{\noalign{\global\savedwidth\arrayrulewidth\global\arrayrulewidth 2pt}%
\hline
\noalign{\global\arrayrulewidth\savedwidth}}

% Remove comment for double spacing
%\usepackage{setspace}
%\doublespacing

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}
\renewcommand{\figurename}{Fig}

% Use the PLoS provided BiBTeX style
% \bibliographystyle{plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
%\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
%\setlength{\headheight}{27.023pt}
%\lhead{\includegraphics[width=2.0in]{PLOS-submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\today}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}

\usepackage{forarray}
\usepackage{xstring}
\newcommand{\getIndex}[2]{
  \ForEach{,}{\IfEq{#1}{\thislevelitem}{\number\thislevelcount\ExitForEach}{}}{#2}
}

\setcounter{secnumdepth}{0}

\newcommand{\getAff}[1]{
  \getIndex{#1}{}
}

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

\begin{document}
\vspace*{0.2in}

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Prediction of PIK3CA mutation with gene expression} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Jun Kang\textsuperscript{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital, College of
Medicine, The Catholic University of Korea, Seoul, South Korea}}\textsuperscript{*}\\
\bigskip
\bigskip
* Corresponding author: jkang.alien@gmail.com\\
\end{flushleft}
% Please keep the abstract below 300 words
\section*{Abstract}
Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget
porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut
dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id
velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus
finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium
diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut
ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit
pulvinar, volutpat risus.

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step.
% Author Summary not valid for PLOS ONE submissions.

\linenumbers

% Use "Eq" instead of "Equation" for equation citations.
\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

A list

\begin{itemize}
\tightlist
\item
  Brevity
\item
  Logic and clarity
\item
  Clean typing
\end{itemize}

The problem

\begin{itemize}
\tightlist
\item
  PIK3CA mutation in selecting drug
\end{itemize}

Breast cancer with PIK3CA mutation has been approved to use PIK3CA
inhibitor in hormone receptor positive HER2 negative subtype. {[}1{]}
Prediction of PIK3CA mutation was done by gene expression data of TCGA.

\begin{itemize}
\tightlist
\item
  Predicting mutation by gene expression data

  \begin{itemize}
  \tightlist
  \item
    What for?\\
  \item
    Feasible?\\
  \item
    Cost?
  \end{itemize}
\item
  Varing frquency of PIK3CA mutation across cancer types

  \begin{itemize}
  \tightlist
  \item
    High in endometrial breast\\
  \item
    not common other cancer type
  \end{itemize}
\end{itemize}

Universial prediction using gene expression data across cancer type for
certain kind of mutation. It's clinically useless now, but we want to
explore the possibilities of the PIK3CA mutation prediction.

RNAseq can be widely used. The mutation status directly predictio

Previous study {[}2{]}

The proposed solution

\hypertarget{materials-and-methods}{%
\section{Materials and Methods}\label{materials-and-methods}}

\hypertarget{dataset}{%
\subsection{Dataset}\label{dataset}}

PIK3CA mutation data was get using cgdsr rpackage. Gene expression data
was get from GDAC firehose using RTCGAToolbox R package. ER immunostain
postive HER2 immunostain negative and/or SISH negative breast cancers
are included. Data of invasive ductal carcinomas were used for training
set and data of invasive lobular carcinoma were used for test set.
Number of observations were 530 in training set and 188 in test set.

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Main message answers the question and main supporting evidence

Critical assessment opinions on - any shortcomings in study design\\
- limitations in methods - flaws in analysis - validity of assumption

Comparison with other studies where inconsistencies are discussed

Conclusions comments on possible biological or clinical implications and
suggestions for further research.\\
Evaluate the results - not the authors

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-andre_2019_alpelisib_nengljmeda}{}%
1. André F, Ciruelos E, Rubovszky G, Campone M, Loibl S, Rugo HS, et al.
Alpelisib for PIK3CA-Mutated, Hormone ReceptorPositive Advanced Breast
Cancer. New England Journal of Medicine. Massachusetts Medical Society;
2019;380: 1929--1940.
doi:\href{https://doi.org/10.1056/NEJMoa1813904}{10.1056/NEJMoa1813904}

\leavevmode\hypertarget{ref-way_2018_machine_cellreports}{}%
2. Way GP, Sanchez-Vega F, La K, Armenia J, Chatila WK, Luna A, et al.
Machine Learning Detects Pan-cancer Ras Pathway Activation in The Cancer
Genome Atlas. Cell Reports. Elsevier; 2018;23: 172--180.e3.
doi:\href{https://doi.org/10.1016/j.celrep.2018.03.046}{10.1016/j.celrep.2018.03.046}

\nolinenumbers

\end{document}

### 2020-04-03 plosOne/plosOne.Rmd
discussion

@@
 ```{r}
-knitr::read_chunk('../data.R')
 knitr::read_chunk('../prediction.R')
-knitr::read_chunk('../logisticRegression.R')
 ```
@@
 
-RNAseq can be widely used. The mutation status directly predictio
+RNAseq can be widely used. The mutation status directly prediction
 
@@
 
+prediction of aberant activation of a certain pathway vs a specific mutation
 The proposed solution
+
+Prediction RAS pathway activation with gene expression data was done in previous study.
+@way_2018_machine_cellreports They trained pancancer The Cancer Genome Atlas (TCGA) data with a supervised elastic net penalized logistic regression classifier with stochastic gradient descent.
+The performance of their model was 84% with an area under the receiver operating characteristic (AUROC) curve and 63% with an area under the precision recall (AUPR) curve.
+We applied their modeling methods in prediction PIK3CA mutation.
 
@@
 
+We used TCGA pancancer dataset.
 PIK3CA mutation data was get using cgdsr rpackage.
@@
 
-To narrow down potential predictors, Genes with a large variance (more than median) and differently expressed by PIK3CA mutation status (wilcox test, *P* $<10^{-8}$) were selected.
-111 out of 20502 genes were included in the modeling process.
+To narrow down potential predictors, Genes with a large the median absolute deviation (more than third-quartiles) were selected. 5000 out of 20502 genes were included in the modeling process.
 
@@
 
-BoxCox transformation was done to correct skewness. Centering and scaling were done. All preprocessing was done using recipe r package.
+Yeo-Johnson transformation was done to correct skewness. Centering and scaling were done. All preprocessing was done using recipe r package.
 
@@
 
-```{r load}
-<<load>>
-```
+## selecting model
 
-```{r}
-<<modeling>>
-```
-
-```{r}
-<<Finalize_model>>
-
-```
-
-Table 1. Frequency of PIK3CA mutation in training set and test set
-
-```{r}
-library(tidyverse)
-train_target <- tibble::tibble(summary(trainSetDiff$PIK3CA_T)) 
-test_target <- tibble::tibble(summary(testSetDiff$PIK3CA_T))
-target <- bind_cols(train_target, test_target) %>%
-  rename(TrainSet = "summary(trainSetDiff$PIK3CA_T)", TestSet = "summary(testSetDiff$PIK3CA_T)")
-knitr::kable(target)
-
-```
-
-```{r}
-<<trainset_prediction>>
-
-knitr::kable(confusion_matrix$table)
-```
-
-## selecting model
 Cross-validation  
@@
 
+Our model showed good performance to predict PIK3CA mutation of various cancer types.
+
 Main message answers the question and main supporting evidence
@@
 - any shortcomings in study design
+
+In this study, the split method was done for testset instead of indepedent testset.
+Because there is no standard analysis method for the RNA-seq gene expression quantification, it is difficult to find gene expression data same as TGCA data.
+Developing standard analysis method for gene expression data is neccessory to apply this prediction model.
+
 - limitations in methods
+
+Case imbalance
+
 - flaws in analysis
+
 - validity of assumption
@@
 # References {#references .unnumbered}
+
+# Figure legends

### 2020-04-03 plosOne/plosOne.tex
discussion

@@
 
-The proposed solution
+prediction of aberant activation of a certain pathway vs a specific
+mutation The proposed solution
+
+Prediction RAS pathway activation with gene expression data was done in
+previous study. {[}2{]} They trained pancancer The Cancer Genome Atlas
+(TCGA) data with a supervised elastic net penalized logistic regression
+classifier with stochastic gradient descent. The performance of their
+model was 84\% with an area under the receiver operating characteristic
+(AUROC) curve and 63\% with an area under the precision recall (AUPR)
+curve. We applied their modeling methods in prediction PIK3CA mutation.
 
@@
 
-PIK3CA mutation data was get using cgdsr rpackage. Gene expression data
-was get from GDAC firehose using RTCGAToolbox R package. ER immunostain
-postive HER2 immunostain negative and/or SISH negative breast cancers
-are included. Data of invasive ductal carcinomas were used for training
-set and data of invasive lobular carcinoma were used for test set.
-Number of observations were 530 in training set and 188 in test set.
+We used TCGA pancancer dataset. PIK3CA mutation data was get using cgdsr
+rpackage. Gene expression data was get from GDAC firehose using
+RTCGAToolbox R package. ER immunostain postive HER2 immunostain negative
+and/or SISH negative breast cancers are included. Data of invasive
+ductal carcinomas were used for training set and data of invasive
+lobular carcinoma were used for test set. Number of observations were
+530 in training set and 188 in test set.
+
+\hypertarget{selecting-variable-for-modeling}{%
+\subsection{Selecting variable for
+modeling}\label{selecting-variable-for-modeling}}
+
+To narrow down potential predictors, Genes with a large the median
+absolute deviation (more than third-quartiles) were selected. 5000 out
+of 20502 genes were included in the modeling process.
+
+\hypertarget{preprocessing}{%
+\subsection{Preprocessing}\label{preprocessing}}
+
+Yeo-Johnson transformation was done to correct skewness. Centering and
+scaling were done. All preprocessing was done using recipe r package.
+
+\hypertarget{modeling}{%
+\subsection{Modeling}\label{modeling}}
+
+Penalized logistic regression was applied to prediction modeling.
+10-fold cross-validation with targe variable stratification was done
+over the hyperparameter grid: \(\lambda\) \{\(10^{-5}\),
+\(10^{-4}\),\(10^{-3}\),\(10^{-2}\),\(10^{-1}\), \(10^{0}\)\},
+\(\alpha\) \{0.0, 0.25, 0.5, 0.75\}. Model performance was evealuated
+with the area under receiver operating characteristic (ROC) and area
+under the ROC (AUROC) and precision recall (PR) and area under curve
+(AUPR).
 
@@
 
+\hypertarget{selecting-model}{%
+\subsection{selecting model}\label{selecting-model}}
+
+Cross-validation\\
+The model showed best performance at lambda = 0.01 and alpha = 1.0
+(Ridge regression).
+
+\hypertarget{prediction-performance}{%
+\subsection{Prediction performance}\label{prediction-performance}}
+
+\hypertarget{performance-of-each-cancer-type}{%
+\subsection{performance of each cancer
+type}\label{performance-of-each-cancer-type}}
+
+\hypertarget{important-predictors}{%
+\subsection{Important predictors}\label{important-predictors}}
+
 \hypertarget{discussion}{%
@@
 
+Our model showed good performance to predict PIK3CA mutation of various
+cancer types.
+
 Main message answers the question and main supporting evidence
 
-Critical assessment opinions on - any shortcomings in study design\\
-- limitations in methods - flaws in analysis - validity of assumption
+Critical assessment opinions on - any shortcomings in study design
+
+In this study, the split method was done for testset instead of
+indepedent testset. Because there is no standard analysis method for the
+RNA-seq gene expression quantification, it is difficult to find gene
+expression data same as TGCA data. Developing standard analysis method
+for gene expression data is neccessory to apply this prediction model.
+
+\begin{itemize}
+\tightlist
+\item
+  limitations in methods
+\end{itemize}
+
+Case imbalance
+
+\begin{itemize}
+\item
+  flaws in analysis
+\item
+  validity of assumption
+\end{itemize}
 
@@
 \addcontentsline{toc}{section}{References}
+
+\hypertarget{figure-legends}{%
+\section*{Figure legends}\label{figure-legends}}
+\addcontentsline{toc}{section}{Figure legends}
 

### 2020-04-03 pancancer_data.R

@@ -58,7 +58,7 @@ mut <- mut %>%
 ## @knitr saveTCGA
 
 data <- inner_join(mut, data_rna_unique, by = "ID")
-save(data, file = "panrna.RData")
+#save(data, file = "panrna.RData")
 
 
 ## @knitr select matched genes
@@ -163,5 +163,5 @@ wfl_final <-
   finalize_workflow(best_glmn) %>%
   fit(data = trainset)
 
-#save(wfl_final, file = "wflFinal.RData")
+save(wfl_final, file = "wflFinal.RData")
 

### 2020-04-03 prediction.R

@@ -1,7 +1,7 @@
 ## @knitr exploreCVResult
 library(tidyverse)
 library(tidymodels)
-load(file = "wflFinal.RData")
+load(file = "D:/rproject/pik3ca/wflFinal.RData")
 load(file = "glmn_tune_pancancer.RData")
 
 
@@ -40,16 +40,18 @@ coeff <- tidy(extract_model(wfl_final)) %>%
   arrange(-abs(estimate)) %>%
   filter(!grepl("type_", term))
 
-head(coeff, n=30) %>%
+coeffPlot <- head(coeff, n=30) %>%
   filter(term != "(Intercept)") %>%
   ggplot(aes(x=reorder(term, -estimate), y=estimate)) +
     geom_bar(stat = "identity") +
     coord_flip()
 
-tidy(extract_model(wfl_final)) %>%
+typeCoeff <- tidy(extract_model(wfl_final)) %>%
   filter(lambda > 0.09 & lambda < 0.104) %>%
   arrange(-abs(estimate)) %>%
-  filter(grepl("type_", term)) %>%
+  filter(grepl("type_", term))
+
+typePlot <- typeCoeff %>%
   ggplot(aes(x=reorder(term, -estimate), y=estimate)) +
   geom_bar(stat = "identity") +
   coord_flip()
@@ -108,7 +110,7 @@ mutationRate <- trainset %>%
   mutate(rate = `TRUE`/(`TRUE`+`FALSE`)) %>%
   bind_cols(roc_auc_train_type)
 
-mutationRate %>%
+mutatioPlot <- mutationRate %>%
   ggplot(aes(x=rate, y=.estimate)) +
   geom_point() +
   geom_smooth(method='lm', formula= y~x) +
@@ -154,7 +156,7 @@ roc_plot_test_type <- roc_auc_test_type %>%
   geom_bar(stat="identity") + 
   coord_flip()
 
-plot(roc_auc_train_type$.estimate, roc_auc_test_type$.estimate)
+test_trainPlot <- plot(roc_auc_train_type$.estimate, roc_auc_test_type$.estimate)
 
 roc_plot_test_type
 
@@ -165,7 +167,7 @@ roc_auc_test
 
 ## @knitr correlation testset
 
-mutationRate <- testset %>%
+mutationRateTest <- testset %>%
   select(type, variant) %>%
   group_by(type) %>%
   count(variant == TRUE) %>%
[…]

### 2020-04-06 prediction.R
aupr

@@ -1,7 +1,7 @@
 ## @knitr exploreCVResult
 library(tidyverse)
 library(tidymodels)
-load(file = "D:/rproject/pik3ca/wflFinal.RData")
+
 load(file = "glmn_tune_pancancer.RData")
 
 
@@ -81,6 +81,13 @@ roc_curve_train <- autoplot(roc_curve(train_probs, obs, .pred_TRUE))
 roc_auc_train <- train_probs %>%
   roc_auc(obs, .pred_TRUE)
 
+pr_auc_train <- train_probs %>%
+  pr_auc(obs, .pred_TRUE)
+
+pr_auc_train
+
+pr_curve_train <- autoplot(pr_curve(train_probs, obs, .pred_TRUE))
+
 roc_auc_train$type <- "ALL"
 
 roc_auc_train_type <- train_probs %>%
@@ -94,10 +101,6 @@ roc_plot_type <- roc_auc_train_type %>%
   geom_bar(stat="identity") + 
   coord_flip()
 
-roc_auc_train <- train_probs %>%
-  roc_auc(obs, .pred_TRUE)
-
-roc_auc_train
 
 ## @knitr correlation
 
@@ -164,6 +167,16 @@ roc_auc_test <- test_probs %>%
   roc_auc(obs, .pred_TRUE)
 
 roc_auc_test
+
+
+pr_auc_test <- test_probs %>%
+  pr_auc(obs, .pred_TRUE)
+
+pr_auc_test
+
+pr_curve_test <- autoplot(pr_curve(test_probs, obs, .pred_TRUE))
+
+pr_curve_test
 
 ## @knitr correlation testset
 

### 2020-04-07 figure.R
down

@@ -1,0 +1,55 @@
+load('result.RData')
+
+library(Cairo)
+library(tidyverse)
+library(tidymodels)
+library(gridExtra)
+
+source("theme_publication.R")
+
+### ROC PR Plots
+ROCPlot <- bind_rows("Trainset" = train_probs,
+          "Testset" = test_probs,
+          .id = "dataset") %>%
+  group_by(dataset) %>%
+  roc_curve(obs, .pred_TRUE) %>%
+    ggplot(aes(x=1-specificity, y=sensitivity, color = dataset)) +
+  geom_line() +
+  labs(title = "ROC curve") +
+  scale_colour_Publication() +
+  theme_Publication()
+  
+
+
+prPlot <-  bind_rows("Trainset" = train_probs,
+                     "Testset" = test_probs,
+                     .id = "dataset") %>%
+  group_by(dataset) %>%
+  pr_curve(obs, .pred_TRUE) %>%
+  ggplot(aes(x=recall, y=precision, color = dataset)) +
+  geom_line() +
+  labs(title = "PR curve") +
+  scale_colour_Publication() +
+  theme_Publication()
+
+CairoPDF(file = 'Figure1',
+         width = 6, height = 3)
+grid.arrange(ROCPlot, prPlot, nrow = 1)
+dev.off()
+
+typePlot +
+  labs(title = "Coefficiency of cancer type",
+       x = "Coefficiency",
+       y = "Cancer type") +
+  scale_colour_Publication() +
+  theme_Publication()
+
+### Train_test ROC 
+
+Train_test_ROC <- test_trainPlot +
+  labs(title = "AUROC of trainset and testset",
+       x = "Trainset",
+       y = "Testset") +
+  scale_colour_Publication() +
+  theme_Publication()
+  

### 2020-04-07 pancancer_data.R

@@ -90,7 +90,9 @@ hist <- clinical %>%
   rename(ID = "bcr_patient_barcode")
 
 dataset <- inner_join(hist, dataLV, by = "ID" )
-dataset$variant <- factor(dataset$variant)
+dataset$variant <- factor(dataset$variant,
+                          levels = c(TRUE, FALSE),
+                          labels = c('Mutant', 'Wild'))
 dataset$type <- factor(dataset$type)
 
 sum(is.na(dataset))
@@ -118,12 +120,26 @@ rec <- recipe(variant ~ ., data = trainset) %>%
   step_YeoJohnson(all_numeric()) %>%
   step_center(all_numeric()) %>%
   step_scale(all_numeric()) %>%
-  step_dummy(type)#%>%
+  step_dummy(type) #%>%
   #step_downsample(variant)
 
 wfl <- 
   workflow() %>%
   add_recipe(rec) %>%
+  add_model(mod)
+
+rec_down <- recipe(variant ~ ., data = trainset) %>%
+  update_role(ID, new_role = "id variable") %>%
+  #step_knnimpute(all_numeric()) %>%
+  step_YeoJohnson(all_numeric()) %>%
+  step_center(all_numeric()) %>%
+  step_scale(all_numeric()) %>%
+  step_dummy(type) %>%
+  step_downsample(variant)
+
+wfl_down <- 
+  workflow() %>%
+  add_recipe(rec_down) %>%
   add_model(mod)
 
 glmn_set <- parameters(penalty(range = c(-5,2), trans = log10_trans()),
@@ -132,7 +148,7 @@ glmn_set <- parameters(penalty(range = c(-5,2), trans = log10_trans()),
 glmn_grid <- 
   grid_regular(glmn_set, levels = c(8, 5))
 
-ctrl <- control_grid(save_pred = FALSE, verbose = TRUE)
+ctrl <- control_grid(save_pred = TRUE, verbose = TRUE)
 
 ## @knitr fitting
 
@@ -142,18 +158,28 @@ glmn_tune <-
             grid = glmn_grid,
             metrics = metric_set(roc_auc, pr_auc),
             control = ctrl)
+glmn_tune_down <- 
+  tune_grid(wfl_down,
+            resamples = cv_splits,
+            grid = glmn_grid,
+            metrics = metric_set(roc_auc, pr_auc),
+            control = ctrl)
 
[…]

### 2020-04-07 prediction.R

@@ -8,18 +8,27 @@ load(file = "glmn_tune_pancancer.RData")
 glmn_tune_tidy <- glmn_tune %>%
   select(-splits)
 
-show_best(glmn_tune, metric = "roc_auc")
-show_best(glmn_tune, metric = "pr_auc")
+best_roc <- show_best(glmn_tune, metric = "roc_auc")
+best_pr <- show_best(glmn_tune, metric = "pr_auc")
+
+best_roc_down <- show_best(glmn_tune_down, metric = "roc_auc")
+best_pr_down <- show_best(glmn_tune_down, metric = "pr_auc")
 
 lasso_glmn <- show_best(glmn_tune)[4,1:2]
 best_glmn <- select_best(glmn_tune, metric = "roc_auc")
 mix <- show_best(glmn_tune)[3,1:2]
 
+best_glmn_down <- select_best(glmn_tune_down, metric = "roc_auc")
 ## @knitr Finalize_model
 
 wfl_final <- 
   wfl %>%
   finalize_workflow(best_glmn) %>%
+  fit(data = trainset)
+
+wfl_down_final <- 
+  wfl_down %>%
+  finalize_workflow(best_glmn_down) %>%
   fit(data = trainset)
 
 roc_vals <-
@@ -79,36 +88,53 @@ confusion_matrix[1,2][[1]]
 roc_curve_train <- autoplot(roc_curve(train_probs, obs, .pred_TRUE))
 
 roc_auc_train <- train_probs %>%
-  roc_auc(obs, .pred_TRUE)
-
-pr_auc_train <- train_probs %>%
-  pr_auc(obs, .pred_TRUE)
-
-pr_auc_train
-
-pr_curve_train <- autoplot(pr_curve(train_probs, obs, .pred_TRUE))
+  roc_auc(obs, .pred_FALSE)
 
 roc_auc_train$type <- "ALL"
 
 roc_auc_train_type <- train_probs %>%
   group_by(type) %>%
-  roc_auc(obs, .pred_TRUE) %>%
+  roc_auc(obs, .pred_Mutant) %>%
   mutate(type = as.character(type))
 
 roc_plot_type <- roc_auc_train_type %>%
+  filter(is.na(.estimate) == FALSE) %>%
   bind_rows(roc_auc_train) %>%
   ggplot(aes(x=reorder(type, -.estimate), y=.estimate)) +
   geom_bar(stat="identity") + 
   coord_flip()
 
[…]

### 2020-04-08 figure.R
aupr

@@ -12,9 +12,11 @@ ROCPlot <- bind_rows("Trainset" = train_probs,
           "Testset" = test_probs,
           .id = "dataset") %>%
   group_by(dataset) %>%
-  roc_curve(obs, .pred_TRUE) %>%
+  roc_curve(obs, .pred_Mutant) %>%
     ggplot(aes(x=1-specificity, y=sensitivity, color = dataset)) +
   geom_line() +
+  geom_text(aes(x=.6, y=.6, label = "AUROC = 0.93")) + 
+  geom_text(aes(x=.6, y=.4, label = "AUROC = 0.84"), color = "#386cb0") +
   labs(title = "ROC curve") +
   scale_colour_Publication() +
   theme_Publication()
@@ -25,9 +27,12 @@ prPlot <-  bind_rows("Trainset" = train_probs,
                      "Testset" = test_probs,
                      .id = "dataset") %>%
   group_by(dataset) %>%
-  pr_curve(obs, .pred_TRUE) %>%
+  pr_curve(obs, .pred_Mutant) %>%
   ggplot(aes(x=recall, y=precision, color = dataset)) +
   geom_line() +
+  geom_hline(yintercept = 0.11, color = 3)+
+  geom_text(aes(x=.7, y=1.0, label = "AUPR = 0.66")) + 
+  geom_text(aes(x=.7, y=.8, label = "AUPR = 0.39"), color = "#386cb0") +
   labs(title = "PR curve") +
   scale_colour_Publication() +
   theme_Publication()
@@ -53,3 +58,9 @@ Train_test_ROC <- test_trainPlot +
   scale_colour_Publication() +
   theme_Publication()
   
+bind_rows("Trainset" = train_probs,
+          "Testset" = test_probs,
+          .id = "dataset") %>%
+  skimr::skim()
+  
+

### 2020-04-08 prediction.R

@@ -196,76 +196,33 @@ pr_curve_test_down <- autoplot(pr_curve(test_probs_down, obs, .pred_Mutant))
 
 pr_curve_test
 
-roc_auc_test$type <- "ALL"
+pr_auc_test$type <- "ALL"
 
-roc_auc_test_type <- test_probs %>%
+pr_auc_test_type <- test_probs %>%
   group_by(type) %>%
-  roc_auc(obs, .pred_TRUE) %>%
+  pr_auc(obs, .pred_Mutant) %>%
   mutate(type = as.character(type))
 
-roc_plot_test_type <- roc_auc_test_type %>%
-  bind_rows(roc_auc_test) %>%
+pr_plot_test_type <- pr_auc_test_type %>%
+  bind_rows(pr_auc_test) %>%
   filter(is.na(.estimate) == FALSE) %>%
   ggplot(aes(x=reorder(type, -.estimate), y=.estimate)) +
   geom_bar(stat="identity") + 
   coord_flip()
 
-roc_train_test <- merge(roc_auc_train_type,
-                        roc_auc_test_type,
-                        by = "type")
+pr_train_test <- merge(pr_auc_train_type,
+                       pr_auc_test_type,
+                       by = "type")
 
-test_trainPlot <- roc_train_test %>%
+pr_test_trainPlot <- pr_train_test %>%
   ggplot(aes(x=.estimate.x, y=.estimate.y)) +
   geom_point() +
   geom_smooth(method='lm', formula= y~x) +
-  ggrepel::geom_text_repel(data = roc_train_test,
+  ggrepel::geom_text_repel(data = pr_train_test,
                            aes(label = type))
 
-test_trainPlot
+pr_test_trainPlot
 
-
-roc_auc_test_type <- test_probs %>%
-  group_by(type) %>%
-  roc_auc(obs, .pred_TRUE) %>%
-  mutate(type = as.character(type))
-
-roc_plot_test_type <- roc_auc_test_type %>%
-  bind_rows(roc_auc_test) %>%
-  filter(is.na(.estimate) == FALSE) %>%
-  ggplot(aes(x=reorder(type, -.estimate), y=.estimate)) +
-  geom_bar(stat="identity") + 
-  coord_flip()
-
-roc_train_test <- merge(roc_auc_train_type,
-                        roc_auc_test_type,
-                        by = "type")
-
-
[…]

### 2020-04-08 figure.R
figure1

@@ -23,6 +23,7 @@ ROCPlot <- bind_rows("Trainset" = train_probs,
   
 
 
+
 prPlot <-  bind_rows("Trainset" = train_probs,
                      "Testset" = test_probs,
                      .id = "dataset") %>%
@@ -37,10 +38,64 @@ prPlot <-  bind_rows("Trainset" = train_probs,
   scale_colour_Publication() +
   theme_Publication()
 
+cor_ROC <- roc_test_trainPlot +
+  labs(title = "AUROC",
+       x = "Trainset",
+       y = "Testset") +
+  scale_colour_Publication() +
+  theme_Publication()
+
+cor_PR <- pr_test_trainPlot +
+  labs(title = "AUPR",
+       x = "Trainset",
+       y = "Testset") +
+  scale_colour_Publication() +
+  theme_Publication()
+
+mutationROC <- mutatioPlot +
+  labs(title = "AUROC",
+       x = "Mutation rate",
+       y = "ROC") +
+  scale_colour_Publication() +
+  theme_Publication()
+
+mutationPR <- mutatioPlotPR +
+  labs(title = "AUPR",
+       x = "Mutation rate",
+       y = "PR") +
+  scale_colour_Publication() +
+  theme_Publication()
+
+geneCoeff <- coeffPlot +
+  labs(title = "Coefficiency of genes",
+       x = "Coefficiency",
+       y = "Gene") +
+  scale_colour_Publication() +
+  theme_Publication()
+
+typeCoeff <- typePlot +
+  labs(title = "Coefficiency of cancer type",
+       x = "Coefficiency",
+       y = "Cancer type") +
+  scale_colour_Publication() +
+  theme_Publication()
+
+
 CairoPDF(file = 'Figure1',
-         width = 6, height = 3)
-grid.arrange(ROCPlot, prPlot, nrow = 1)
+         width = 8, height = 12, pointsize=10)
+grid.arrange(ROCPlot, prPlot,
+             cor_ROC, cor_PR,
+             mutationROC, mutationPR,
+             nrow = 3)
 dev.off()
+
+CairoPDF(file = 'Figure2',
+         width = 8, height = 12)
+grid.arrange(geneCoeff, typeCoeff,
+             nrow = 1)
+dev.off()
+
+
 
 typePlot +
   labs(title = "Coefficiency of cancer type",
@@ -48,19 +103,3 @@ typePlot +
        y = "Cancer type") +
   scale_colour_Publication() +
[…]

### 2020-04-08 prediction.R

@@ -59,6 +59,7 @@ typeCoeff <- tidy(extract_model(wfl_final)) %>%
   filter(lambda > 0.09 & lambda < 0.104) %>%
   arrange(-abs(estimate)) %>%
   filter(grepl("type_", term))
+typeCoeff$term <- gsub("type_", "", typeCoeff$term)
 
 typePlot <- typeCoeff %>%
   ggplot(aes(x=reorder(term, -estimate), y=estimate)) +
@@ -145,6 +146,22 @@ mutatioPlot <- mutationRate %>%
   geom_smooth(method='lm', formula= y~x) +
   ggrepel::geom_text_repel(data = mutationRate, aes(label = type))
 
+
+mutationRatePR <- trainset %>%
+  select(type, variant) %>%
+  group_by(type) %>%
+  count(variant == "Mutant") %>%
+  rename(variant = 'variant == "Mutant"') %>%
+  spread(variant, n) %>%
+  mutate(rate = `TRUE`/(`TRUE`+`FALSE`)) %>%
+  bind_cols(pr_auc_train_type)
+
+mutatioPlotPR <- mutationRatePR %>%
+  ggplot(aes(x=rate, y=.estimate)) +
+  geom_point() +
+  geom_smooth(method='lm', formula= y~x) +
+  ggrepel::geom_text_repel(data = mutationRatePR, aes(label = type))
+
 ## @knitr testset_prediction
 
 test_probs <- 
@@ -223,23 +240,18 @@ pr_test_trainPlot <- pr_train_test %>%
 
 pr_test_trainPlot
 
+roc_train_test <- merge(roc_auc_train_type,
+                       roc_auc_test_type,
+                       by = "type")
 
-## @knitr correlation testset
-
-mutationRateTest <- testset %>%
-  select(type, variant) %>%
-  group_by(type) %>%
-  count(variant == TRUE) %>%
-  rename(variant = 'variant == TRUE') %>%
-  spread(variant, n) %>%
-  mutate(rate = `TRUE`/(`TRUE`+`FALSE`)) %>%
-  bind_cols(roc_auc_test_type)
-
-mutationTestPlot <- mutationRateTest %>%
-  ggplot(aes(x=rate, y=.estimate)) +
+roc_test_trainPlot <- roc_train_test %>%
+  ggplot(aes(x=.estimate.x, y=.estimate.y)) +
   geom_point() +
   geom_smooth(method='lm', formula= y~x) +
[…]

### 2020-04-08 plosOne/S1Appendix.Rmd
figure legend

---
title: "Supplimentary"
author: "JK"
date: '2020 4 3 '
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

Data repository

Pancancer TCGA RNA data: http://api.gdc.cancer.gov/data/3586c0da-64d0-4b74-a449-5ff4d9136611

```{r TCGA abbreviations}
library(tidyverse)

read_delim("../TCGAAbbreviations.txt", delim = "\t") %>%
  knitr::kable()

```

### 2020-04-08 plosOne/plosOne.Rmd
aupr
figure legend

@@
 abstract: |
-  Lorem ipsum dolor sit amet, consectetur adipiscing elit.
-Curabitur eget porta erat.
-Morbi consectetur est vel gravida pretium.
-Suspendisse ut dui eu ante cursus gravida non sed sem.
-Nullam sapien tellus, commodo id velit id, eleifend volutpat quam.
-Phasellus mauris velit, dapibus finibus elementum vel, pulvinar non tellus.
-Nunc pellentesque pretium diam, quis maximus dolor faucibus id.
-Nunc convallis sodales ante, ut ullamcorper est egestas vitae.
-Nam sit amet enim ultrices, ultrices elit pulvinar, volutpat risus.
 
@@
 knitr::opts_chunk$set(echo = FALSE, message = FALSE)
-```
-
-```{r}
-knitr::read_chunk('../prediction.R')
 ```
@@
 
+The trainset AUROC was 0.93 and the testset AUROC was 0.84. The AUPR of trainset was 0.74 and the testset AUPR was 0.771.
+
+```{r}
+load("../result.RData")
+```
+
 ## performance of each cancer type
@@
 
+# Figure legends
+
+Figure 1. Summary of modeling results
+
+(A) Left: receiver operating characteristic (ROC) curve right: precision recall (PR) curve of trainset and testset.
+The horizontal green line is the PIK3CA mutation rate (0.11)  (B) Correlation between trainset and testset of area under receiver operating characteristic curve (AUROC) and area under precision recall curve (AUPR) among cancer type.
+The abbreviations are explained in S1 appendix.
+(C) Correlation between the PIK3CA mutation rate of area under receiver operating characteristic curve (AUROC) and area under precision recall curve (AUPR).
+
+Figure 2. Coefficients of model
+
+(A) Top 30 high coefficients of mRNA. (B) Coefficients of cancer types. The abbreviations are explained in S1 appendix.
+
+# Supporting information
+
+S1 Appendix. 
+S2 Figure.
+S1 Table.  
+S2 Table.  
+S3 Table.
+
 # References {#references .unnumbered}
-
-# Figure legends

### 2020-04-08 plosOne/plosOne.Rmd
aupr figure

@@
     corresponding: jkang.alien@gmail.com
-
 abstract: |
-
 bibliography: mybibfile.bib
@@
 
+PIK3CA prediction model performed better than RAS activation prediction model in terms of AUROC. The overall AUPR was
+
 Conclusions comments on possible biological or clinical implications and suggestions for further research.  

### 2020-04-08 plosOne/plosOne.tex
aupr figure

@@
 % Please keep the abstract below 300 words
-\section*{Abstract}
-Lorem ipsum dolor sit amet, consectetur adipiscing elit. Curabitur eget
-porta erat. Morbi consectetur est vel gravida pretium. Suspendisse ut
-dui eu ante cursus gravida non sed sem. Nullam sapien tellus, commodo id
-velit id, eleifend volutpat quam. Phasellus mauris velit, dapibus
-finibus elementum vel, pulvinar non tellus. Nunc pellentesque pretium
-diam, quis maximus dolor faucibus id. Nunc convallis sodales ante, ut
-ullamcorper est egestas vitae. Nam sit amet enim ultrices, ultrices elit
-pulvinar, volutpat risus.
 
@@
 
+The trainset AUROC was 0.93 and the testset AUROC was 0.84. The AUPR of
+trainset was 0.74 and the testset AUPR was 0.771.
+
 \hypertarget{performance-of-each-cancer-type}{%
@@
 
+\hypertarget{figure-legends}{%
+\section{Figure legends}\label{figure-legends}}
+
+Figure 1. Summary of modeling results
+
+\begin{enumerate}
+\def\labelenumi{(\Alph{enumi})}
+\tightlist
+\item
+  Left: receiver operating characteristic (ROC) curve right: precision
+  recall (PR) curve of trainset and testset. The horizontal green line
+  is the PIK3CA mutation rate (0.11) (B) Correlation between trainset
+  and testset of area under receiver operating characteristic curve
+  (AUROC) and area under precision recall curve (AUPR) among cancer
+  type. The abbreviations are explained in S1 appendix. (C) Correlation
+  between the PIK3CA mutation rate of area under receiver operating
+  characteristic curve (AUROC) and area under precision recall curve
+  (AUPR).
+\end{enumerate}
+
+Figure 2. Coefficients of model
+
+\begin{enumerate}
+\def\labelenumi{(\Alph{enumi})}
+\tightlist
+\item
+  Top 30 high coefficients of mRNA. (B) Coefficients of cancer types.
+  The abbreviations are explained in S1 appendix.
+\end{enumerate}
+
+\hypertarget{supporting-information}{%
+\section{Supporting information}\label{supporting-information}}
+
+S1 Appendix. S2 Figure. S1 Table.\\
+S2 Table.\\
+S3 Table.
+
 \hypertarget{references}{%
@@
 \addcontentsline{toc}{section}{References}
-
-\hypertarget{figure-legends}{%
-\section*{Figure legends}\label{figure-legends}}
-\addcontentsline{toc}{section}{Figure legends}
 

### 2020-04-08 figure.R

@@ -52,18 +52,21 @@ cor_PR <- pr_test_trainPlot +
   scale_colour_Publication() +
   theme_Publication()
 
-mutationROC <- mutatioPlot +
+mutationROC <- mutatioPlotTest +
   labs(title = "AUROC",
        x = "Mutation rate",
        y = "ROC") +
   scale_colour_Publication() +
   theme_Publication()
 
-mutationPR <- mutatioPlotPR +
+mutationPR <- mutatioPlotPRTest +
   labs(title = "AUPR",
        x = "Mutation rate",
        y = "PR") +
+  scale_x_continuous(limits = c(0, 0.6)) +
+  scale_y_continuous(limits = c(0, 0.6)) +
   scale_colour_Publication() +
+  geom_abline(slope = 1, intercept = 0, color = 3) +
   theme_Publication()
 
 geneCoeff <- coeffPlot +

### 2020-04-08 prediction.R

@@ -162,6 +162,39 @@ mutatioPlotPR <- mutationRatePR %>%
   geom_smooth(method='lm', formula= y~x) +
   ggrepel::geom_text_repel(data = mutationRatePR, aes(label = type))
 
+## @knitr testset
+
+mutationRateTest <- testset %>%
+  select(type, variant) %>%
+  group_by(type) %>%
+  count(variant == "Mutant") %>%
+  rename(variant = 'variant == "Mutant"') %>%
+  spread(variant, n) %>%
+  mutate(rate = `TRUE`/(`TRUE`+`FALSE`)) %>%
+  bind_cols(roc_auc_test_type)
+
+mutatioPlotTest <- mutationRateTest %>%
+  ggplot(aes(x=rate, y=.estimate)) +
+  geom_point() +
+  geom_smooth(method='lm', formula= y~x) +
+  ggrepel::geom_text_repel(data = mutationRateTest, aes(label = type))
+
+
+mutationRatePRTest <- testset %>%
+  select(type, variant) %>%
+  group_by(type) %>%
+  count(variant == "Mutant") %>%
+  rename(variant = 'variant == "Mutant"') %>%
+  spread(variant, n) %>%
+  mutate(rate = `TRUE`/(`TRUE`+`FALSE`)) %>%
+  bind_cols(pr_auc_test_type)
+
+mutatioPlotPRTest <- mutationRatePRTest %>%
+  ggplot(aes(x=rate, y=.estimate)) +
+  geom_point() +
+  #geom_smooth(method='lm', formula= y~x) +
+  ggrepel::geom_text_repel(data = mutationRatePRTest, aes(label = type))
+
 ## @knitr testset_prediction
 
 test_probs <- 

### 2020-04-08 plosOne/plosOne.tex
pdf

@@
 
+PIK3CA prediction model performed better than RAS activation prediction
+model in terms of AUROC. AUPR and the PIK3CA mutation rate
+interpretation. {[}3{]}
+
 Conclusions comments on possible biological or clinical implications and
@@
 
+\leavevmode\hypertarget{ref-davis_2006_relationship_proc23rdintconfmachlearn}{}%
+3. Davis J, Goadrich M. The relationship between Precision-Recall and
+ROC curves. Proceedings of the 23rd international conference on Machine
+learning. Pittsburgh, Pennsylvania, USA: Association for Computing
+Machinery; 2006. pp. 233--240.
+doi:\href{https://doi.org/10.1145/1143844.1143874}{10.1145/1143844.1143874}
+
 \nolinenumbers

### 2020-04-22 plosOne/plosOne.Rmd
pdf
rpackage citations

@@
 We used TCGA pancancer dataset.
-PIK3CA mutation data was get using cgdsr rpackage.
-Gene expression data was get from GDAC firehose using RTCGAToolbox R package.
-ER immunostain postive HER2 immunostain negative and/or SISH negative breast cancers are included.
+PIK3CA mutation data was get using cgdsr rpackage.[@jacobsenCgdsrRbasedAPI2019] Gene expression data was get from GDAC firehose using RTCGAToolbox R package.
+[@samurRTCGAToolboxNewTool2014] ER immunostain postive HER2 immunostain negative and/or SISH negative breast cancers are included.
 Data of invasive ductal carcinomas were used for training set and data of invasive lobular carcinoma were used for test set.
@@
 
-## Selecting variable for modeling
+## Modeling
 
@@
 
-## Preprocessing
-
-Yeo-Johnson transformation was done to correct skewness. Centering and scaling were done. All preprocessing was done using recipe r package.
-
-## Modeling
+Yeo-Johnson transformation was done to correct skewness. Centering and scaling were done. All preprocessing was done using recipe r package. [@kuhnRecipesPreprocessingTools2020]
 
@@
 Model performance was evealuated with the area under receiver operating characteristic (ROC) and area under the ROC (AUROC) and precision recall (PR) and area under curve (AUPR).
+The modelin process was done with tidymodels rpackage.
+[@kuhnTidymodelsEasilyInstall2020]
 
@@
 
-The trainset AUROC was 0.93 and the testset AUROC was 0.84. The AUPR of trainset was 0.74 and the testset AUPR was 0.771.
-
-```{r}
-load("../result.RData")
-```
+The trainset AUROC was 0.93 and the testset AUROC was 0.84. The AUPR of trainset was 0.66 and the testset AUPR was 0.39. (Figure 1A)
 
 ## performance of each cancer type  
+Because tht prevalence of PIK3CA mutation is vary across the cancer type, the performance of each cancer type was investigated.
+The AUROC and AUPR were positively correlated between train set and test set in cancer type subanalysis.
+(Figure 1B) The AUPR was high in cancer type with high PIK3CA mutation rate such as colon, brest, Uterus cancer types.
+The AUROC did not correlated with PIK3CA mutation rate of each cancer types.
+(Figure 1C)
 
@@
 
-PIK3CA prediction model performed better than RAS activation prediction model in terms of AUROC. The overall AUPR was
+PIK3CA prediction model performed better than RAS activation prediction model in terms of AUROC. AUPR and the PIK3CA mutation rate interpretation. @davis_2006_relationship_proc23rdintconfmachlearn
 

### 2020-04-22 plosOne/plosOne.tex
rpackage citations

@@
 We used TCGA pancancer dataset. PIK3CA mutation data was get using cgdsr
-rpackage. Gene expression data was get from GDAC firehose using
-RTCGAToolbox R package. ER immunostain postive HER2 immunostain negative
-and/or SISH negative breast cancers are included. Data of invasive
-ductal carcinomas were used for training set and data of invasive
-lobular carcinoma were used for test set. Number of observations were
-530 in training set and 188 in test set.
+rpackage.{[}3{]} Gene expression data was get from GDAC firehose using
+RTCGAToolbox R package. {[}4{]} ER immunostain postive HER2 immunostain
+negative and/or SISH negative breast cancers are included. Data of
+invasive ductal carcinomas were used for training set and data of
+invasive lobular carcinoma were used for test set. Number of
+observations were 530 in training set and 188 in test set.
 
-\hypertarget{selecting-variable-for-modeling}{%
-\subsection{Selecting variable for
-modeling}\label{selecting-variable-for-modeling}}
+\hypertarget{modeling}{%
+\subsection{Modeling}\label{modeling}}
 
@@
 
-\hypertarget{preprocessing}{%
-\subsection{Preprocessing}\label{preprocessing}}
-
 Yeo-Johnson transformation was done to correct skewness. Centering and
 scaling were done. All preprocessing was done using recipe r package.
-
-\hypertarget{modeling}{%
-\subsection{Modeling}\label{modeling}}
+{[}5{]}
 
@@
 under the ROC (AUROC) and precision recall (PR) and area under curve
-(AUPR).
+(AUPR). The modelin process was done with tidymodels rpackage. {[}6{]}
 
@@
 The trainset AUROC was 0.93 and the testset AUROC was 0.84. The AUPR of
-trainset was 0.74 and the testset AUPR was 0.771.
+trainset was 0.66 and the testset AUPR was 0.39. (Figure 1A)
 
@@
 type}\label{performance-of-each-cancer-type}}
+
+Because tht prevalence of PIK3CA mutation is vary across the cancer
+type, the performance of each cancer type was investigated. The AUROC
+and AUPR were positively correlated between train set and test set in
+cancer type subanalysis. (Figure 1B) The AUPR was high in cancer type
+with high PIK3CA mutation rate such as colon, brest, Uterus cancer
+types. The AUROC did not correlated with PIK3CA mutation rate of each
+cancer types. (Figure 1C)
 
@@
 model in terms of AUROC. AUPR and the PIK3CA mutation rate
-interpretation. {[}3{]}
+interpretation. {[}7{]}
 
@@
 
-S1 Appendix. S2 Figure. S1 Table.\\
+S1 Appendix.\\
+S2 Figure.\\
+S1 Table.\\
 S2 Table.\\
@@
 
+\leavevmode\hypertarget{ref-jacobsenCgdsrRbasedAPI2019}{}%
+3. Jacobsen A, Luna A. Cgdsr: R-based API for accessing the MSKCC cancer
+genomics data server (CGDS). 2019.
+
+\leavevmode\hypertarget{ref-samurRTCGAToolboxNewTool2014}{}%
+4. Samur MK. RTCGAToolbox: A new tool for exporting TCGA Firehose data.
+PLoS One. 2014;9(9):e106397.
+
+\leavevmode\hypertarget{ref-kuhnRecipesPreprocessingTools2020}{}%
+5. Kuhn M, Wickham H. Recipes: Preprocessing tools to create design
+matrices. 2020.
+
+\leavevmode\hypertarget{ref-kuhnTidymodelsEasilyInstall2020}{}%
+6. Kuhn M, Wickham H. Tidymodels: Easily install and load the
+'tidymodels' packages. 2020.
+
 \leavevmode\hypertarget{ref-davis_2006_relationship_proc23rdintconfmachlearn}{}%
-3. Davis J, Goadrich M. The relationship between Precision-Recall and
+7. Davis J, Goadrich M. The relationship between Precision-Recall and
 ROC curves. Proceedings of the 23rd international conference on Machine

### 2020-04-22 plosOne/plosOne.Rmd
discussion RAS

@@
 abstract: |
-bibliography: mybibfile.bib
+bibliography: [mybibfile.bib, pik3ca.bib] 
+indent: true
 output: rticles::plos_article
@@
 
-A list
-
 - Brevity
@@
 
-The problem
-
-- PIK3CA mutation in selecting drug
+- The problem  
+  * PIK3CA mutation in selecting drug
 
@@
 
-RNAseq can be widely used. The mutation status directly prediction
+- RNAseq can be widely used. The mutation status directly prediction
 
-Previous study @way_2018_machine_cellreports
+- Previous study @way_2018_machine_cellreports
 
-prediction of aberant activation of a certain pathway vs a specific mutation
-The proposed solution
+- Prediction of aberant activation of a certain pathway vs a specific mutation
+
+- The proposed solution
 
@@
 
-To narrow down potential predictors, Genes with a large the median absolute deviation (more than third-quartiles) were selected. 5000 out of 20502 genes were included in the modeling process.
-
-Yeo-Johnson transformation was done to correct skewness. Centering and scaling were done. All preprocessing was done using recipe r package. [@kuhnRecipesPreprocessingTools2020]
-
-Penalized logistic regression was applied to prediction modeling.
+To narrow down potential predictors, Genes with a large the median absolute deviation (more than third-quartiles) were selected.
+5000 out of 20502 genes were included in the modeling process.
+Yeo-Johnson transformation was done to correct skewness.
+Centering and scaling were done.
+All preprocessing was done using recipe r package.
+[@kuhnRecipesPreprocessingTools2020] Penalized logistic regression was applied to prediction modeling.
 10-fold cross-validation with targe variable stratification was done over the hyperparameter grid: $\lambda$ {$10^{-5}$, $10^{-4}$,$10^{-3}$,$10^{-2}$,$10^{-1}$, $10^{0}$}, $\alpha$ {0.0, 0.25, 0.5, 0.75}.
@@
 
-## selecting model
+## Selecting model
 
@@
 
-## performance of each cancer type  
+## Performance of each cancer type  
 Because tht prevalence of PIK3CA mutation is vary across the cancer type, the performance of each cancer type was investigated.
@@
 
+Figure 2 shows top 30 important predictors.
+IGF1R mRNA expression was the strongest negative predictor and PTEN was the strongest positive predictor.
+Both genes and PIK3CA are key players in tyrosin kanase pathway.
+The cancer type was important predictors.
+Some cancer types including uterine carcinosarcoma (UCS), bladder urothelial carcinoma (BLCA), pancreatic adenocarcinoma (PAAD), lymphoid neoplasm diffuse large B-cell lymphoma (DLBC) are strongest predictors.
+
 # Discussion
@@
 
-Main message answers the question and main supporting evidence
+- Main message answers the question and main supporting evidence
 
-Critical assessment opinions on 
-- any shortcomings in study design
+- Critical assessment opinions on any shortcomings in study design
 
@@
 
+In this study, the prevalence of PIK3CA mutation was 11%.
+The low prevalence of PIK3CA results in imbalaced dataset which makes the prediction difficult.
+The AUROC is optimistic in terms of performance.
+The AUPR is more informative than AUROC on imbalaced datasets.
+[@saitoPrecisionRecallPlotMore2015]
+
 - limitations in methods
 
-Case imbalance
+  * Case imbalance  
+  * flaws in analysis  
+  * validity of assumption
 
-- flaws in analysis
+- Comparison with other studies where inconsistencies are discussed  
+- Evaluate the results - not the authors
 
-- validity of assumption
+Our PIK3CA prediction model performed better than RAS activation prediction model of previous study in terms of both AUROC (0.84 vs 0.75) and AUPR (0.39 vs 0.24) on testset prediction.
+Our testset is corresponding to the samples initially filtered from training.
+The target variable of our study is more specific than the previous study.
+The specific important mutations can effect stronger downsteam gene expression than the broad events pathway activation.
+The previous study might be more difficult prediction problum.
 
-Comparison with other studies where inconsistencies are discussed
-
-PIK3CA prediction model performed better than RAS activation prediction model in terms of AUROC. AUPR and the PIK3CA mutation rate interpretation. @davis_2006_relationship_proc23rdintconfmachlearn
-
-Conclusions comments on possible biological or clinical implications and suggestions for further research.  
-Evaluate the results - not the authors
+- Conclusions comments on possible biological or clinical implications and suggestions for further research.
 
@@
 
-Figure 1. Summary of modeling results
+- Figure 1. Summary of modeling results
 
@@
 
-Figure 2. Coefficients of model
+- Figure 2. Coefficients of model
 
@@
 
-S1 Appendix.  
-S2 Figure.  
-S1 Table.  
-S2 Table.  
-S3 Table.
+- S1 Appendix.  
+- S2 Figure.  
+- S1 Table.  
+- S2 Table.  
+- S3 Table.
 

### 2020-04-22 plosOne/plosOne.tex
discussion RAS

@@
 
-A list
-
 \begin{itemize}
-\tightlist
 \item
@@
   Clean typing
-\end{itemize}
-
+\item
   The problem
@@
     PIK3CA mutation in selecting drug
+  \end{itemize}
 \end{itemize}
@@
 
+\begin{itemize}
+\item
   RNAseq can be widely used. The mutation status directly prediction
-
+\item
   Previous study {[}2{]}
-
-prediction of aberant activation of a certain pathway vs a specific
-mutation The proposed solution
+\item
+  Prediction of aberant activation of a certain pathway vs a specific
+  mutation
+\item
+  The proposed solution
+\end{itemize}
 
@@
 absolute deviation (more than third-quartiles) were selected. 5000 out
-of 20502 genes were included in the modeling process.
-
-Yeo-Johnson transformation was done to correct skewness. Centering and
-scaling were done. All preprocessing was done using recipe r package.
-{[}5{]}
-
+of 20502 genes were included in the modeling process. Yeo-Johnson
+transformation was done to correct skewness. Centering and scaling were
+done. All preprocessing was done using recipe r package. {[}5{]}
 Penalized logistic regression was applied to prediction modeling.
@@
 \hypertarget{selecting-model}{%
-\subsection{selecting model}\label{selecting-model}}
+\subsection{Selecting model}\label{selecting-model}}
 
@@
 \hypertarget{performance-of-each-cancer-type}{%
-\subsection{performance of each cancer
+\subsection{Performance of each cancer
 type}\label{performance-of-each-cancer-type}}
@@
 
+Figure 2 shows top 30 important predictors. IGF1R mRNA expression was
+the strongest negative predictor and PTEN was the strongest positive
+predictor. Both genes and PIK3CA are key players in tyrosin kanase
+pathway. The cancer type was important predictors. Some cancer types
+including uterine carcinosarcoma (UCS), bladder urothelial carcinoma
+(BLCA), pancreatic adenocarcinoma (PAAD), lymphoid neoplasm diffuse
+large B-cell lymphoma (DLBC) are strongest predictors.
+
 \hypertarget{discussion}{%
@@
 
+\begin{itemize}
+\item
   Main message answers the question and main supporting evidence
-
-Critical assessment opinions on - any shortcomings in study design
+\item
+  Critical assessment opinions on any shortcomings in study design
+\end{itemize}
 
@@
 
+In this study, the prevalence of PIK3CA mutation was 11\%. The low
+prevalence of PIK3CA results in imbalaced dataset which makes the
+prediction difficult. The AUROC is optimistic in terms of performance.
+The AUPR is more informative than AUROC on imbalaced datasets. {[}7{]}
+
+\begin{itemize}
+\item
+  limitations in methods
+
   \begin{itemize}
@@
   \item
-  limitations in methods
-\end{itemize}
-
-Case imbalance
-
-\begin{itemize}
+    Case imbalance\\
   \item
-  flaws in analysis
+    flaws in analysis\\
   \item
@@
   \end{itemize}
-
-Comparison with other studies where inconsistencies are discussed
+\item
+  Comparison with other studies where inconsistencies are discussed\\
+\item
+  Evaluate the results - not the authors
+\end{itemize}
 
@@
 
-Conclusions comments on possible biological or clinical implications and
-suggestions for further research.\\
-Evaluate the results - not the authors
+\begin{itemize}
+\tightlist
+\item
+  Conclusions comments on possible biological or clinical implications
+  and suggestions for further research.
+\end{itemize}
 
@@
 
+\begin{itemize}
+\tightlist
+\item
   Figure 1. Summary of modeling results
+\end{itemize}
 
@@
 
+\begin{itemize}
+\tightlist
+\item
   Figure 2. Coefficients of model
+\end{itemize}
 
@@
 
+\begin{itemize}
+\tightlist
+\item
   S1 Appendix.\\
+\item
   S2 Figure.\\
+\item
   S1 Table.\\
+\item
   S2 Table.\\
+\item
   S3 Table.
+\end{itemize}
 
@@
 
-\leavevmode\hypertarget{ref-davis_2006_relationship_proc23rdintconfmachlearn}{}%
-7. Davis J, Goadrich M. The relationship between Precision-Recall and
-ROC curves. Proceedings of the 23rd international conference on Machine
-learning. Pittsburgh, Pennsylvania, USA: Association for Computing
-Machinery; 2006. pp. 233--240.
-doi:\href{https://doi.org/10.1145/1143844.1143874}{10.1145/1143844.1143874}
+\leavevmode\hypertarget{ref-saitoPrecisionRecallPlotMore2015}{}%
+7. Saito T, Rehmsmeier M. The Precision-Recall Plot Is More Informative
+than the ROC Plot When Evaluating Binary Classifiers on Imbalanced
+Datasets. PLOS ONE. Public Library of Science; 10: e0118432.
+doi:\href{https://doi.org/10.1371/journal.pone.0118432}{10.1371/journal.pone.0118432}
 

### 2020-04-23 plosOne/plosOne.Rmd
rpackage.bib

@@
 abstract: |
-bibliography: [mybibfile.bib, pik3ca.bib] 
+bibliography: [mybibfile.bib, pik3ca.bib, rpackage.bib] 
 indent: true
@@
 PIK3CA mutation data was get using cgdsr rpackage.[@jacobsenCgdsrRbasedAPI2019] Gene expression data was get from GDAC firehose using RTCGAToolbox R package.
-[@samurRTCGAToolboxNewTool2014] ER immunostain postive HER2 immunostain negative and/or SISH negative breast cancers are included.
-Data of invasive ductal carcinomas were used for training set and data of invasive lobular carcinoma were used for test set.
-Number of observations were 530 in training set and 188 in test set.
+[@samurRTCGAToolboxNewTool2014]
 
@@
 10-fold cross-validation with targe variable stratification was done over the hyperparameter grid: $\lambda$ {$10^{-5}$, $10^{-4}$,$10^{-3}$,$10^{-2}$,$10^{-1}$, $10^{0}$}, $\alpha$ {0.0, 0.25, 0.5, 0.75}.
-Model performance was evealuated with the area under receiver operating characteristic (ROC) and area under the ROC (AUROC) and precision recall (PR) and area under curve (AUPR).
+Lambda is  penalty scaling parameter and alpha is mixing parameter of penalty function ($(1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1$).
+[@friedmanRegularizationPathsGeneralized2010] Model performance was evealuated with the area under receiver operating characteristic (ROC) and area under the ROC (AUROC) and precision recall (PR) and area under curve (AUPR).
 The modelin process was done with tidymodels rpackage.
@@
 
-## Selecting model
+## Selecting model and performance estimation
 
-Cross-validation  
-The model showed best performance at lambda = 0.01 and alpha = 1.0 (Ridge regression).
-
-## Prediction performance
+In 10-fold cross-validation, the model with lambda = 0.01 and alpha = 1.0 (Ridge regression) were best performance in terms of AUROC.
+The final model was trained with the selected hyperparameters with all trainset.
 
@@
 ## Performance of each cancer type
+
 Because tht prevalence of PIK3CA mutation is vary across the cancer type, the performance of each cancer type was investigated.
@@
 
+OUr model includes cancer type predictor and they are stronger than gene expression data.
+The varing prevalence of PIC3CA mutation across cancer type might reason of the strong cancer type predictor.
+If the cancer type was wrong or can not be determined, our model performance can be poor.
+
 - Conclusions comments on possible biological or clinical implications and suggestions for further research.
+
+- Proof of concept study  
+- Reproducibility of gene expression prediction model
 

### 2020-04-23 plosOne/plosOne.tex
rpackage.bib

@@
 rpackage.{[}3{]} Gene expression data was get from GDAC firehose using
-RTCGAToolbox R package. {[}4{]} ER immunostain postive HER2 immunostain
-negative and/or SISH negative breast cancers are included. Data of
-invasive ductal carcinomas were used for training set and data of
-invasive lobular carcinoma were used for test set. Number of
-observations were 530 in training set and 188 in test set.
+RTCGAToolbox R package. {[}4{]}
 
@@
 \(10^{-4}\),\(10^{-3}\),\(10^{-2}\),\(10^{-1}\), \(10^{0}\)\},
-\(\alpha\) \{0.0, 0.25, 0.5, 0.75\}. Model performance was evealuated
-with the area under receiver operating characteristic (ROC) and area
-under the ROC (AUROC) and precision recall (PR) and area under curve
-(AUPR). The modelin process was done with tidymodels rpackage. {[}6{]}
+\(\alpha\) \{0.0, 0.25, 0.5, 0.75\}. Lambda is penalty scaling parameter
+and alpha is mixing parameter of penalty function
+(\((1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1\)).
+Model performance was evealuated with the area under receiver operating
+characteristic (ROC) and area under the ROC (AUROC) and precision recall
+(PR) and area under curve (AUPR). The modelin process was done with
+tidymodels rpackage. {[}6{]}
 
@@
 
-\hypertarget{selecting-model}{%
-\subsection{Selecting model}\label{selecting-model}}
+\hypertarget{selecting-model-and-performance-estimation}{%
+\subsection{Selecting model and performance
+estimation}\label{selecting-model-and-performance-estimation}}
 
-Cross-validation\\
-The model showed best performance at lambda = 0.01 and alpha = 1.0
-(Ridge regression).
-
-\hypertarget{prediction-performance}{%
-\subsection{Prediction performance}\label{prediction-performance}}
+In 10-fold cross-validation, the model with lambda = 0.01 and alpha =
+1.0 (Ridge regression) were best performance in terms of AUROC. The
+final model was trained with the selected hyperparameters with all
+trainset.
 
@@
 
-PIK3CA prediction model performed better than RAS activation prediction
-model in terms of AUROC. AUPR and the PIK3CA mutation rate
-interpretation. {[}7{]}
+Our PIK3CA prediction model performed better than RAS activation
+prediction model of previous study in terms of both AUROC (0.84 vs 0.75)
+and AUPR (0.39 vs 0.24) on testset prediction. Our testset is
+corresponding to the samples initially filtered from training. The
+target variable of our study is more specific than the previous study.
+The specific important mutations can effect stronger downsteam gene
+expression than the broad events pathway activation. The previous study
+might be more difficult prediction problum.
+
+OUr model includes cancer type predictor and they are stronger than gene
+expression data. The varing prevalence of PIC3CA mutation across cancer
+type might reason of the strong cancer type predictor. If the cancer
+type was wrong or can not be determined, our model performance can be
+poor.
 
 \begin{itemize}
-\tightlist
 \item
@@
   and suggestions for further research.
+\item
+  Proof of concept study\\
+\item
+  Reproducibility of gene expression prediction model
 \end{itemize}

### 2020-04-23 plosOne/plosOne.Rmd
prevalance of pik3ca

@@
 
-## Modeling
+10845 cases were available both PIK3CA mutation and mRNAD expression data.
+5128 out of 20502 genes were included in the modeling process after filtering with median absolute deviation as described at modelinf process method.
+33 cancer type dummy variables were included in predictor variables.
+
+The target variable was PIK3CA mutation status.
+The status of  PIK3CA was considered as positive when the case has following PIK3CA variants which is the target variables of the therascreen PIK3CA RGQ PCR Kit; C420R, E542K, E545A, E545D, E545G, E545K, Q546E, Q546R, H1047L, H1047R, H1047Y.
+The therascreen PIK3CA RGQ PCR Kit was approved by U.S.
+Food and Drug adminidtration for PIK3CA inhibitor.
+We splited the three quarters of dataset for the trainset and one quarter for testset.
+
+## Modeling process
 
 To narrow down potential predictors, Genes with a large the median absolute deviation (more than third-quartiles) were selected.
-5000 out of 20502 genes were included in the modeling process.
 Yeo-Johnson transformation was done to correct skewness.
@@
 # Results
+
+## Dataset summary 
+- PIK3CA prevalence
+-
 
@@
 Figure 2 shows top 30 important predictors.
+The coefficient is the parameter of the predictor which represent the effect of the predictor on prediction.
 IGF1R mRNA expression was the strongest negative predictor and PTEN was the strongest positive predictor.

### 2020-04-23 plosOne/plosOne.tex
prevalance of pik3ca

@@
 
-\hypertarget{modeling}{%
-\subsection{Modeling}\label{modeling}}
+10845 cases were available both PIK3CA mutation and mRNAD expression
+data. 5128 out of 20502 genes were included in the modeling process
+after filtering with median absolute deviation as described at modelinf
+process method. 33 cancer type dummy variables were included in
+predictor variables.
+
+The target variable was PIK3CA mutation status. The status of PIK3CA was
+considered as positive when the case has following PIK3CA variants which
+is the target variables of the therascreen PIK3CA RGQ PCR Kit; C420R,
+E542K, E545A, E545D, E545G, E545K, Q546E, Q546R, H1047L, H1047R, H1047Y.
+The therascreen PIK3CA RGQ PCR Kit was approved by U.S. Food and Drug
+adminidtration for PIK3CA inhibitor. We splited the three quarters of
+dataset for the trainset and one quarter for testset.
+
+\hypertarget{modeling-process}{%
+\subsection{Modeling process}\label{modeling-process}}
 
 To narrow down potential predictors, Genes with a large the median
-absolute deviation (more than third-quartiles) were selected. 5000 out
-of 20502 genes were included in the modeling process. Yeo-Johnson
-transformation was done to correct skewness. Centering and scaling were
-done. All preprocessing was done using recipe r package. {[}5{]}
-Penalized logistic regression was applied to prediction modeling.
-10-fold cross-validation with targe variable stratification was done
-over the hyperparameter grid: \(\lambda\) \{\(10^{-5}\),
+absolute deviation (more than third-quartiles) were selected.
+Yeo-Johnson transformation was done to correct skewness. Centering and
+scaling were done. All preprocessing was done using recipe r package.
+{[}5{]} Penalized logistic regression was applied to prediction
+modeling. 10-fold cross-validation with targe variable stratification
+was done over the hyperparameter grid: \(\lambda\) \{\(10^{-5}\),
 \(10^{-4}\),\(10^{-3}\),\(10^{-2}\),\(10^{-1}\), \(10^{0}\)\},
@@
 (\((1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1\)).
-Model performance was evealuated with the area under receiver operating
-characteristic (ROC) and area under the ROC (AUROC) and precision recall
-(PR) and area under curve (AUPR). The modelin process was done with
-tidymodels rpackage. {[}6{]}
+{[}6{]} Model performance was evealuated with the area under receiver
+operating characteristic (ROC) and area under the ROC (AUROC) and
+precision recall (PR) and area under curve (AUPR). The modelin process
+was done with tidymodels rpackage. {[}7{]}
 
@@
 \section{Results}\label{results}}
+
+\hypertarget{dataset-summary}{%
+\subsection{Dataset summary}\label{dataset-summary}}
+
+\begin{itemize}
+\item
+  PIK3CA prevalence
+\item
+\end{itemize}
 
@@
 
-Figure 2 shows top 30 important predictors. IGF1R mRNA expression was
-the strongest negative predictor and PTEN was the strongest positive
-predictor. Both genes and PIK3CA are key players in tyrosin kanase
-pathway. The cancer type was important predictors. Some cancer types
-including uterine carcinosarcoma (UCS), bladder urothelial carcinoma
-(BLCA), pancreatic adenocarcinoma (PAAD), lymphoid neoplasm diffuse
-large B-cell lymphoma (DLBC) are strongest predictors.
+Figure 2 shows top 30 important predictors. The coefficient is the
+parameter of the predictor which represent the effect of the predictor
+on prediction. IGF1R mRNA expression was the strongest negative
+predictor and PTEN was the strongest positive predictor. Both genes and
+PIK3CA are key players in tyrosin kanase pathway. The cancer type was
+important predictors. Some cancer types including uterine carcinosarcoma
+(UCS), bladder urothelial carcinoma (BLCA), pancreatic adenocarcinoma
+(PAAD), lymphoid neoplasm diffuse large B-cell lymphoma (DLBC) are
+strongest predictors.
 
@@
 prediction difficult. The AUROC is optimistic in terms of performance.
-The AUPR is more informative than AUROC on imbalaced datasets. {[}7{]}
+The AUPR is more informative than AUROC on imbalaced datasets. {[}8{]}
 
@@
 
+\leavevmode\hypertarget{ref-friedmanRegularizationPathsGeneralized2010}{}%
+6. Friedman J, Hastie T, Tibshirani R. Regularization Paths for
+Generalized Linear Models via Coordinate Descent. J Stat Softw. 2010;33:
+1--22.
+
 \leavevmode\hypertarget{ref-kuhnTidymodelsEasilyInstall2020}{}%
-6. Kuhn M, Wickham H. Tidymodels: Easily install and load the
+7. Kuhn M, Wickham H. Tidymodels: Easily install and load the
 'tidymodels' packages. 2020.
@@
 \leavevmode\hypertarget{ref-saitoPrecisionRecallPlotMore2015}{}%
-7. Saito T, Rehmsmeier M. The Precision-Recall Plot Is More Informative
+8. Saito T, Rehmsmeier M. The Precision-Recall Plot Is More Informative
 than the ROC Plot When Evaluating Binary Classifiers on Imbalanced

### 2020-04-23 figure.R

@@ -1,13 +1,47 @@
 load('result.RData')
-
+load('dataset.RData')
 library(Cairo)
 library(tidyverse)
 library(tidymodels)
 library(gridExtra)
 
 source("theme_publication.R")
+### PIK3CA prevalence across cancer type
+prevalancePlot <- dataset %>%
+  select(variant, type) %>%
+  group_by(type) %>%
+  count(variant) %>%
+  spread(variant, n) %>%
+  replace_na(list(Mutant=0)) %>%
+  mutate(prevalance = Mutant/(Mutant + Wild)) %>%
+  ggplot(aes(x=reorder(type, -prevalance), y=prevalance)) +
+  geom_bar(stat = "identity") +
+  coord_flip() + 
+  theme_Publication() +
+  labs(title = "PIK3CA Prevalance",
+       x = "Prevalance",
+       y = "Cancer type")
 
+CairoPDF(file = 'Figure1',
+         width = 4, height = 6, pointsize=8)
+dataset %>%
+  select(variant, type) %>%
+  group_by(type) %>%
+  count(variant) %>%
+  spread(variant, n) %>%
+  replace_na(list(Mutant=0)) %>%
+  mutate(prevalance = Mutant/(Mutant + Wild)) %>%
+  ggplot(aes(x=reorder(type, -prevalance), y=prevalance)) +
+  geom_bar(stat = "identity") +
+  coord_flip() + 
+  theme_Publication() +
+  labs(title = "PIK3CA Prevalance",
+       x = "Prevalance",
+       y = "Cancer type")
+
+dev.off()
 ### ROC PR Plots
+  
 ROCPlot <- bind_rows("Trainset" = train_probs,
           "Testset" = test_probs,
           .id = "dataset") %>%
@@ -21,9 +55,6 @@ ROCPlot <- bind_rows("Trainset" = train_probs,
   scale_colour_Publication() +
   theme_Publication()
   
-
-
-
 prPlot <-  bind_rows("Trainset" = train_probs,
                      "Testset" = test_probs,
                      .id = "dataset") %>%
@@ -84,7 +115,7 @@ typeCoeff <- typePlot +
   theme_Publication()
 
 
-CairoPDF(file = 'Figure1',
+CairoPDF(file = 'Figure2',
          width = 8, height = 12, pointsize=10)
 grid.arrange(ROCPlot, prPlot,
              cor_ROC, cor_PR,
[…]

### 2020-05-07 plosOne/plosOne.Rmd
introduction machine learning

@@
 
+Targeted thrapy becomes standard treatment in many cancer patients.
+To treat with the targeted therapy, the The patients should be confirmed the indicated genomic alterations in their cancer.
+Many direct test for the genomic alteration has been developed and prooved for their clical utility to discriminate which patients will be response the targeted therapy.
+
+The machine learning approach has been actively researched recently to detect the genomic alterations.
+The machine learning can build a prediction model from a large number of predictors such as radiolomic features @dercleIdentificationNonSmall2020, pathology image or gene expression data @wayMachineLearningDetects2018.
+The machine learning approach has limitations to apply to clinical practice.
+However the machine learning prediction can be a second best when the direct test fails.
+
+PIK3CA is encodes the p110$\alpha$ catalytic subunit of phosphatidylinositol 3′-kinase (PI3K).
+PI3K is a protein kinase which phosphorylates phosphatidylinositol 4,5-biphosphate (PIP~2~) to make phosphatidylinositol 3,4,5-triphosphate (PIP~3~).
+Phosphatase and tensin homolog (PTEN) changes PIP~2~ to PIP~3~ in contrast PI3K.
+PIP~3~ is a second mesenger to activate protein kinase B (AKT) which is a serine/threonine-specific protein kinase.
+AKT inhibits apoptosis and promote cell proliferation.
+
 Breast cancer with PIK3CA mutation has been approved to use PIK3CA inhibitor in hormone receptor positive HER2 negative subtype.
-@andre_2019_alpelisib_nengljmeda Prediction of PIK3CA mutation was done by gene expression data of TCGA.
-
-- Predicting mutation by gene expression data
-  * What for?  
-  * Feasible?  
-  * Cost?
-
-- Varing frquency of PIK3CA mutation across cancer types  
-  * High in endometrial breast  
-  * not common other cancer type
-
-Universial prediction using gene expression data across cancer type for certain kind of mutation.
-It's clinically useless now, but we want to explore the possibilities of the PIK3CA mutation prediction.
-
-- RNAseq can be widely used. The mutation status directly prediction
-
-- Previous study @way_2018_machine_cellreports
-
-- Prediction of aberant activation of a certain pathway vs a specific mutation
-
-- The proposed solution
+@andre_2019_alpelisib_nengljmeda The PIK3CA mutation is second most driver mutation after TP53.
+The PIK3CA mutation is most frequantly founded in endometrial carcinoma (45%), and followed by breast invasive carcinoma (24%), cervical squamous cell carcinoma and endocervical adenocarcinoma (20%) and colon adenocarcinoma (16%).
 
@@
 
-10845 cases were available both PIK3CA mutation and mRNAD expression data.
+10845 cases were available both PIK3CA mutation and mRNA expression data.
 5128 out of 20502 genes were included in the modeling process after filtering with median absolute deviation as described at modelinf process method.
@@
 ## Dataset summary 
-- PIK3CA prevalence
--
+Prevalcence of PIK3CA was 0.11 in all cases. The PIK3CA prevalance of each cancer type was vary. The median prevalence of PIK3CA of each cancer types  was 0.03 (range 0-0.33) (Figure 1).
 

### 2020-05-07 plosOne/plosOne.tex
introduction machine learning

@@
 
-Breast cancer with PIK3CA mutation has been approved to use PIK3CA
-inhibitor in hormone receptor positive HER2 negative subtype. {[}1{]}
-Prediction of PIK3CA mutation was done by gene expression data of TCGA.
+Targeted thrapy becomes standard treatment in many cancer patients. To
+treat with the targeted therapy, the The patients should be confirmed
+the indicated genomic alterations in their cancer. Many direct test for
+the genomic alteration has been developed and prooved for their clical
+utility to discriminate which patients will be response the targeted
+therapy.
+
+The machine learning approach has been actively researched recently to
+detect the genomic alterations. Breast cancer with PIK3CA mutation has
+been approved to use PIK3CA inhibitor in hormone receptor positive HER2
+negative subtype. {[}1{]}
+
+\begin{itemize}
+\tightlist
+\item
+  PIK3CA summary
+\end{itemize}
+
+PIK3CA is encodes the p110\(\alpha\) catalytic subunit of
+phosphatidylinositol 3′-kinase (PI3K). PI3K is a protein kinase which
+phosphorylates phosphatidylinositol 4,5-biphosphate
+(PIP\textsubscript{2}) to make phosphatidylinositol 3,4,5-triphosphate
+(PIP\_\{3\}). Phosphatase and tensin homolog (PTEN) changes PIP\_\{2\}
+to PIP\_\{3\} in contrast PI3K. PIP\_\{3\} is a second mesenger to
+activate protein kinase B (AKT) which is a serine/threonine-specific
+protein kinase. AKT inhibits apoptosis and promote cell proliferation.
+
+The PIK3CA mutation is second most driver mutation after TP53. the
+PIK3CA mutation is most frequantly founded in endometrial carcinoma
+(45\%), and followed by breast invasive carcinoma (24\%), cervical
+squamous cell carcinoma and endocervical adenocarcinoma (20\%) and colon
+adenocarcinoma (16\%).
 
@@
 
-10845 cases were available both PIK3CA mutation and mRNAD expression
+10845 cases were available both PIK3CA mutation and mRNA expression
 data. 5128 out of 20502 genes were included in the modeling process
@@
 
-\begin{itemize}
-\item
-  PIK3CA prevalence
-\item
-\end{itemize}
+Prevalcence of PIK3CA was 0.11 in all cases. The PIK3CA prevalance of
+each cancer type was vary. The median prevalence of PIK3CA of each
+cancer types was 0.03 (range 0-0.33) (Figure 1).
 

### 2020-05-07 figure.R

@@ -7,20 +7,6 @@ library(gridExtra)
 
 source("theme_publication.R")
 ### PIK3CA prevalence across cancer type
-prevalancePlot <- dataset %>%
-  select(variant, type) %>%
-  group_by(type) %>%
-  count(variant) %>%
-  spread(variant, n) %>%
-  replace_na(list(Mutant=0)) %>%
-  mutate(prevalance = Mutant/(Mutant + Wild)) %>%
-  ggplot(aes(x=reorder(type, -prevalance), y=prevalance)) +
-  geom_bar(stat = "identity") +
-  coord_flip() + 
-  theme_Publication() +
-  labs(title = "PIK3CA Prevalance",
-       x = "Prevalance",
-       y = "Cancer type")
 
 CairoPDF(file = 'Figure1',
          width = 4, height = 6, pointsize=8)
@@ -33,13 +19,25 @@ dataset %>%
   mutate(prevalance = Mutant/(Mutant + Wild)) %>%
   ggplot(aes(x=reorder(type, -prevalance), y=prevalance)) +
   geom_bar(stat = "identity") +
-  coord_flip() + 
-  theme_Publication() +
+  coord_flip() +
+  theme_Publication() + 
   labs(title = "PIK3CA Prevalance",
        x = "Prevalance",
-       y = "Cancer type")
-
+       y = "Cancer type") 
 dev.off()
+## Prevalence of PIK3CA 
+
+
+dfPrevalance <- dataset %>%
+  select(variant, type) %>%
+  group_by(type) %>%
+  count(variant) %>%
+  spread(variant, n) %>%
+  replace_na(list(Mutant=0)) %>%
+  mutate(prevalance = Mutant/(Mutant + Wild))
+
+summary(dfPrevalance$prevalance)
+summary(dataset$variant)
 ### ROC PR Plots
   
 ROCPlot <- bind_rows("Trainset" = train_probs,

### 2020-05-07 plosOne/plosOne.Rmd
edit 5/7

@@
 Targeted thrapy becomes standard treatment in many cancer patients.
-To treat with the targeted therapy, the The patients should be confirmed the indicated genomic alterations in their cancer.
+Many targeted therapy requires test for a specific cancer genomic alteration, to treat the patients.
 Many direct test for the genomic alteration has been developed and prooved for their clical utility to discriminate which patients will be response the targeted therapy.
 
-The machine learning approach has been actively researched recently to detect the genomic alterations.
-The machine learning can build a prediction model from a large number of predictors such as radiolomic features @dercleIdentificationNonSmall2020, pathology image or gene expression data @wayMachineLearningDetects2018.
-The machine learning approach has limitations to apply to clinical practice.
+Machine learning approach has been actively researched to detect the genomic alterations.
+Machine learning can build a prediction model from a large number of predictors such as radiolomic features @dercleIdentificationNonSmall2020, pathology image or gene expression data @wayMachineLearningDetects2018.
+Because most direct genomic test are more specific and sensitive than predictive models, machine learning approach might has limited role in clinical practice.
 However the machine learning prediction can be a second best when the direct test fails.
+
+Prediction RAS pathway activation with gene expression data was done in previous study.
+@way_2018_machine_cellreports They trained pancancer The Cancer Genome Atlas (TCGA) data with a supervised elastic net penalized logistic regression classifier with stochastic gradient descent.
+The performance of their model was 84% with an area under the receiver operating characteristic (AUROC) curve and 63% with an area under the precision recall (AUPR) curve.
+The authors suggested their approach can be applied to other genomic alterations.
 
@@
 
-Prediction RAS pathway activation with gene expression data was done in previous study.
-@way_2018_machine_cellreports They trained pancancer The Cancer Genome Atlas (TCGA) data with a supervised elastic net penalized logistic regression classifier with stochastic gradient descent.
-The performance of their model was 84% with an area under the receiver operating characteristic (AUROC) curve and 63% with an area under the precision recall (AUPR) curve.
-We applied their modeling methods in prediction PIK3CA mutation.
+We apply a supervised elastic net penalized logistic regression model in prediction PIK3CA mutation.
+The purpose of this study is to know this prediction model approach can be applied not only RAS pathway activation but also PIK3CA mutation across many cancer types.
 
@@
 10845 cases were available both PIK3CA mutation and mRNA expression data.
-5128 out of 20502 genes were included in the modeling process after filtering with median absolute deviation as described at modelinf process method.
+5128 out of 20502 genes were included in the modeling process after filtering with median absolute deviation as described at modeling process method.
 33 cancer type dummy variables were included in predictor variables.
@@
 The status of  PIK3CA was considered as positive when the case has following PIK3CA variants which is the target variables of the therascreen PIK3CA RGQ PCR Kit; C420R, E542K, E545A, E545D, E545G, E545K, Q546E, Q546R, H1047L, H1047R, H1047Y.
-The therascreen PIK3CA RGQ PCR Kit was approved by U.S.
-Food and Drug adminidtration for PIK3CA inhibitor.
+The therascreen PIK3CA RGQ PCR Kit was approved as a companion diagnosis to treat with PIK3CA inhibitor by U.S.
+Food and Drug adminidtration.
 We splited the three quarters of dataset for the trainset and one quarter for testset.
@@
 
-## Dataset summary 
-Prevalcence of PIK3CA was 0.11 in all cases. The PIK3CA prevalance of each cancer type was vary. The median prevalence of PIK3CA of each cancer types  was 0.03 (range 0-0.33) (Figure 1).
+## prevalance rate rate of PIK3CA mutation   
+prevalance rate rate of PIK3CA was 0.11 in all cases. The PIK3CA prevalance rate of each cancer type was vary. The median prevalence of PIK3CA of each cancer types  was 0.03 (range 0-0.33) (Figure 1).
 
@@
 
-In 10-fold cross-validation, the model with lambda = 0.01 and alpha = 1.0 (Ridge regression) were best performance in terms of AUROC.
+In 10-fold cross-validation, the model with $\lambda$ = 0.01 and $\alpha$ = 1.0 (Ridge regression) showed best performance in terms of AUROC.
 The final model was trained with the selected hyperparameters with all trainset.
@@
 
+- Main message answers the question and main supporting evidence
+
 Our model showed good performance to predict PIK3CA mutation of various cancer types.
-
-- Main message answers the question and main supporting evidence
+This result shows that the supervised elastic net penalized logistic regression model can be applied not only RAS activation pathway but also PIK3CA mutation.
+Both RAS activation pathway and PIK3CA mutation are important and common cancer genomic alterations.
+They have significant effect on gene expression in cacer cells.
+It might be challenging prediction of genomic alterations which are infrequent or have weak effect on gene expression.
 
@@
 
-OUr model includes cancer type predictor and they are stronger than gene expression data.
+Our model includes cancer type predictor and they are stronger than gene expression data.
 The varing prevalence of PIC3CA mutation across cancer type might reason of the strong cancer type predictor.
@@
 
-- Figure 1. Summary of modeling results
+- Figure 1. prevalance rate rate of PIK3CA across cancer types
+The abbreviations of cancer types are explained in S1 appendix.
+
+- Figure 2. Summary of modeling results
 
@@
 
-- Figure 2. Coefficients of model
+- Figure 3. Coefficients of model
 
-(A) Top 30 high coefficients of mRNA. (B) Coefficients of cancer types. The abbreviations are explained in S1 appendix.
+(A) Top 30 high coefficients of mRNA. (B) Coefficients of cancer types. The abbreviations of cancer types are explained in S1 appendix.
 

### 2020-05-07 plosOne/plosOne.tex
edit 5/7

@@
 The machine learning approach has been actively researched recently to
-detect the genomic alterations. Breast cancer with PIK3CA mutation has
-been approved to use PIK3CA inhibitor in hormone receptor positive HER2
-negative subtype. {[}1{]}
-
-\begin{itemize}
-\tightlist
-\item
-  PIK3CA summary
-\end{itemize}
+detect the genomic alterations. The machine learning can build a
+prediction model from a large number of predictors such as radiolomic
+features {[}1{]}, pathology image or gene expression data {[}2{]}. The
+machine learning approach has limitations to apply to clinical practice.
+However the machine learning prediction can be a second best when the
+direct test fails.
 
@@
 (PIP\textsubscript{2}) to make phosphatidylinositol 3,4,5-triphosphate
-(PIP\_\{3\}). Phosphatase and tensin homolog (PTEN) changes PIP\_\{2\}
-to PIP\_\{3\} in contrast PI3K. PIP\_\{3\} is a second mesenger to
-activate protein kinase B (AKT) which is a serine/threonine-specific
-protein kinase. AKT inhibits apoptosis and promote cell proliferation.
+(PIP\textsubscript{3}). Phosphatase and tensin homolog (PTEN) changes
+PIP\textsubscript{2} to PIP\textsubscript{3} in contrast PI3K.
+PIP\textsubscript{3} is a second mesenger to activate protein kinase B
+(AKT) which is a serine/threonine-specific protein kinase. AKT inhibits
+apoptosis and promote cell proliferation.
 
-The PIK3CA mutation is second most driver mutation after TP53. the
+Breast cancer with PIK3CA mutation has been approved to use PIK3CA
+inhibitor in hormone receptor positive HER2 negative subtype. {[}3{]}
+The PIK3CA mutation is second most driver mutation after TP53. The
 PIK3CA mutation is most frequantly founded in endometrial carcinoma
@@
 
-\begin{itemize}
-\tightlist
-\item
-  Predicting mutation by gene expression data
-
-  \begin{itemize}
-  \tightlist
-  \item
-    What for?\\
-  \item
-    Feasible?\\
-  \item
-    Cost?
-  \end{itemize}
-\item
-  Varing frquency of PIK3CA mutation across cancer types
-
-  \begin{itemize}
-  \tightlist
-  \item
-    High in endometrial breast\\
-  \item
-    not common other cancer type
-  \end{itemize}
-\end{itemize}
-
-Universial prediction using gene expression data across cancer type for
-certain kind of mutation. It's clinically useless now, but we want to
-explore the possibilities of the PIK3CA mutation prediction.
-
-\begin{itemize}
-\item
-  RNAseq can be widely used. The mutation status directly prediction
-\item
-  Previous study {[}2{]}
-\item
-  Prediction of aberant activation of a certain pathway vs a specific
-  mutation
-\item
-  The proposed solution
-\end{itemize}
-
 Prediction RAS pathway activation with gene expression data was done in
-previous study. {[}2{]} They trained pancancer The Cancer Genome Atlas
+previous study. {[}4{]} They trained pancancer The Cancer Genome Atlas
 (TCGA) data with a supervised elastic net penalized logistic regression
@@
 We used TCGA pancancer dataset. PIK3CA mutation data was get using cgdsr
-rpackage.{[}3{]} Gene expression data was get from GDAC firehose using
-RTCGAToolbox R package. {[}4{]}
+rpackage.{[}5{]} Gene expression data was get from GDAC firehose using
+RTCGAToolbox R package. {[}6{]}
 
@@
 scaling were done. All preprocessing was done using recipe r package.
-{[}5{]} Penalized logistic regression was applied to prediction
+{[}7{]} Penalized logistic regression was applied to prediction
 modeling. 10-fold cross-validation with targe variable stratification
@@
 (\((1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1\)).
-{[}6{]} Model performance was evealuated with the area under receiver
+{[}8{]} Model performance was evealuated with the area under receiver
 operating characteristic (ROC) and area under the ROC (AUROC) and
 precision recall (PR) and area under curve (AUPR). The modelin process
-was done with tidymodels rpackage. {[}7{]}
+was done with tidymodels rpackage. {[}9{]}
 
@@
 prediction difficult. The AUROC is optimistic in terms of performance.
-The AUPR is more informative than AUROC on imbalaced datasets. {[}8{]}
+The AUPR is more informative than AUROC on imbalaced datasets. {[}10{]}
 
@@
 
-OUr model includes cancer type predictor and they are stronger than gene
+Our model includes cancer type predictor and they are stronger than gene
 expression data. The varing prevalence of PIC3CA mutation across cancer
@@
 \begin{itemize}
-\tightlist
 \item
-  Figure 1. Summary of modeling results
+  Figure 1. Prevalance rate of PIK3CA across cancer types The
+  abbreviations of cancer types are explained in S1 appendix.
+\item
+  Figure 2. Summary of modeling results
 \end{itemize}
@@
 \item
-  Figure 2. Coefficients of model
+  Figure 3. Coefficients of model
 \end{itemize}
@@
   Top 30 high coefficients of mRNA. (B) Coefficients of cancer types.
-  The abbreviations are explained in S1 appendix.
+  The abbreviations of cancer types are explained in S1 appendix.
 \end{enumerate}
@@
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-dercleIdentificationNonSmall2020}{}%
+1. Dercle L, Fronheiser M, Lu L, Du S, Hayes W, Leung DK, et al.
+Identification of NonSmall Cell Lung Cancer Sensitive to Systemic Cancer
+Therapies Using Radiomics. Clin Cancer Res. American Association for
+Cancer Research; 2020;26: 2151--2162.
+doi:\href{https://doi.org/10.1158/1078-0432.CCR-19-2942}{10.1158/1078-0432.CCR-19-2942}
+
+\leavevmode\hypertarget{ref-wayMachineLearningDetects2018}{}%
+2. Way GP, Sanchez-Vega F, La K, Armenia J, Chatila WK, Luna A, et al.
+Machine Learning Detects Pan-cancer Ras Pathway Activation in The Cancer
+Genome Atlas. Cell Reports. Elsevier; 2018;23: 172--180.e3.
+doi:\href{https://doi.org/10.1016/j.celrep.2018.03.046}{10.1016/j.celrep.2018.03.046}
+
 \leavevmode\hypertarget{ref-andre_2019_alpelisib_nengljmeda}{}%
-1. André F, Ciruelos E, Rubovszky G, Campone M, Loibl S, Rugo HS, et al.
+3. André F, Ciruelos E, Rubovszky G, Campone M, Loibl S, Rugo HS, et al.
 Alpelisib for PIK3CA-Mutated, Hormone ReceptorPositive Advanced Breast
@@
 \leavevmode\hypertarget{ref-way_2018_machine_cellreports}{}%
-2. Way GP, Sanchez-Vega F, La K, Armenia J, Chatila WK, Luna A, et al.
+4. Way GP, Sanchez-Vega F, La K, Armenia J, Chatila WK, Luna A, et al.
 Machine Learning Detects Pan-cancer Ras Pathway Activation in The Cancer
@@
 \leavevmode\hypertarget{ref-jacobsenCgdsrRbasedAPI2019}{}%
-3. Jacobsen A, Luna A. Cgdsr: R-based API for accessing the MSKCC cancer
+5. Jacobsen A, Luna A. Cgdsr: R-based API for accessing the MSKCC cancer
 genomics data server (CGDS). 2019.
@@
 \leavevmode\hypertarget{ref-samurRTCGAToolboxNewTool2014}{}%
-4. Samur MK. RTCGAToolbox: A new tool for exporting TCGA Firehose data.
+6. Samur MK. RTCGAToolbox: A new tool for exporting TCGA Firehose data.
 PLoS One. 2014;9(9):e106397.
@@
 \leavevmode\hypertarget{ref-kuhnRecipesPreprocessingTools2020}{}%
-5. Kuhn M, Wickham H. Recipes: Preprocessing tools to create design
+7. Kuhn M, Wickham H. Recipes: Preprocessing tools to create design
 matrices. 2020.
@@
 \leavevmode\hypertarget{ref-friedmanRegularizationPathsGeneralized2010}{}%
-6. Friedman J, Hastie T, Tibshirani R. Regularization Paths for
+8. Friedman J, Hastie T, Tibshirani R. Regularization Paths for
 Generalized Linear Models via Coordinate Descent. J Stat Softw. 2010;33:
@@
 \leavevmode\hypertarget{ref-kuhnTidymodelsEasilyInstall2020}{}%
-7. Kuhn M, Wickham H. Tidymodels: Easily install and load the
+9. Kuhn M, Wickham H. Tidymodels: Easily install and load the
 'tidymodels' packages. 2020.
@@
 \leavevmode\hypertarget{ref-saitoPrecisionRecallPlotMore2015}{}%
-8. Saito T, Rehmsmeier M. The Precision-Recall Plot Is More Informative
+10. Saito T, Rehmsmeier M. The Precision-Recall Plot Is More Informative
 than the ROC Plot When Evaluating Binary Classifiers on Imbalanced

### 2020-05-07 figure.R

@@ -22,8 +22,8 @@ dataset %>%
   coord_flip() +
   theme_Publication() + 
   labs(title = "PIK3CA Prevalance",
-       x = "Prevalance",
-       y = "Cancer type") 
+       y = "Prevalance rate",
+       x = "Cancer type") 
 dev.off()
 ## Prevalence of PIK3CA 
 

### 2020-05-12 plosOne/plosOne.Rmd
gene predictors

@@
 abstract: |
-bibliography: [mybibfile.bib, pik3ca.bib, rpackage.bib] 
+bibliography: [mybibfile.bib, pik3ca.bib, rpackage.bib, genes.bib] 
 indent: true
@@
 Machine learning approach has been actively researched to detect the genomic alterations.
-Machine learning can build a prediction model from a large number of predictors such as radiolomic features @dercleIdentificationNonSmall2020, pathology image or gene expression data @wayMachineLearningDetects2018.
+Machine learning can build a prediction model from a large number of predictors such as radiolomic features @dercleIdentificationNonSmall2020, pathology image [@coudrayClassificationMutationPrediction2018] or gene expression data [@wayMachineLearningDetects2018 @loiPIK3CAMutationsAssociated2010].
 Because most direct genomic test are more specific and sensitive than predictive models, machine learning approach might has limited role in clinical practice.
@@
 AKT inhibits apoptosis and promote cell proliferation.
+@cantleyPhosphoinositide3KinasePathway2002
 
@@
 We used TCGA pancancer dataset.
+TCGA is a cancer genomic consotium that archives data of exom sequencing, gene expression, DNA methylation, protein expression and clicial data of more than 10000 cancer samples across 33 common cancer types.
+The gene expression TCGA pancancer dataset is batch-corrected with normalization.
+TCGA dataset is publically available.
 PIK3CA mutation data was get using cgdsr rpackage.[@jacobsenCgdsrRbasedAPI2019] Gene expression data was get from GDAC firehose using RTCGAToolbox R package.
@@
 Lambda is  penalty scaling parameter and alpha is mixing parameter of penalty function ($(1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1$).
-[@friedmanRegularizationPathsGeneralized2010] Model performance was evealuated with the area under receiver operating characteristic (ROC) and area under the ROC (AUROC) and precision recall (PR) and area under curve (AUPR).
-The modelin process was done with tidymodels rpackage.
+[@friedmanRegularizationPathsGeneralized2010]
+
+## Accessing model performance
+
+Model performance was evealuated with the area under receiver operating characteristic (ROC) and area under the ROC (AUROC) and precision recall (PR) and area under curve (AUPR).
+The prevalence of PIK3CA mutation is low.
+The low prevalence of PIK3CA results in imbalaced dataset which makes the prediction difficult.
+The AUROC is optimistic in terms of performance.
+The AUPR is more informative than AUROC on imbalaced datasets.
+[@saitoPrecisionRecallPlotMore2015]   The modeling process was done with tidymodels rpackage.
 [@kuhnTidymodelsEasilyInstall2020]
@@
 ## prevalance rate rate of PIK3CA mutation   
-prevalance rate rate of PIK3CA was 0.11 in all cases. The PIK3CA prevalance rate of each cancer type was vary. The median prevalence of PIK3CA of each cancer types  was 0.03 (range 0-0.33) (Figure 1).
+Prevalance rate of PIK3CA was 0.11 in all cases. The PIK3CA prevalance rate of each cancer type was vary. The median prevalance rate of PIK3CA of each cancer types  was 0.03 (range 0-0.33) (Figure 1).
 
@@
 
-In this study, the split method was done for testset instead of indepedent testset.
-Because there is no standard analysis method for the RNA-seq gene expression quantification, it is difficult to find gene expression data same as TGCA data.
-Developing standard analysis method for gene expression data is neccessory to apply this prediction model.
-
-In this study, the prevalence of PIK3CA mutation was 11%.
-The low prevalence of PIK3CA results in imbalaced dataset which makes the prediction difficult.
-The AUROC is optimistic in terms of performance.
-The AUPR is more informative than AUROC on imbalaced datasets.
-[@saitoPrecisionRecallPlotMore2015]
+Prediction modeling from TCGA pancancer dataset has limitations regarding to data preprocessing.
+Methods for gancer gene expression has been developed for research.
+To stastical analysis, the gene expression data is processed between-sample normalization to remove batch effect.
+If the model has been trained from the between-sample normalization, a new sample can not be exactly processed like trainset.
+A model based on gene expression TCGA pancacner dataset has limitation on preprocessing.
+It is nessessory developing preprocessing method which is independent with dataset to apply the gene expression data to prediction model.
 
@@
 The specific important mutations can effect stronger downsteam gene expression than the broad events pathway activation.
-The previous study might be more difficult prediction problum.
+The previous study might be more difficult prediction problem.
 
@@
 If the cancer type was wrong or can not be determined, our model performance can be poor.
+
+Some significant gene expression predictors were closely related with PTEN and the PI3K pathway.
+PTEN and IGFR1R are the strongest gene expression predictor which have negative and positive predictive power.
+IGF1R is a tyrosine kinase receptor which activates PI3K.
+[@leroithInsulinlikeGrowthFactor2003] Insulin receptor substrate-2 (IRS2) is the adaptor protein of IGF1R.
+[@heInteractionInsulinReceptor1996] PTEN is an important regulator of PIP~3~ by dephosphorylating  PIP~3~ in constrast PI3K.[@cantleyPhosphoinositide3KinasePathway2002]
+
+Another study of PIK3CA mutation prediction showed good performance AUROC 0.71 in independent testset.
+They made gene-expression signature which is sum of the average of the logarithmic gene expression.
+[@cizkovaGeneExpressionProfiling2010; @loiPIK3CAMutationsAssociated2010]
+
+Another study predicted copy numbear alterations with gene expression using multinomial logistic regression model with least absolute shrinkage and selection operator (LASSO).
+The prediction of 1p/19q codel was very good with an AUROC of 0.997.
+The gene level prediction was good with an AUROC 0.75.
+[@muCNAPEMachineLearning2019]
+
+A Hidden Markov Model Approach for Prediction of Genomic Alterations from Gene Expression Profiling.[@gengHiddenMarkovModel2008]
+
+[@heGeneSignaturesAssociated2020]
 

### 2020-05-12 plosOne/plosOne.tex
gene predictors

@@
 
-Targeted thrapy becomes standard treatment in many cancer patients. To
-treat with the targeted therapy, the The patients should be confirmed
-the indicated genomic alterations in their cancer. Many direct test for
-the genomic alteration has been developed and prooved for their clical
-utility to discriminate which patients will be response the targeted
-therapy.
+Targeted thrapy becomes standard treatment in many cancer patients. Many
+targeted therapy requires test for a specific cancer genomic alteration,
+to treat the patients. Many direct test for the genomic alteration has
+been developed and prooved for their clical utility to discriminate
+which patients will be response the targeted therapy.
 
-The machine learning approach has been actively researched recently to
-detect the genomic alterations. The machine learning can build a
-prediction model from a large number of predictors such as radiolomic
-features {[}1{]}, pathology image or gene expression data {[}2{]}. The
-machine learning approach has limitations to apply to clinical practice.
-However the machine learning prediction can be a second best when the
-direct test fails.
+Machine learning approach has been actively researched to detect the
+genomic alterations. Machine learning can build a prediction model from
+a large number of predictors such as radiolomic features {[}1{]},
+pathology image {[}2{]} or gene expression data {[}3{]}. Because most
+direct genomic test are more specific and sensitive than predictive
+models, machine learning approach might has limited role in clinical
+practice. However the machine learning prediction can be a second best
+when the direct test fails.
+
+Prediction RAS pathway activation with gene expression data was done in
+previous study. {[}5{]} They trained pancancer The Cancer Genome Atlas
+(TCGA) data with a supervised elastic net penalized logistic regression
+classifier with stochastic gradient descent. The performance of their
+model was 84\% with an area under the receiver operating characteristic
+(AUROC) curve and 63\% with an area under the precision recall (AUPR)
+curve. The authors suggested their approach can be applied to other
+genomic alterations.
 
@@
 (AKT) which is a serine/threonine-specific protein kinase. AKT inhibits
-apoptosis and promote cell proliferation.
+apoptosis and promote cell proliferation. {[}6{]}
 
 Breast cancer with PIK3CA mutation has been approved to use PIK3CA
-inhibitor in hormone receptor positive HER2 negative subtype. {[}3{]}
+inhibitor in hormone receptor positive HER2 negative subtype. {[}7{]}
 The PIK3CA mutation is second most driver mutation after TP53. The
@@
 
-Prediction RAS pathway activation with gene expression data was done in
-previous study. {[}4{]} They trained pancancer The Cancer Genome Atlas
-(TCGA) data with a supervised elastic net penalized logistic regression
-classifier with stochastic gradient descent. The performance of their
-model was 84\% with an area under the receiver operating characteristic
-(AUROC) curve and 63\% with an area under the precision recall (AUPR)
-curve. We applied their modeling methods in prediction PIK3CA mutation.
+We apply a supervised elastic net penalized logistic regression model in
+prediction PIK3CA mutation. The purpose of this study is to know this
+prediction model approach can be applied not only RAS pathway activation
+but also PIK3CA mutation across many cancer types.
 
@@
 
-We used TCGA pancancer dataset. PIK3CA mutation data was get using cgdsr
-rpackage.{[}5{]} Gene expression data was get from GDAC firehose using
-RTCGAToolbox R package. {[}6{]}
+We used TCGA pancancer dataset. TCGA is a cancer genomic consotium that
+archives data of exom sequencing, gene expression, DNA methylation,
+protein expression and clicial data of more than 10000 cancer samples
+across 33 common cancer types. The gene expression TCGA pancancer
+dataset is batch-corrected with normalization. TCGA dataset is
+publically available. PIK3CA mutation data was get using cgdsr
+rpackage.{[}8{]} Gene expression data was get from GDAC firehose using
+RTCGAToolbox R package. {[}9{]}
 
@@
 data. 5128 out of 20502 genes were included in the modeling process
-after filtering with median absolute deviation as described at modelinf
+after filtering with median absolute deviation as described at modeling
 process method. 33 cancer type dummy variables were included in
@@
 E542K, E545A, E545D, E545G, E545K, Q546E, Q546R, H1047L, H1047R, H1047Y.
-The therascreen PIK3CA RGQ PCR Kit was approved by U.S. Food and Drug
-adminidtration for PIK3CA inhibitor. We splited the three quarters of
-dataset for the trainset and one quarter for testset.
+The therascreen PIK3CA RGQ PCR Kit was approved as a companion diagnosis
+to treat with PIK3CA inhibitor by U.S. Food and Drug adminidtration. We
+splited the three quarters of dataset for the trainset and one quarter
+for testset.
 
@@
 scaling were done. All preprocessing was done using recipe r package.
-{[}7{]} Penalized logistic regression was applied to prediction
+{[}10{]} Penalized logistic regression was applied to prediction
 modeling. 10-fold cross-validation with targe variable stratification
@@
 (\((1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1\)).
-{[}8{]} Model performance was evealuated with the area under receiver
-operating characteristic (ROC) and area under the ROC (AUROC) and
-precision recall (PR) and area under curve (AUPR). The modelin process
-was done with tidymodels rpackage. {[}9{]}
+{[}11{]}
+
+\hypertarget{accessing-model-performance}{%
+\subsection{Accessing model
+performance}\label{accessing-model-performance}}
+
+Model performance was evealuated with the area under receiver operating
+characteristic (ROC) and area under the ROC (AUROC) and precision recall
+(PR) and area under curve (AUPR). The prevalence of PIK3CA mutation is
+low. The low prevalence of PIK3CA results in imbalaced dataset which
+makes the prediction difficult. The AUROC is optimistic in terms of
+performance. The AUPR is more informative than AUROC on imbalaced
+datasets. {[}12{]} The modeling process was done with tidymodels
+rpackage. {[}13{]}
 
@@
 
-\hypertarget{dataset-summary}{%
-\subsection{Dataset summary}\label{dataset-summary}}
+\hypertarget{prevalance-rate-rate-of-pik3ca-mutation}{%
+\subsection{prevalance rate rate of PIK3CA
+mutation}\label{prevalance-rate-rate-of-pik3ca-mutation}}
 
-Prevalcence of PIK3CA was 0.11 in all cases. The PIK3CA prevalance of
-each cancer type was vary. The median prevalence of PIK3CA of each
-cancer types was 0.03 (range 0-0.33) (Figure 1).
+Prevalance rate of PIK3CA was 0.11 in all cases. The PIK3CA prevalance
+rate of each cancer type was vary. The median prevalance rate of PIK3CA
+of each cancer types was 0.03 (range 0-0.33) (Figure 1).
 
@@
 
-In 10-fold cross-validation, the model with lambda = 0.01 and alpha =
-1.0 (Ridge regression) were best performance in terms of AUROC. The
-final model was trained with the selected hyperparameters with all
-trainset.
+In 10-fold cross-validation, the model with \(\lambda\) = 0.01 and
+\(\alpha\) = 1.0 (Ridge regression) showed best performance in terms of
+AUROC. The final model was trained with the selected hyperparameters
+with all trainset.
 
@@
 
+\begin{itemize}
+\tightlist
+\item
+  Main message answers the question and main supporting evidence
+\end{itemize}
+
 Our model showed good performance to predict PIK3CA mutation of various
-cancer types.
+cancer types. This result shows that the supervised elastic net
+penalized logistic regression model can be applied not only RAS
+activation pathway but also PIK3CA mutation. Both RAS activation pathway
+and PIK3CA mutation are important and common cancer genomic alterations.
+They have significant effect on gene expression in cacer cells. It might
+be challenging prediction of genomic alterations which are infrequent or
+have weak effect on gene expression.
 
 \begin{itemize}
-\item
-  Main message answers the question and main supporting evidence
+\tightlist
 \item
@@
 
-In this study, the split method was done for testset instead of
-indepedent testset. Because there is no standard analysis method for the
-RNA-seq gene expression quantification, it is difficult to find gene
-expression data same as TGCA data. Developing standard analysis method
-for gene expression data is neccessory to apply this prediction model.
-
-In this study, the prevalence of PIK3CA mutation was 11\%. The low
-prevalence of PIK3CA results in imbalaced dataset which makes the
-prediction difficult. The AUROC is optimistic in terms of performance.
-The AUPR is more informative than AUROC on imbalaced datasets. {[}10{]}
+Prediction modeling from TCGA pancancer dataset has limitations
+regarding to data preprocessing. Methods for gancer gene expression has
+been developed for research. To stastical analysis, the gene expression
+data is processed between-sample normalization to remove batch effect.
+If the model has been trained from the between-sample normalization, a
+new sample can not be exactly processed like trainset. A model based on
+gene expression TCGA pancacner dataset has limitation on preprocessing.
+It is nessessory developing preprocessing method which is independent
+with dataset to apply the gene expression data to prediction model.
 
@@
 expression than the broad events pathway activation. The previous study
-might be more difficult prediction problum.
+might be more difficult prediction problem.
 
@@
 poor.
+
+Some significant gene expression predictors were closely related with
+PTEN and the PI3K pathway. PTEN and IGFR1R are the strongest gene
+expression predictor which have negative and positive predictive power.
+IGF1R is a tyrosine kinase receptor which activates PI3K. {[}14{]}
+Insulin receptor substrate-2 (IRS2) is the adaptor protein of IGF1R.
+{[}15{]} PTEN is an important regulator of PIP\textsubscript{3} by
+dephosphorylating PIP\textsubscript{3} in constrast PI3K.{[}6{]}
+
+Another study of PIK3CA mutation prediction showed good performance
+AUROC 0.71 in independent testset. They made gene-expression signature
+which is sum of the average of the logarithmic gene expression.
+{[}4,16{]}
+
+Another study predicted copy numbear alterations with gene expression
+using multinomial logistic regression model with least absolute
+shrinkage and selection operator (LASSO). The prediction of 1p/19q codel
+was very good with an AUROC of 0.997. The gene level prediction was good
+with an AUROC 0.75. {[}17{]}
+
+A Hidden Markov Model Approach for Prediction of Genomic Alterations
+from Gene Expression Profiling.{[}18{]}
+
+{[}19{]}
 
@@
 \item
-  Figure 1. Prevalance rate of PIK3CA across cancer types The
+  Figure 1. prevalance rate rate of PIK3CA across cancer types The
   abbreviations of cancer types are explained in S1 appendix.
@@
 
+\leavevmode\hypertarget{ref-coudrayClassificationMutationPrediction2018}{}%
+2. Coudray N, Ocampo PS, Sakellaropoulos T, Narula N, Snuderl M, Fenyö
+D, et al. Classification and mutation prediction from nonSmall cell lung
+cancer histopathology images using deep learning. Nature Medicine.
+Nature Publishing Group; 2018;24: 1559--1567.
+doi:\href{https://doi.org/10.1038/s41591-018-0177-5}{10.1038/s41591-018-0177-5}
+
 \leavevmode\hypertarget{ref-wayMachineLearningDetects2018}{}%
-2. Way GP, Sanchez-Vega F, La K, Armenia J, Chatila WK, Luna A, et al.
+3. Way GP, Sanchez-Vega F, La K, Armenia J, Chatila WK, Luna A, et al.
 Machine Learning Detects Pan-cancer Ras Pathway Activation in The Cancer
@@
 
+\leavevmode\hypertarget{ref-loiPIK3CAMutationsAssociated2010}{}%
+4. Loi S, Haibe-Kains B, Majjaj S, Lallemand F, Durbecq V, Larsimont D,
+et al. PIK3CA mutations associated with gene signature of low mTORC1
+signaling and better outcomes in estrogen receptorPositive breast
+cancer. PNAS. National Academy of Sciences; 2010;107: 10208--10213.
+doi:\href{https://doi.org/10.1073/pnas.0907011107}{10.1073/pnas.0907011107}
+
+\leavevmode\hypertarget{ref-way_2018_machine_cellreports}{}%
+5. Way GP, Sanchez-Vega F, La K, Armenia J, Chatila WK, Luna A, et al.
+Machine Learning Detects Pan-cancer Ras Pathway Activation in The Cancer
+Genome Atlas. Cell Reports. Elsevier; 2018;23: 172--180.e3.
+doi:\href{https://doi.org/10.1016/j.celrep.2018.03.046}{10.1016/j.celrep.2018.03.046}
+
+\leavevmode\hypertarget{ref-cantleyPhosphoinositide3KinasePathway2002}{}%
+6. Cantley LC. The Phosphoinositide 3-Kinase Pathway. Science. American
+Association for the Advancement of Science; 2002;296: 1655--1657.
+doi:\href{https://doi.org/10.1126/science.296.5573.1655}{10.1126/science.296.5573.1655}
+
 \leavevmode\hypertarget{ref-andre_2019_alpelisib_nengljmeda}{}%
-3. André F, Ciruelos E, Rubovszky G, Campone M, Loibl S, Rugo HS, et al.
+7. André F, Ciruelos E, Rubovszky G, Campone M, Loibl S, Rugo HS, et al.
 Alpelisib for PIK3CA-Mutated, Hormone ReceptorPositive Advanced Breast
@@
 
-\leavevmode\hypertarget{ref-way_2018_machine_cellreports}{}%
-4. Way GP, Sanchez-Vega F, La K, Armenia J, Chatila WK, Luna A, et al.
-Machine Learning Detects Pan-cancer Ras Pathway Activation in The Cancer
-Genome Atlas. Cell Reports. Elsevier; 2018;23: 172--180.e3.
-doi:\href{https://doi.org/10.1016/j.celrep.2018.03.046}{10.1016/j.celrep.2018.03.046}
-
 \leavevmode\hypertarget{ref-jacobsenCgdsrRbasedAPI2019}{}%
-5. Jacobsen A, Luna A. Cgdsr: R-based API for accessing the MSKCC cancer
+8. Jacobsen A, Luna A. Cgdsr: R-based API for accessing the MSKCC cancer
 genomics data server (CGDS). 2019.
@@
 \leavevmode\hypertarget{ref-samurRTCGAToolboxNewTool2014}{}%
-6. Samur MK. RTCGAToolbox: A new tool for exporting TCGA Firehose data.
+9. Samur MK. RTCGAToolbox: A new tool for exporting TCGA Firehose data.
 PLoS One. 2014;9(9):e106397.
@@
 \leavevmode\hypertarget{ref-kuhnRecipesPreprocessingTools2020}{}%
-7. Kuhn M, Wickham H. Recipes: Preprocessing tools to create design
+10. Kuhn M, Wickham H. Recipes: Preprocessing tools to create design
 matrices. 2020.
@@
 \leavevmode\hypertarget{ref-friedmanRegularizationPathsGeneralized2010}{}%
-8. Friedman J, Hastie T, Tibshirani R. Regularization Paths for
+11. Friedman J, Hastie T, Tibshirani R. Regularization Paths for
 Generalized Linear Models via Coordinate Descent. J Stat Softw. 2010;33:
@@
 
-\leavevmode\hypertarget{ref-kuhnTidymodelsEasilyInstall2020}{}%
-9. Kuhn M, Wickham H. Tidymodels: Easily install and load the
-'tidymodels' packages. 2020.
-
 \leavevmode\hypertarget{ref-saitoPrecisionRecallPlotMore2015}{}%
-10. Saito T, Rehmsmeier M. The Precision-Recall Plot Is More Informative
+12. Saito T, Rehmsmeier M. The Precision-Recall Plot Is More Informative
 than the ROC Plot When Evaluating Binary Classifiers on Imbalanced
@@
 
+\leavevmode\hypertarget{ref-kuhnTidymodelsEasilyInstall2020}{}%
+13. Kuhn M, Wickham H. Tidymodels: Easily install and load the
+'tidymodels' packages. 2020.
+
+\leavevmode\hypertarget{ref-leroithInsulinlikeGrowthFactor2003}{}%
+14. LeRoith D, Roberts CT. The insulin-like growth factor system and
+cancer. Cancer Letters. 2003;195: 127--137.
+doi:\href{https://doi.org/10.1016/S0304-3835(03)00159-9}{10.1016/S0304-3835(03)00159-9}
+
+\leavevmode\hypertarget{ref-heInteractionInsulinReceptor1996}{}%
+15. He W, Craparo A, Zhu Y, O'Neill TJ, Wang L-M, Pierce JH, et al.
+Interaction of Insulin Receptor Substrate-2 (IRS-2) with the Insulin and
+Insulin-like Growth Factor I Receptors EVIDENCE FOR TWO DISTINCT
+PHOSPHOTYROSINE-DEPENDENT INTERACTION DOMAINS WITHIN IRS-2. J Biol Chem.
+American Society for Biochemistry and Molecular Biology; 1996;271:
+11641--11645.
+doi:\href{https://doi.org/10.1074/jbc.271.20.11641}{10.1074/jbc.271.20.11641}
+
+\leavevmode\hypertarget{ref-cizkovaGeneExpressionProfiling2010}{}%
+16. Cizkova M, Cizeron-Clairac G, Vacher S, Susini A, Andrieu C,
+Lidereau R, et al. Gene Expression Profiling Reveals New Aspects of
+PIK3CA Mutation in ERalpha-Positive Breast Cancer: Major Implication of
+the Wnt Signaling Pathway. PLOS ONE. Public Library of Science; 5:
+e15647.
+doi:\href{https://doi.org/10.1371/journal.pone.0015647}{10.1371/journal.pone.0015647}
+
+\leavevmode\hypertarget{ref-muCNAPEMachineLearning2019}{}%
+17. Mu Q, Wang J. CNAPE: A Machine Learning Method for Copy Number
+Alteration Prediction from Gene Expression. IEEE/ACM Transactions on
+Computational Biology and Bioinformatics. 2019; 1--1.
+doi:\href{https://doi.org/10.1109/TCBB.2019.2944827}{10.1109/TCBB.2019.2944827}
+
+\leavevmode\hypertarget{ref-gengHiddenMarkovModel2008}{}%
+18. Geng H, Ali HH, Chan WC. A Hidden Markov Model Approach for
+Prediction of Genomic Alterations from Gene Expression Profiling. In:
+Măndoiu I, Sunderraman R, Zelikovsky A, editors. Bioinformatics Research
+and Applications. Berlin, Heidelberg: Springer; 2008. pp. 414--425.
+doi:\href{https://doi.org/10.1007/978-3-540-79450-9_38}{10.1007/978-3-540-79450-9\_38}
+
+\leavevmode\hypertarget{ref-heGeneSignaturesAssociated2020}{}%
+19. He X, Qin C, Zhao Y, Zou L, Zhao H, Cheng C. Gene signatures
+associated with genomic aberrations predict prognosis in neuroblastoma.
+Cancer Communications. 2020;40: 105--118.
+doi:\href{https://doi.org/10.1002/cac2.12016}{10.1002/cac2.12016}
+
 \nolinenumbers

### 2020-05-12 figure.R

@@ -100,15 +100,15 @@ mutationPR <- mutatioPlotPRTest +
 
 geneCoeff <- coeffPlot +
   labs(title = "Coefficiency of genes",
-       x = "Coefficiency",
-       y = "Gene") +
+       y = "Coefficiency",
+       x = "Gene") +
   scale_colour_Publication() +
   theme_Publication()
 
 typeCoeff <- typePlot +
   labs(title = "Coefficiency of cancer type",
-       x = "Coefficiency",
-       y = "Cancer type") +
+       y = "Coefficiency",
+       x = "Cancer type") +
   scale_colour_Publication() +
   theme_Publication()
 

### 2020-05-14 plosOne/plosOne.Rmd
aurp discussion

--- 2020-05-07 plosOne/plosOne.Rmd
@@
 
-Targeted thrapy becomes standard treatment in many cancer patients.
-To treat with the targeted therapy, the The patients should be confirmed the indicated genomic alterations in their cancer.
-Many direct test for the genomic alteration has been developed and prooved for their clical utility to discriminate which patients will be response the targeted therapy.
+Targeted therapy becomes a standard treatment in many cancer patients.
+Much targeted therapy requires a test for a specific cancer genomic alteration, to treat the patients.
+Many direct tests for the genomic alteration have been developed and prooved for their clinical utility to find which patients will be responding to the targeted therapy.
 
-The machine learning approach has been actively researched recently to detect the genomic alterations.
-The machine learning can build a prediction model from a large number of predictors such as radiolomic features @dercleIdentificationNonSmall2020, pathology image or gene expression data @wayMachineLearningDetects2018.
-The machine learning approach has limitations to apply to clinical practice.
-However the machine learning prediction can be a second best when the direct test fails.
+It has been actively researched that the machine learnig approach can be applied to detect genomic alterations.
+Machine learning algorithm can build a prediction model from a large number of predictors such as radiolomic features @dercleIdentificationNonSmall2020, pathology image [@coudrayClassificationMutationPrediction2018] or gene expression data [@wayMachineLearningDetects2018 @loiPIK3CAMutationsAssociated2010].
+Because the most direct genomic test is more specific and sensitive than predictive models, the machine learning approach might have a limited role in clinical practice.
+However, the machine learning prediction can be a second-best when the direct test fails.
 
-PIK3CA is encodes the p110$\alpha$ catalytic subunit of phosphatidylinositol 3′-kinase (PI3K).
-PI3K is a protein kinase which phosphorylates phosphatidylinositol 4,5-biphosphate (PIP~2~) to make phosphatidylinositol 3,4,5-triphosphate (PIP~3~).
-Phosphatase and tensin homolog (PTEN) changes PIP~2~ to PIP~3~ in contrast PI3K.
-PIP~3~ is a second mesenger to activate protein kinase B (AKT) which is a serine/threonine-specific protein kinase.
-AKT inhibits apoptosis and promote cell proliferation.
+Prediction RAS pathway activation with gene expression data was done in the previous study.
+@way_2018_machine_cellreports They trained pan-cancer The Cancer Genome Atlas (TCGA) data with a supervised elastic net penalized logistic regression classifier with stochastic gradient descent.
+The performance of their model was 84% with an area under the receiver operating characteristic (AUROC) curve and 63% with an area under the precision-recall (AUPR) curve.
+The authors suggested their approach can be applied to other genomic alterations.
 
-Breast cancer with PIK3CA mutation has been approved to use PIK3CA inhibitor in hormone receptor positive HER2 negative subtype.
-@andre_2019_alpelisib_nengljmeda The PIK3CA mutation is second most driver mutation after TP53.
-The PIK3CA mutation is most frequantly founded in endometrial carcinoma (45%), and followed by breast invasive carcinoma (24%), cervical squamous cell carcinoma and endocervical adenocarcinoma (20%) and colon adenocarcinoma (16%).
+PIK3CA encodes the p110$\alpha$ catalytic subunit of phosphatidylinositol 3′-kinase (PI3K).
+PI3K is a protein kinase that phosphorylates phosphatidylinositol 4,5-biphosphate (PIP~2~) to make phosphatidylinositol 3,4,5-triphosphate (PIP~3~).
+Phosphatase and tensin homolog (PTEN) changes PIP~2~ to PIP~3~ in contrast, PI3K.
+PIP~3~ is a second messenger to activate protein kinase B (AKT) which is a serine/threonine-specific protein kinase.
+AKT inhibits apoptosis and promotes cell proliferation.
+@cantleyPhosphoinositide3KinasePathway2002
 
-Prediction RAS pathway activation with gene expression data was done in previous study.
-@way_2018_machine_cellreports They trained pancancer The Cancer Genome Atlas (TCGA) data with a supervised elastic net penalized logistic regression classifier with stochastic gradient descent.
-The performance of their model was 84% with an area under the receiver operating characteristic (AUROC) curve and 63% with an area under the precision recall (AUPR) curve.
-We applied their modeling methods in prediction PIK3CA mutation.
+Breast cancer with PIK3CA mutation has been approved to use PIK3CA inhibitor in hormone receptor-positive HER2 negative subtype.
+@andre_2019_alpelisib_nengljmeda The PIK3CA mutation is the second most driver mutation after TP53.
+The PIK3CA mutation is most frequently founded in endometrial carcinoma (45%), and followed by breast invasive carcinoma (24%), cervical squamous cell carcinoma, and endocervical adenocarcinoma (20%) and colon adenocarcinoma (16%).
+
+We apply a supervised elastic net penalized logistic regression model in prediction PIK3CA mutation.
+The purpose of this study is to know this prediction model approach can be applied not only to RAS pathway activation but also PIK3CA mutation across many cancer types.
 
@@
 
-We used TCGA pancancer dataset.
-PIK3CA mutation data was get using cgdsr rpackage.[@jacobsenCgdsrRbasedAPI2019] Gene expression data was get from GDAC firehose using RTCGAToolbox R package.
-[@samurRTCGAToolboxNewTool2014]
+We used the TCGA pan-cancer dataset.
+TCGA is a cancer genomic consortium that archives data of exome sequencing, gene expression, DNA methylation, protein expression, and clinical data of more than 10000 cancer samples across 33 common cancer types.
+The gene expression TCGA pan-cancer dataset is batch-corrected with normalization.
+The TCGA dataset is publically available.
+PIK3CA mutation data was got using cgdsr rpackage.[@jacobsenCgdsrRbasedAPI2019] Gene expression data was downloaded from the National Cancer Institute (NCI)'s Genomic Data Commons (GDC) website that archives data used for The Pan-Cancer Atlas initiative.
+(https://gdc.cancer.gov/about-data/publications/pancanatlas)
 
-10845 cases were available both PIK3CA mutation and mRNA expression data.
-5128 out of 20502 genes were included in the modeling process after filtering with median absolute deviation as described at modelinf process method.
-33 cancer type dummy variables were included in predictor variables.
-
-The target variable was PIK3CA mutation status.
-The status of  PIK3CA was considered as positive when the case has following PIK3CA variants which is the target variables of the therascreen PIK3CA RGQ PCR Kit; C420R, E542K, E545A, E545D, E545G, E545K, Q546E, Q546R, H1047L, H1047R, H1047Y.
-The therascreen PIK3CA RGQ PCR Kit was approved by U.S.
-Food and Drug adminidtration for PIK3CA inhibitor.
-We splited the three quarters of dataset for the trainset and one quarter for testset.
+The target variable was the PIK3CA mutation status.
+The status of  PIK3CA was considered as positive when the case has following PIK3CA variants which are the target variables of the therascreen PIK3CA RGQ PCR Kit; C420R, E542K, E545A, E545D, E545G, E545K, Q546E, Q546R, H1047L, H1047R, H1047Y.
+The therascreen PIK3CA RGQ PCR Kit was approved as a companion diagnostics to treat with PIK3CA inhibitor by the U.S.
+Food and Drug Administration.
+We split the three-quarters of the dataset for the trainset and one quarter for the test set.
 
@@
 
-To narrow down potential predictors, Genes with a large the median absolute deviation (more than third-quartiles) were selected.
+To narrow down potential predictors, genes with a large the median absolute deviation (more than third-quartiles) were selected.
+33 cancer type dummy variables were included in predictor variables.
 Yeo-Johnson transformation was done to correct skewness.
@@
 Lambda is  penalty scaling parameter and alpha is mixing parameter of penalty function ($(1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1$).
-[@friedmanRegularizationPathsGeneralized2010] Model performance was evealuated with the area under receiver operating characteristic (ROC) and area under the ROC (AUROC) and precision recall (PR) and area under curve (AUPR).
-The modelin process was done with tidymodels rpackage.
+[@friedmanRegularizationPathsGeneralized2010]
+
+## Accessing model performance
+
+Model performance was evaluated with the area under the receiver operating characteristic (ROC) and area under the ROC (AUROC) and precision-recall (PR) and area under the curve (AUPR).
+The AUPR is more informative than AUROC on imbalanced datasets.
+[@saitoPrecisionRecallPlotMore2015] The modeling process and accessing model performance were done with tidymodels rpackage.
 [@kuhnTidymodelsEasilyInstall2020]
@@
 ## Dataset summary
-Prevalcence of PIK3CA was 0.11 in all cases. The PIK3CA prevalance of each cancer type was vary. The median prevalence of PIK3CA of each cancer types  was 0.03 (range 0-0.33) (Figure 1).
+
+10845 cases were available for both PIK3CA mutation and mRNA expression data.
+5128 out of 20502 genes were included in the modeling process after filtering with median absolute deviation as described at the modeling process method.
+The prevalence rate of PIK3CA was 0.11 in all cases.
+The PIK3CA prevalence rate of each cancer type varied.
+The median prevalence rate of PIK3CA of each cancer type was 0.03 (range 0-0.33) (Figure 1).
 
@@
 
-In 10-fold cross-validation, the model with lambda = 0.01 and alpha = 1.0 (Ridge regression) were best performance in terms of AUROC.
-The final model was trained with the selected hyperparameters with all trainset.
-
-The trainset AUROC was 0.93 and the testset AUROC was 0.84. The AUPR of trainset was 0.66 and the testset AUPR was 0.39. (Figure 1A)
+In 10-fold cross-validation, the model with $\lambda$ = 0.01 and $\alpha$ = 1.0 (Ridge regression) showed best performance in terms of AUROC.
+The final model was trained with the selected hyperparameters with the entire trainset.
+The trainset AUROC was 0.93 and the testset AUROC was 0.84.
+The AUPR of trainset was 0.66 and the testset AUPR was 0.39.
+(Figure 2A)
 
@@
 
-Because tht prevalence of PIK3CA mutation is vary across the cancer type, the performance of each cancer type was investigated.
-The AUROC and AUPR were positively correlated between train set and test set in cancer type subanalysis.
-(Figure 1B) The AUPR was high in cancer type with high PIK3CA mutation rate such as colon, brest, Uterus cancer types.
-The AUROC did not correlated with PIK3CA mutation rate of each cancer types.
-(Figure 1C)
+Because the prevalence of PIK3CA mutation varies across the cancer type, the performance of each cancer type was investigated.
+The AUROC and AUPR were positively correlated between the train set and test set in cancer type subanalysis.
+(Figure 2B) The AUPR was high in cancer type with high PIK3CA mutation rate such as colon, breast, uterus cancer types.
+The AUROC did not correlate with the PIK3CA mutation rate of each cancer type.
+(Figure 2C)
 
@@
 
-Figure 2 shows top 30 important predictors.
-The coefficient is the parameter of the predictor which represent the effect of the predictor on prediction.
+Figure 3 shows the top 30 important predictors.
+The coefficient is the parameter of the predictor which represents the effect of the predictor on prediction.
 IGF1R mRNA expression was the strongest negative predictor and PTEN was the strongest positive predictor.
-Both genes and PIK3CA are key players in tyrosin kanase pathway.
-The cancer type was important predictors.
-Some cancer types including uterine carcinosarcoma (UCS), bladder urothelial carcinoma (BLCA), pancreatic adenocarcinoma (PAAD), lymphoid neoplasm diffuse large B-cell lymphoma (DLBC) are strongest predictors.
+Both IGFR1R and PTEN are key players in the tyrosine kinase pathway.
+The cancer type was an important predictor.
+Some cancer types including uterine carcinosarcoma (UCS), bladder urothelial carcinoma (BLCA), pancreatic adenocarcinoma (PAAD), lymphoid neoplasm diffuse large B-cell lymphoma (DLBC) are the strongest predictors.
 
@@
 
-Our model showed good performance to predict PIK3CA mutation of various cancer types.
+- The main message answers the question and main supporting evidence
 
-- Main message answers the question and main supporting evidence
+Our model showed good performance to predict the PIK3CA mutation of various cancer types.
+This result suggests that the supervised elastic net penalized logistic regression model can be applied not only to the RAS activation pathway but also other genomic alterations.
+Both the RAS activation pathway and PIK3CA mutation are important and common cancer genomic alterations.
+They have a significant effect on gene expression in cancer cells.
+It can not be generalized that the supervised elastic net penalized logistic regression model can be applied to other genomic alteration.
+It might be a challenging prediction of genomic alterations that are infrequent or have a weak effect on gene expression.
 
@@
 
-In this study, the split method was done for testset instead of indepedent testset.
-Because there is no standard analysis method for the RNA-seq gene expression quantification, it is difficult to find gene expression data same as TGCA data.
-Developing standard analysis method for gene expression data is neccessory to apply this prediction model.
-
-In this study, the prevalence of PIK3CA mutation was 11%.
-The low prevalence of PIK3CA results in imbalaced dataset which makes the prediction difficult.
-The AUROC is optimistic in terms of performance.
-The AUPR is more informative than AUROC on imbalaced datasets.
-[@saitoPrecisionRecallPlotMore2015]
+Prediction modeling from the TCGA pan-cancer dataset has limitations regarding data preprocessing.
+Methods for cancer gene expression have been developed for research.
+The gene expression data is processed between-sample normalization to remove the batch effect.
+If the model has been trained from the between-sample normalization, a new sample can not be exactly preprocessed with normalization which was done on trainset.
+A model based on gene expression TCGA pan-cancer dataset has a limitation on data preprocessing.
+It is necessary for developing a preprocessing method that is independent with a dataset to apply the gene expression data to the prediction model.
 
@@
 
-Our PIK3CA prediction model performed better than RAS activation prediction model of previous study in terms of both AUROC (0.84 vs 0.75) and AUPR (0.39 vs 0.24) on testset prediction.
-Our testset is corresponding to the samples initially filtered from training.
-The target variable of our study is more specific than the previous study.
-The specific important mutations can effect stronger downsteam gene expression than the broad events pathway activation.
-The previous study might be more difficult prediction problum.
+Our PIK3CA prediction model performed similar with the RAS activation prediction model of the previous study in terms of AUROC (0.84).
+However AUPR of our model for PIK3CA is lower than model for RAS activation (0.39 vs 0.63).
+The reason of lower AUPR of our model can be explained by low prevalence rate of PIK3CA mutation and imbalanced dataset.
+The model for RAS activation trained with cancer types with more than 0.05 prevalence of RAS activation to avoid imbalance classification problem.
+We included all cancer types into modeling process.
+The lower prevalance rate of target variable mean our dataset has a lower baseline of AUPR.
+In the subanalysis for performance of each cancer types, the cancer types with higher PIK3CA mutation rate showe better AUPR.
+The imbalanced data of our model
 
-OUr model includes cancer type predictor and they are stronger than gene expression data.
-The varing prevalence of PIC3CA mutation across cancer type might reason of the strong cancer type predictor.
+Our model includes cancer type predictor and they are stronger than gene expression data.
+The varying prevalence of PIC3CA mutation across cancer type might reason for the strong cancer type predictor.
 If the cancer type was wrong or can not be determined, our model performance can be poor.
+
+Some significant gene expression predictors were closely related to PTEN and the PI3K pathway.
+PTEN and IGFR1R are the strongest gene expression predictor which have negative and positive predictive power.
+IGF1R is a tyrosine kinase receptor that activates PI3K.
+[@leroithInsulinlikeGrowthFactor2003] Insulin receptor substrate-2 (IRS2) is the adaptor protein of IGF1R.
+[@heInteractionInsulinReceptor1996] PTEN is an important regulator of PIP~3~ by dephosphorylating  PIP~3~ in contrast PI3K.[@cantleyPhosphoinositide3KinasePathway2002]
+
+There is some study trying to predict genomic alterations from gene expression data.
+A study performed PIK3CA mutation prediction by a gene-expression signature which is a sum of the average of the logarithmic gene expression.
+The model showed good performance AUROC 0.71 in an independent test set.
+[@cizkovaGeneExpressionProfiling2010; @loiPIK3CAMutationsAssociated2010] Another study predicted copy number alterations with gene expression using a multinomial logistic regression model with least absolute shrinkage and selection operator (LASSO).
+The prediction of the 1p/19q codel was very good with an AUROC of 0.997.
+The gene-level prediction was good with an AUROC 0.75.
+[@muCNAPEMachineLearning2019] A Hidden Markov Model Approach for Prediction of Genomic Alterations from Gene Expression Profiling.[@gengHiddenMarkovModel2008] A logistic regression model was used for MYCN gene amplification in neuroblastoma.
+[@heGeneSignaturesAssociated2020]
 
@@
 
-- Proof of concept study  
-- Reproducibility of gene expression prediction model
+Our study suggests that the prediction of genomic alteration with gene expression data is possible with good performance.
+However, improvement of performance is required for applying to clinical test and the standardization of generation processing of gene expression data is also needed.
 
@@
 
-- Figure 1. Summary of modeling results
+- Figure 1. prevalence rate of PIK3CA across cancer types
+The abbreviations of cancer types are explained in the S1 Appendix.
 
-(A) Left: receiver operating characteristic (ROC) curve right: precision recall (PR) curve of trainset and testset.
-The horizontal green line is the PIK3CA mutation rate (0.11)  (B) Correlation between trainset and testset of area under receiver operating characteristic curve (AUROC) and area under precision recall curve (AUPR) among cancer type.
-The abbreviations are explained in S1 appendix.
-(C) Correlation between the PIK3CA mutation rate of area under receiver operating characteristic curve (AUROC) and area under precision recall curve (AUPR).
+- Figure 2. Summary of modeling results
 
-- Figure 2. Coefficients of model
+(A) Left: receiver operating characteristic (ROC) curve right: precision-recall (PR) curve of trainset and test set.
+The horizontal green line is the PIK3CA mutation rate (0.11)  (B) Correlation between trainset and test set of the area under the receiver operating characteristic curve (AUROC) and area under the precision-recall curve (AUPR) among cancer type.
+The gray band is 95% confidence interval.
+The abbreviations are explained in the S1 Appendix.
+(C) Correlation between the PIK3CA mutation rate of area under the receiver operating characteristic curve (AUROC) and area under the precision-recall curve (AUPR).
 
-(A) Top 30 high coefficients of mRNA. (B) Coefficients of cancer types. The abbreviations are explained in S1 appendix.
+- Figure 3. Coefficients of model
+
+(A) Top 30 high coefficients of mRNA. (B) Coefficients of cancer types. The abbreviations of cancer types are explained in the S1 Appendix.
 

### 2020-05-14 plosOne/plosOne.tex
aurp discussion

@@
 
-Targeted thrapy becomes standard treatment in many cancer patients. Many
-targeted therapy requires test for a specific cancer genomic alteration,
-to treat the patients. Many direct test for the genomic alteration has
-been developed and prooved for their clical utility to discriminate
-which patients will be response the targeted therapy.
+Targeted therapy becomes a standard treatment in many cancer patients.
+Much targeted therapy requires a test for a specific cancer genomic
+alteration, to treat the patients. Many direct tests for the genomic
+alteration have been developed and prooved for their clinical utility to
+find which patients will be responding to the targeted therapy.
 
-Machine learning approach has been actively researched to detect the
+The machine learning approach has been actively researched to detect
 genomic alterations. Machine learning can build a prediction model from
 a large number of predictors such as radiolomic features {[}1{]},
-pathology image {[}2{]} or gene expression data {[}3{]}. Because most
-direct genomic test are more specific and sensitive than predictive
-models, machine learning approach might has limited role in clinical
-practice. However the machine learning prediction can be a second best
-when the direct test fails.
+pathology image {[}2{]} or gene expression data {[}3{]}. Because the
+most direct genomic test is more specific and sensitive than predictive
+models, the machine learning approach might have a limited role in
+clinical practice. However, the machine learning prediction can be a
+second-best when the direct test fails.
 
 Prediction RAS pathway activation with gene expression data was done in
-previous study. {[}5{]} They trained pancancer The Cancer Genome Atlas
-(TCGA) data with a supervised elastic net penalized logistic regression
-classifier with stochastic gradient descent. The performance of their
-model was 84\% with an area under the receiver operating characteristic
-(AUROC) curve and 63\% with an area under the precision recall (AUPR)
-curve. The authors suggested their approach can be applied to other
-genomic alterations.
+the previous study. {[}5{]} They trained pan-cancer The Cancer Genome
+Atlas (TCGA) data with a supervised elastic net penalized logistic
+regression classifier with stochastic gradient descent. The performance
+of their model was 84\% with an area under the receiver operating
+characteristic (AUROC) curve and 63\% with an area under the
+precision-recall (AUPR) curve. The authors suggested their approach can
+be applied to other genomic alterations.
 
-PIK3CA is encodes the p110\(\alpha\) catalytic subunit of
-phosphatidylinositol 3′-kinase (PI3K). PI3K is a protein kinase which
+PIK3CA encodes the p110\(\alpha\) catalytic subunit of
+phosphatidylinositol 3′-kinase (PI3K). PI3K is a protein kinase that
 phosphorylates phosphatidylinositol 4,5-biphosphate
@@
 (PIP\textsubscript{3}). Phosphatase and tensin homolog (PTEN) changes
-PIP\textsubscript{2} to PIP\textsubscript{3} in contrast PI3K.
-PIP\textsubscript{3} is a second mesenger to activate protein kinase B
+PIP\textsubscript{2} to PIP\textsubscript{3} in contrast, PI3K.
+PIP\textsubscript{3} is a second messenger to activate protein kinase B
 (AKT) which is a serine/threonine-specific protein kinase. AKT inhibits
-apoptosis and promote cell proliferation. {[}6{]}
+apoptosis and promotes cell proliferation. {[}6{]}
 
 Breast cancer with PIK3CA mutation has been approved to use PIK3CA
-inhibitor in hormone receptor positive HER2 negative subtype. {[}7{]}
-The PIK3CA mutation is second most driver mutation after TP53. The
-PIK3CA mutation is most frequantly founded in endometrial carcinoma
+inhibitor in hormone receptor-positive HER2 negative subtype. {[}7{]}
+The PIK3CA mutation is the second most driver mutation after TP53. The
+PIK3CA mutation is most frequently founded in endometrial carcinoma
 (45\%), and followed by breast invasive carcinoma (24\%), cervical
-squamous cell carcinoma and endocervical adenocarcinoma (20\%) and colon
-adenocarcinoma (16\%).
+squamous cell carcinoma, and endocervical adenocarcinoma (20\%) and
+colon adenocarcinoma (16\%).
 
@@
 prediction PIK3CA mutation. The purpose of this study is to know this
-prediction model approach can be applied not only RAS pathway activation
-but also PIK3CA mutation across many cancer types.
+prediction model approach can be applied not only to RAS pathway
+activation but also PIK3CA mutation across many cancer types.
 
@@
 
-We used TCGA pancancer dataset. TCGA is a cancer genomic consotium that
-archives data of exom sequencing, gene expression, DNA methylation,
-protein expression and clicial data of more than 10000 cancer samples
-across 33 common cancer types. The gene expression TCGA pancancer
-dataset is batch-corrected with normalization. TCGA dataset is
-publically available. PIK3CA mutation data was get using cgdsr
-rpackage.{[}8{]} Gene expression data was get from GDAC firehose using
-RTCGAToolbox R package. {[}9{]}
+We used the TCGA pan-cancer dataset. TCGA is a cancer genomic consortium
+that archives data of exome sequencing, gene expression, DNA
+methylation, protein expression, and clinical data of more than 10000
+cancer samples across 33 common cancer types. The gene expression TCGA
+pan-cancer dataset is batch-corrected with normalization. The TCGA
+dataset is publically available. PIK3CA mutation data was got using
+cgdsr rpackage.{[}8{]} Gene expression data was got from the GDAC
+firehose using the RTCGAToolbox R package. {[}9{]}
 
-10845 cases were available both PIK3CA mutation and mRNA expression
+10845 cases were available for both PIK3CA mutation and mRNA expression
 data. 5128 out of 20502 genes were included in the modeling process
-after filtering with median absolute deviation as described at modeling
-process method. 33 cancer type dummy variables were included in
+after filtering with median absolute deviation as described at the
+modeling process method. 33 cancer type dummy variables were included in
 predictor variables.
 
-The target variable was PIK3CA mutation status. The status of PIK3CA was
-considered as positive when the case has following PIK3CA variants which
-is the target variables of the therascreen PIK3CA RGQ PCR Kit; C420R,
-E542K, E545A, E545D, E545G, E545K, Q546E, Q546R, H1047L, H1047R, H1047Y.
-The therascreen PIK3CA RGQ PCR Kit was approved as a companion diagnosis
-to treat with PIK3CA inhibitor by U.S. Food and Drug adminidtration. We
-splited the three quarters of dataset for the trainset and one quarter
-for testset.
+The target variable was the PIK3CA mutation status. The status of PIK3CA
+was considered as positive when the case has following PIK3CA variants
+which are the target variables of the therascreen PIK3CA RGQ PCR Kit;
+C420R, E542K, E545A, E545D, E545G, E545K, Q546E, Q546R, H1047L, H1047R,
+H1047Y. The therascreen PIK3CA RGQ PCR Kit was approved as a companion
+diagnostics to treat with PIK3CA inhibitor by the U.S. Food and Drug
+Administration. We split the three-quarters of the dataset for the
+trainset and one quarter for the test set.
 
@@
 
-Model performance was evealuated with the area under receiver operating
-characteristic (ROC) and area under the ROC (AUROC) and precision recall
-(PR) and area under curve (AUPR). The prevalence of PIK3CA mutation is
-low. The low prevalence of PIK3CA results in imbalaced dataset which
-makes the prediction difficult. The AUROC is optimistic in terms of
-performance. The AUPR is more informative than AUROC on imbalaced
-datasets. {[}12{]} The modeling process was done with tidymodels
-rpackage. {[}13{]}
+Model performance was evaluated with the area under the receiver
+operating characteristic (ROC) and area under the ROC (AUROC) and
+precision-recall (PR) and area under the curve (AUPR). The prevalence of
+PIK3CA mutation is low. The low prevalence of PIK3CA results in an
+imbalanced dataset which makes the prediction difficult. The AUROC is
+optimistic in terms of performance. The AUPR is more informative than
+AUROC on imbalanced datasets. {[}12{]} The modeling process was done
+with tidymodels rpackage. {[}13{]}
 
@@
 
-\hypertarget{prevalance-rate-rate-of-pik3ca-mutation}{%
-\subsection{prevalance rate rate of PIK3CA
-mutation}\label{prevalance-rate-rate-of-pik3ca-mutation}}
+\hypertarget{prevalence-rate-of-pik3ca-mutation}{%
+\subsection{prevalence rate of PIK3CA
+mutation}\label{prevalence-rate-of-pik3ca-mutation}}
 
-Prevalance rate of PIK3CA was 0.11 in all cases. The PIK3CA prevalance
-rate of each cancer type was vary. The median prevalance rate of PIK3CA
-of each cancer types was 0.03 (range 0-0.33) (Figure 1).
+The prevalence rate of PIK3CA was 0.11 in all cases. The PIK3CA
+prevalence rate of each cancer type varied. The median prevalence rate
+of PIK3CA of each cancer type was 0.03 (range 0-0.33) (Figure 1).
 
@@
 AUROC. The final model was trained with the selected hyperparameters
-with all trainset.
+with the entire trainset.
 
@@
 
-Because tht prevalence of PIK3CA mutation is vary across the cancer
-type, the performance of each cancer type was investigated. The AUROC
-and AUPR were positively correlated between train set and test set in
-cancer type subanalysis. (Figure 1B) The AUPR was high in cancer type
-with high PIK3CA mutation rate such as colon, brest, Uterus cancer
-types. The AUROC did not correlated with PIK3CA mutation rate of each
-cancer types. (Figure 1C)
+Because the prevalence of PIK3CA mutation varies across the cancer type,
+the performance of each cancer type was investigated. The AUROC and AUPR
+were positively correlated between the train set and test set in cancer
+type subanalysis. (Figure 1B) The AUPR was high in cancer type with high
+PIK3CA mutation rate such as colon, breast, Uterus cancer types. The
+AUROC did not correlate with the PIK3CA mutation rate of each cancer
+type. (Figure 1C)
 
@@
 
-Figure 2 shows top 30 important predictors. The coefficient is the
-parameter of the predictor which represent the effect of the predictor
+Figure 2 shows the top 30 important predictors. The coefficient is the
+parameter of the predictor which represents the effect of the predictor
 on prediction. IGF1R mRNA expression was the strongest negative
 predictor and PTEN was the strongest positive predictor. Both genes and
-PIK3CA are key players in tyrosin kanase pathway. The cancer type was
-important predictors. Some cancer types including uterine carcinosarcoma
-(UCS), bladder urothelial carcinoma (BLCA), pancreatic adenocarcinoma
-(PAAD), lymphoid neoplasm diffuse large B-cell lymphoma (DLBC) are
-strongest predictors.
+PIK3CA are key players in the tyrosine kinase pathway. The cancer type
+was an important predictor. Some cancer types including uterine
+carcinosarcoma (UCS), bladder urothelial carcinoma (BLCA), pancreatic
+adenocarcinoma (PAAD), lymphoid neoplasm diffuse large B-cell lymphoma
+(DLBC) are the strongest predictors.
 
@@
 \item
-  Main message answers the question and main supporting evidence
+  The main message answers the question and main supporting evidence
 \end{itemize}
 
-Our model showed good performance to predict PIK3CA mutation of various
-cancer types. This result shows that the supervised elastic net
-penalized logistic regression model can be applied not only RAS
-activation pathway but also PIK3CA mutation. Both RAS activation pathway
-and PIK3CA mutation are important and common cancer genomic alterations.
-They have significant effect on gene expression in cacer cells. It might
-be challenging prediction of genomic alterations which are infrequent or
-have weak effect on gene expression.
+Our model showed good performance to predict the PIK3CA mutation of
+various cancer types. This result shows that the supervised elastic net
+penalized logistic regression model can be applied not only to the RAS
+activation pathway but also PIK3CA mutation. Both the RAS activation
+pathway and PIK3CA mutation are important and common cancer genomic
+alterations. They have a significant effect on gene expression in cancer
+cells. It might be a challenging prediction of genomic alterations that
+are infrequent or have a weak effect on gene expression.
 
@@
 
-Prediction modeling from TCGA pancancer dataset has limitations
-regarding to data preprocessing. Methods for gancer gene expression has
-been developed for research. To stastical analysis, the gene expression
-data is processed between-sample normalization to remove batch effect.
-If the model has been trained from the between-sample normalization, a
-new sample can not be exactly processed like trainset. A model based on
-gene expression TCGA pancacner dataset has limitation on preprocessing.
-It is nessessory developing preprocessing method which is independent
-with dataset to apply the gene expression data to prediction model.
+Prediction modeling from the TCGA pan-cancer dataset has limitations
+regarding data preprocessing. Methods for cancer gene expression have
+been developed for research. To statistical analysis, the gene
+expression data is processed between-sample normalization to remove the
+batch effect. If the model has been trained from the between-sample
+normalization, a new sample can not be exactly processed like trainset.
+A model based on gene expression TCGA pan-cancer dataset has a
+limitation on preprocessing. It is necessary for developing a
+preprocessing method that is independent with a dataset to apply the
+gene expression data to the prediction model.
 
@@
 
-Our PIK3CA prediction model performed better than RAS activation
-prediction model of previous study in terms of both AUROC (0.84 vs 0.75)
-and AUPR (0.39 vs 0.24) on testset prediction. Our testset is
+Our PIK3CA prediction model performed better than the RAS activation
+prediction model of the previous study in terms of both AUROC (0.84 vs
+0.75) and AUPR (0.39 vs 0.24) on test set prediction. Our test set is
 corresponding to the samples initially filtered from training. The
 target variable of our study is more specific than the previous study.
-The specific important mutations can effect stronger downsteam gene
+The specific important mutations can affect stronger downstream gene
 expression than the broad events pathway activation. The previous study
-might be more difficult prediction problem.
+might be a more difficult prediction problem.
 
 Our model includes cancer type predictor and they are stronger than gene
-expression data. The varing prevalence of PIC3CA mutation across cancer
-type might reason of the strong cancer type predictor. If the cancer
+expression data. The varying prevalence of PIC3CA mutation across cancer
+type might reason for the strong cancer type predictor. If the cancer
 type was wrong or can not be determined, our model performance can be
@@
 
-Some significant gene expression predictors were closely related with
-PTEN and the PI3K pathway. PTEN and IGFR1R are the strongest gene
-expression predictor which have negative and positive predictive power.
-IGF1R is a tyrosine kinase receptor which activates PI3K. {[}14{]}
-Insulin receptor substrate-2 (IRS2) is the adaptor protein of IGF1R.
-{[}15{]} PTEN is an important regulator of PIP\textsubscript{3} by
-dephosphorylating PIP\textsubscript{3} in constrast PI3K.{[}6{]}
+Some significant gene expression predictors were closely related to PTEN
+and the PI3K pathway. PTEN and IGFR1R are the strongest gene expression
+predictor which have negative and positive predictive power. IGF1R is a
+tyrosine kinase receptor that activates PI3K. {[}14{]} Insulin receptor
+substrate-2 (IRS2) is the adaptor protein of IGF1R. {[}15{]} PTEN is an
+important regulator of PIP\textsubscript{3} by dephosphorylating
+PIP\textsubscript{3} in contrast PI3K.{[}6{]}
 
-Another study of PIK3CA mutation prediction showed good performance
-AUROC 0.71 in independent testset. They made gene-expression signature
-which is sum of the average of the logarithmic gene expression.
-{[}4,16{]}
-
-Another study predicted copy numbear alterations with gene expression
-using multinomial logistic regression model with least absolute
-shrinkage and selection operator (LASSO). The prediction of 1p/19q codel
-was very good with an AUROC of 0.997. The gene level prediction was good
-with an AUROC 0.75. {[}17{]}
-
-A Hidden Markov Model Approach for Prediction of Genomic Alterations
-from Gene Expression Profiling.{[}18{]}
-
+There is some study trying to predict genomic alterations from gene
+expression data. A study performed PIK3CA mutation prediction by a
+gene-expression signature which is a sum of the average of the
+logarithmic gene expression. The model showed good performance AUROC
+0.71 in an independent test set. {[}4,16{]} Another study predicted copy
+number alterations with gene expression using a multinomial logistic
+regression model with least absolute shrinkage and selection operator
+(LASSO). The prediction of the 1p/19q codel was very good with an AUROC
+of 0.997. The gene-level prediction was good with an AUROC 0.75.
+{[}17{]} A Hidden Markov Model Approach for Prediction of Genomic
+Alterations from Gene Expression Profiling.{[}18{]} A logistic
+regression model was used for MYCN gene amplification in neuroblastoma.
 {[}19{]}
@@
 \begin{itemize}
+\tightlist
 \item
@@
   and suggestions for further research.
-\item
-  Proof of concept study\\
-\item
-  Reproducibility of gene expression prediction model
 \end{itemize}
+
+Our study suggests that the prediction of genomic alteration with gene
+expression data is possible with good performance. However, improvement
+of performance is required for applying to clinical test and the
+standardization of generation processing of gene expression data is also
+needed.
 
@@
 \item
-  Figure 1. prevalance rate rate of PIK3CA across cancer types The
-  abbreviations of cancer types are explained in S1 appendix.
+  Figure 1. prevalence rate of PIK3CA across cancer types The
+  abbreviations of cancer types are explained in the S1 Appendix.
 \item
@@
 \item
-  Left: receiver operating characteristic (ROC) curve right: precision
-  recall (PR) curve of trainset and testset. The horizontal green line
-  is the PIK3CA mutation rate (0.11) (B) Correlation between trainset
-  and testset of area under receiver operating characteristic curve
-  (AUROC) and area under precision recall curve (AUPR) among cancer
-  type. The abbreviations are explained in S1 appendix. (C) Correlation
-  between the PIK3CA mutation rate of area under receiver operating
-  characteristic curve (AUROC) and area under precision recall curve
-  (AUPR).
+  Left: receiver operating characteristic (ROC) curve right:
+  precision-recall (PR) curve of trainset and test set. The horizontal
+  green line is the PIK3CA mutation rate (0.11) (B) Correlation between
+  trainset and test set of the area under the receiver operating
+  characteristic curve (AUROC) and area under the precision-recall curve
+  (AUPR) among cancer type. The abbreviations are explained in the S1
+  Appendix. (C) Correlation between the PIK3CA mutation rate of area
+  under the receiver operating characteristic curve (AUROC) and area
+  under the precision-recall curve (AUPR).
 \end{enumerate}
@@
   Top 30 high coefficients of mRNA. (B) Coefficients of cancer types.
-  The abbreviations of cancer types are explained in S1 appendix.
+  The abbreviations of cancer types are explained in the S1 Appendix.
 \end{enumerate}

### 2020-05-14 figure.R

@@ -16,27 +16,27 @@ dataset %>%
   count(variant) %>%
   spread(variant, n) %>%
   replace_na(list(Mutant=0)) %>%
-  mutate(prevalance = Mutant/(Mutant + Wild)) %>%
-  ggplot(aes(x=reorder(type, -prevalance), y=prevalance)) +
+  mutate(prevalence = Mutant/(Mutant + Wild)) %>%
+  ggplot(aes(x=reorder(type, -prevalence), y=prevalence)) +
   geom_bar(stat = "identity") +
   coord_flip() +
   theme_Publication() + 
-  labs(title = "PIK3CA Prevalance",
-       y = "Prevalance rate",
+  labs(title = "PIK3CA prevalence",
+       y = "Prevalence rate",
        x = "Cancer type") 
 dev.off()
 ## Prevalence of PIK3CA 
 
 
-dfPrevalance <- dataset %>%
+dfprevalence <- dataset %>%
   select(variant, type) %>%
   group_by(type) %>%
   count(variant) %>%
   spread(variant, n) %>%
   replace_na(list(Mutant=0)) %>%
-  mutate(prevalance = Mutant/(Mutant + Wild))
+  mutate(prevalence = Mutant/(Mutant + Wild))
 
-summary(dfPrevalance$prevalance)
+summary(dfprevalence$prevalence)
 summary(dataset$variant)
 ### ROC PR Plots
   
@@ -99,15 +99,15 @@ mutationPR <- mutatioPlotPRTest +
   theme_Publication()
 
 geneCoeff <- coeffPlot +
-  labs(title = "Coefficiency of genes",
-       y = "Coefficiency",
+  labs(title = "Coefficient of genes",
+       y = "Coefficient",
        x = "Gene") +
   scale_colour_Publication() +
   theme_Publication()
 
 typeCoeff <- typePlot +
-  labs(title = "Coefficiency of cancer type",
-       y = "Coefficiency",
+  labs(title = "Coefficient of cancer type",
+       y = "Coefficient",
        x = "Cancer type") +
   scale_colour_Publication() +
   theme_Publication()
@@ -130,8 +130,8 @@ dev.off()
 
 
 typePlot +
-  labs(title = "Coefficiency of cancer type",
-       x = "Coefficiency",
+  labs(title = "Coefficient of cancer type",
+       x = "Coefficient",
        y = "Cancer type") +
   scale_colour_Publication() +
   theme_Publication()

### 2020-05-20 plosOne/S1Appendix.Rmd
author edit

@@
 title: "Supplimentary"
-author: "JK"
-date: '2020 4 3 '
 output: html_document
@@
 ```
-
-Data repository
-
-Pancancer TCGA RNA data: http://api.gdc.cancer.gov/data/3586c0da-64d0-4b74-a449-5ff4d9136611
 

### 2020-05-20 plosOne/plosOne.Rmd
author edit

@@
     affiliation: Department of Hospital Pathology, Seoul St. Mary’s Hospital, College of Medicine, The Catholic University of Korea, Seoul, South Korea
-    corresponding: jkang.alien@gmail.com
+
+  - name: Ahwon Lee
+    email: klee@catholic.ac.kr
+    affiliation: Department of Hospital Pathology, Seoul St. Mary’s Hospital, College of Medicine, The Catholic University of Korea, Seoul, South Korea
+
+  - name: Youn Soo Lee
+    email: lys9908@catholic.ac.kr
+    affiliation: Department of Hospital Pathology, Seoul St. Mary’s Hospital, College of Medicine, The Catholic University of Korea, Seoul, South Korea
+    corresponding: lys9908@catholic.ac.kr
+
 abstract: | 
+  Breast cancer with PIK3CA mutation has been approved to use PIK3CA inhibitor in hormone receptor-positive HER2 negative subtype.
+We apply a supervised elastic net penalized logistic regression model in prediction PIK3CA mutation from gene expression data.
+We used the TCGA pan-cancer dataset.
+Penalized logistic regression was applied to prediction modeling.
+10-fold cross-validation with targe variable stratification was done over the hyperparameter grid.
+10845 cases were available for both PIK3CA mutation and mRNA expression data.
+In 10-fold cross-validation, the model with $\lambda$ = 0.01 and $\alpha$ = 1.0 (Ridge regression) showed best performance in terms of AUROC.
+The final model was trained with the selected hyperparameters with the entire trainset.
+The trainset AUROC was 0.93 and the testset AUROC was 0.84.
+The AUPR of trainset was 0.66 and the testset AUPR was 0.39.
+Both IGFR1R and PTEN are key players in the tyrosine kinase pathway.
+The cancer type was an important predictor.
+Our study suggests that the prediction of genomic alteration with gene expression data is possible with good performance.
+
 bibliography: [mybibfile.bib, pik3ca.bib, rpackage.bib, genes.bib] 
@@
 # Introduction
-
-- Brevity
-- Logic and clarity
-- Clean typing
-
-- The problem  
-  * PIK3CA mutation in selecting drug
 
@@
 
-- The main message answers the question and main supporting evidence
-
 Our model showed good performance to predict the PIK3CA mutation of various cancer types.
@@
 It might be a challenging prediction of genomic alterations that are infrequent or have a weak effect on gene expression.
-
-- Critical assessment opinions on any shortcomings in study design
 
@@
 
-- limitations in methods
-
-  * Case imbalance  
-  * flaws in analysis  
-  * validity of assumption
-
-- Comparison with other studies where inconsistencies are discussed  
-- Evaluate the results - not the authors
-
 Our PIK3CA prediction model performed similar with the RAS activation prediction model of the previous study in terms of AUROC (0.84).
@@
 In the subanalysis for performance of each cancer types, the cancer types with higher PIK3CA mutation rate showe better AUPR.
-The imbalanced data of our model
 
@@
 [@heGeneSignaturesAssociated2020]
-
-- Conclusions comments on possible biological or clinical implications and suggestions for further research.
 
@@
 - S1 Appendix.
-- S2 Figure.  
-- S1 Table.  
-- S2 Table.  
-- S3 Table.
 

### 2020-05-20 plosOne/plosOne.tex
author edit

@@
 Jun Kang\textsuperscript{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital, College of
+Medicine, The Catholic University of Korea, Seoul, South Korea}},
+Ahwon Lee\textsuperscript{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital, College of
+Medicine, The Catholic University of Korea, Seoul, South Korea}},
+Youn Soo Lee\textsuperscript{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital, College of
 Medicine, The Catholic University of Korea, Seoul, South Korea}}\textsuperscript{*}\\
@@
 \bigskip
-* Corresponding author: jkang.alien@gmail.com\\
+* Corresponding author: lys9908@catholic.ac.kr\\
 \end{flushleft}
 % Please keep the abstract below 300 words
+\section*{Abstract}
+Breast cancer with PIK3CA mutation has been approved to use PIK3CA
+inhibitor in hormone receptor-positive HER2 negative subtype. We apply a
+supervised elastic net penalized logistic regression model in prediction
+PIK3CA mutation from gene expression data. We used the TCGA pan-cancer
+dataset. Penalized logistic regression was applied to prediction
+modeling. 10-fold cross-validation with targe variable stratification
+was done over the hyperparameter grid. 10845 cases were available for
+both PIK3CA mutation and mRNA expression data. In 10-fold
+cross-validation, the model with \(\lambda\) = 0.01 and \(\alpha\) = 1.0
+(Ridge regression) showed best performance in terms of AUROC. The final
+model was trained with the selected hyperparameters with the entire
+trainset. The trainset AUROC was 0.93 and the testset AUROC was 0.84.
+The AUPR of trainset was 0.66 and the testset AUPR was 0.39. Both IGFR1R
+and PTEN are key players in the tyrosine kinase pathway. The cancer type
+was an important predictor. Our study suggests that the prediction of
+genomic alteration with gene expression data is possible with good
+performance.
 
@@
 
-\begin{itemize}
-\item
-  Brevity
-\item
-  Logic and clarity
-\item
-  Clean typing
-\item
-  The problem
-
-  \begin{itemize}
-  \tightlist
-  \item
-    PIK3CA mutation in selecting drug
-  \end{itemize}
-\end{itemize}
-
 Targeted therapy becomes a standard treatment in many cancer patients.
@@
 
-The machine learning approach has been actively researched to detect
-genomic alterations. Machine learning can build a prediction model from
-a large number of predictors such as radiolomic features {[}1{]},
-pathology image {[}2{]} or gene expression data {[}3{]}. Because the
-most direct genomic test is more specific and sensitive than predictive
-models, the machine learning approach might have a limited role in
-clinical practice. However, the machine learning prediction can be a
-second-best when the direct test fails.
+It has been actively researched that the machine learnig approach can be
+applied to detect genomic alterations. Machine learning algorithm can
+build a prediction model from a large number of predictors such as
+radiolomic features {[}1{]}, pathology image {[}2{]} or gene expression
+data {[}3{]}. Because the most direct genomic test is more specific and
+sensitive than predictive models, the machine learning approach might
+have a limited role in clinical practice. However, the machine learning
+prediction can be a second-best when the direct test fails.
 
@@
 dataset is publically available. PIK3CA mutation data was got using
-cgdsr rpackage.{[}8{]} Gene expression data was got from the GDAC
-firehose using the RTCGAToolbox R package. {[}9{]}
-
-10845 cases were available for both PIK3CA mutation and mRNA expression
-data. 5128 out of 20502 genes were included in the modeling process
-after filtering with median absolute deviation as described at the
-modeling process method. 33 cancer type dummy variables were included in
-predictor variables.
+cgdsr rpackage.{[}8{]} Gene expression data was downloaded from the
+National Cancer Institute (NCI)'s Genomic Data Commons (GDC) website
+that archives data used for The Pan-Cancer Atlas initiative.
+(https://gdc.cancer.gov/about-data/publications/pancanatlas)
 
@@
 
-To narrow down potential predictors, Genes with a large the median
-absolute deviation (more than third-quartiles) were selected.
-Yeo-Johnson transformation was done to correct skewness. Centering and
-scaling were done. All preprocessing was done using recipe r package.
-{[}10{]} Penalized logistic regression was applied to prediction
-modeling. 10-fold cross-validation with targe variable stratification
-was done over the hyperparameter grid: \(\lambda\) \{\(10^{-5}\),
+To narrow down potential predictors, genes with a large the median
+absolute deviation (more than third-quartiles) were selected. 33 cancer
+type dummy variables were included in predictor variables. Yeo-Johnson
+transformation was done to correct skewness. Centering and scaling were
+done. All preprocessing was done using recipe r package. {[}9{]}
+Penalized logistic regression was applied to prediction modeling.
+10-fold cross-validation with targe variable stratification was done
+over the hyperparameter grid: \(\lambda\) \{\(10^{-5}\),
 \(10^{-4}\),\(10^{-3}\),\(10^{-2}\),\(10^{-1}\), \(10^{0}\)\},
@@
 (\((1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1\)).
-{[}11{]}
+{[}10{]}
 
@@
 operating characteristic (ROC) and area under the ROC (AUROC) and
-precision-recall (PR) and area under the curve (AUPR). The prevalence of
-PIK3CA mutation is low. The low prevalence of PIK3CA results in an
-imbalanced dataset which makes the prediction difficult. The AUROC is
-optimistic in terms of performance. The AUPR is more informative than
-AUROC on imbalanced datasets. {[}12{]} The modeling process was done
-with tidymodels rpackage. {[}13{]}
+precision-recall (PR) and area under the curve (AUPR). The AUPR is more
+informative than AUROC on imbalanced datasets. {[}11{]} The modeling
+process and accessing model performance were done with tidymodels
+rpackage. {[}12{]}
 
@@
 
-\hypertarget{prevalence-rate-of-pik3ca-mutation}{%
-\subsection{prevalence rate of PIK3CA
-mutation}\label{prevalence-rate-of-pik3ca-mutation}}
+\hypertarget{dataset-summary}{%
+\subsection{Dataset summary}\label{dataset-summary}}
 
-The prevalence rate of PIK3CA was 0.11 in all cases. The PIK3CA
-prevalence rate of each cancer type varied. The median prevalence rate
-of PIK3CA of each cancer type was 0.03 (range 0-0.33) (Figure 1).
+10845 cases were available for both PIK3CA mutation and mRNA expression
+data. 5128 out of 20502 genes were included in the modeling process
+after filtering with median absolute deviation as described at the
+modeling process method. The prevalence rate of PIK3CA was 0.11 in all
+cases. The PIK3CA prevalence rate of each cancer type varied. The median
+prevalence rate of PIK3CA of each cancer type was 0.03 (range 0-0.33)
+(Figure 1).
 
@@
 AUROC. The final model was trained with the selected hyperparameters
-with the entire trainset.
-
-The trainset AUROC was 0.93 and the testset AUROC was 0.84. The AUPR of
-trainset was 0.66 and the testset AUPR was 0.39. (Figure 1A)
+with the entire trainset. The trainset AUROC was 0.93 and the testset
+AUROC was 0.84. The AUPR of trainset was 0.66 and the testset AUPR was
+0.39. (Figure 2A)
 
@@
 were positively correlated between the train set and test set in cancer
-type subanalysis. (Figure 1B) The AUPR was high in cancer type with high
-PIK3CA mutation rate such as colon, breast, Uterus cancer types. The
+type subanalysis. (Figure 2B) The AUPR was high in cancer type with high
+PIK3CA mutation rate such as colon, breast, uterus cancer types. The
 AUROC did not correlate with the PIK3CA mutation rate of each cancer
-type. (Figure 1C)
+type. (Figure 2C)
 
@@
 
-Figure 2 shows the top 30 important predictors. The coefficient is the
+Figure 3 shows the top 30 important predictors. The coefficient is the
 parameter of the predictor which represents the effect of the predictor
 on prediction. IGF1R mRNA expression was the strongest negative
-predictor and PTEN was the strongest positive predictor. Both genes and
-PIK3CA are key players in the tyrosine kinase pathway. The cancer type
-was an important predictor. Some cancer types including uterine
+predictor and PTEN was the strongest positive predictor. Both IGFR1R and
+PTEN are key players in the tyrosine kinase pathway. The cancer type was
+an important predictor. Some cancer types including uterine
 carcinosarcoma (UCS), bladder urothelial carcinoma (BLCA), pancreatic
@@
 
-\begin{itemize}
-\tightlist
-\item
-  The main message answers the question and main supporting evidence
-\end{itemize}
-
 Our model showed good performance to predict the PIK3CA mutation of
-various cancer types. This result shows that the supervised elastic net
-penalized logistic regression model can be applied not only to the RAS
-activation pathway but also PIK3CA mutation. Both the RAS activation
-pathway and PIK3CA mutation are important and common cancer genomic
-alterations. They have a significant effect on gene expression in cancer
-cells. It might be a challenging prediction of genomic alterations that
-are infrequent or have a weak effect on gene expression.
-
-\begin{itemize}
-\tightlist
-\item
-  Critical assessment opinions on any shortcomings in study design
-\end{itemize}
+various cancer types. This result suggests that the supervised elastic
+net penalized logistic regression model can be applied not only to the
+RAS activation pathway but also other genomic alterations. Both the RAS
+activation pathway and PIK3CA mutation are important and common cancer
+genomic alterations. They have a significant effect on gene expression
+in cancer cells. It can not be generalized that the supervised elastic
+net penalized logistic regression model can be applied to other genomic
+alteration. It might be a challenging prediction of genomic alterations
+that are infrequent or have a weak effect on gene expression.
 
@@
 regarding data preprocessing. Methods for cancer gene expression have
-been developed for research. To statistical analysis, the gene
-expression data is processed between-sample normalization to remove the
-batch effect. If the model has been trained from the between-sample
-normalization, a new sample can not be exactly processed like trainset.
-A model based on gene expression TCGA pan-cancer dataset has a
-limitation on preprocessing. It is necessary for developing a
+been developed for research. The gene expression data is processed
+between-sample normalization to remove the batch effect. If the model
+has been trained from the between-sample normalization, a new sample can
+not be exactly preprocessed with normalization which was done on
+trainset. A model based on gene expression TCGA pan-cancer dataset has a
+limitation on data preprocessing. It is necessary for developing a
 preprocessing method that is independent with a dataset to apply the
@@
 
-\begin{itemize}
-\item
-  limitations in methods
-
-  \begin{itemize}
-  \tightlist
-  \item
-    Case imbalance\\
-  \item
-    flaws in analysis\\
-  \item
-    validity of assumption
-  \end{itemize}
-\item
-  Comparison with other studies where inconsistencies are discussed\\
-\item
-  Evaluate the results - not the authors
-\end{itemize}
-
-Our PIK3CA prediction model performed better than the RAS activation
-prediction model of the previous study in terms of both AUROC (0.84 vs
-0.75) and AUPR (0.39 vs 0.24) on test set prediction. Our test set is
-corresponding to the samples initially filtered from training. The
-target variable of our study is more specific than the previous study.
-The specific important mutations can affect stronger downstream gene
-expression than the broad events pathway activation. The previous study
-might be a more difficult prediction problem.
+Our PIK3CA prediction model performed similar with the RAS activation
+prediction model of the previous study in terms of AUROC (0.84). However
+AUPR of our model for PIK3CA is lower than model for RAS activation
+(0.39 vs 0.63). The reason of lower AUPR of our model can be explained
+by low prevalence rate of PIK3CA mutation and imbalanced dataset. The
+model for RAS activation trained with cancer types with more than 0.05
+prevalence of RAS activation to avoid imbalance classification problem.
+We included all cancer types into modeling process. The lower prevalance
+rate of target variable mean our dataset has a lower baseline of AUPR.
+In the subanalysis for performance of each cancer types, the cancer
+types with higher PIK3CA mutation rate showe better AUPR.
 
@@
 predictor which have negative and positive predictive power. IGF1R is a
-tyrosine kinase receptor that activates PI3K. {[}14{]} Insulin receptor
-substrate-2 (IRS2) is the adaptor protein of IGF1R. {[}15{]} PTEN is an
+tyrosine kinase receptor that activates PI3K. {[}13{]} Insulin receptor
+substrate-2 (IRS2) is the adaptor protein of IGF1R. {[}14{]} PTEN is an
 important regulator of PIP\textsubscript{3} by dephosphorylating
@@
 logarithmic gene expression. The model showed good performance AUROC
-0.71 in an independent test set. {[}4,16{]} Another study predicted copy
+0.71 in an independent test set. {[}4,15{]} Another study predicted copy
 number alterations with gene expression using a multinomial logistic
@@
 of 0.997. The gene-level prediction was good with an AUROC 0.75.
-{[}17{]} A Hidden Markov Model Approach for Prediction of Genomic
-Alterations from Gene Expression Profiling.{[}18{]} A logistic
+{[}16{]} A Hidden Markov Model Approach for Prediction of Genomic
+Alterations from Gene Expression Profiling.{[}17{]} A logistic
 regression model was used for MYCN gene amplification in neuroblastoma.
-{[}19{]}
-
-\begin{itemize}
-\tightlist
-\item
-  Conclusions comments on possible biological or clinical implications
-  and suggestions for further research.
-\end{itemize}
+{[}18{]}
 
@@
   characteristic curve (AUROC) and area under the precision-recall curve
-  (AUPR) among cancer type. The abbreviations are explained in the S1
-  Appendix. (C) Correlation between the PIK3CA mutation rate of area
-  under the receiver operating characteristic curve (AUROC) and area
-  under the precision-recall curve (AUPR).
+  (AUPR) among cancer type. The gray band is 95\% confidence interval.
+  The abbreviations are explained in the S1 Appendix. (C) Correlation
+  between the PIK3CA mutation rate of area under the receiver operating
+  characteristic curve (AUROC) and area under the precision-recall curve
+  (AUPR).
 \end{enumerate}
@@
 \item
-  S1 Appendix.\\
-\item
-  S2 Figure.\\
-\item
-  S1 Table.\\
-\item
-  S2 Table.\\
-\item
-  S3 Table.
+  S1 Appendix.
 \end{itemize}
@@
 
-\leavevmode\hypertarget{ref-samurRTCGAToolboxNewTool2014}{}%
-9. Samur MK. RTCGAToolbox: A new tool for exporting TCGA Firehose data.
-PLoS One. 2014;9(9):e106397.
-
 \leavevmode\hypertarget{ref-kuhnRecipesPreprocessingTools2020}{}%
-10. Kuhn M, Wickham H. Recipes: Preprocessing tools to create design
+9. Kuhn M, Wickham H. Recipes: Preprocessing tools to create design
 matrices. 2020.
@@
 \leavevmode\hypertarget{ref-friedmanRegularizationPathsGeneralized2010}{}%
-11. Friedman J, Hastie T, Tibshirani R. Regularization Paths for
+10. Friedman J, Hastie T, Tibshirani R. Regularization Paths for
 Generalized Linear Models via Coordinate Descent. J Stat Softw. 2010;33:
@@
 \leavevmode\hypertarget{ref-saitoPrecisionRecallPlotMore2015}{}%
-12. Saito T, Rehmsmeier M. The Precision-Recall Plot Is More Informative
+11. Saito T, Rehmsmeier M. The Precision-Recall Plot Is More Informative
 than the ROC Plot When Evaluating Binary Classifiers on Imbalanced
@@
 \leavevmode\hypertarget{ref-kuhnTidymodelsEasilyInstall2020}{}%
-13. Kuhn M, Wickham H. Tidymodels: Easily install and load the
+12. Kuhn M, Wickham H. Tidymodels: Easily install and load the
 'tidymodels' packages. 2020.
@@
 \leavevmode\hypertarget{ref-leroithInsulinlikeGrowthFactor2003}{}%
-14. LeRoith D, Roberts CT. The insulin-like growth factor system and
+13. LeRoith D, Roberts CT. The insulin-like growth factor system and
 cancer. Cancer Letters. 2003;195: 127--137.
@@
 \leavevmode\hypertarget{ref-heInteractionInsulinReceptor1996}{}%
-15. He W, Craparo A, Zhu Y, O'Neill TJ, Wang L-M, Pierce JH, et al.
+14. He W, Craparo A, Zhu Y, O'Neill TJ, Wang L-M, Pierce JH, et al.
 Interaction of Insulin Receptor Substrate-2 (IRS-2) with the Insulin and
@@
 \leavevmode\hypertarget{ref-cizkovaGeneExpressionProfiling2010}{}%
-16. Cizkova M, Cizeron-Clairac G, Vacher S, Susini A, Andrieu C,
+15. Cizkova M, Cizeron-Clairac G, Vacher S, Susini A, Andrieu C,
 Lidereau R, et al. Gene Expression Profiling Reveals New Aspects of
@@
 \leavevmode\hypertarget{ref-muCNAPEMachineLearning2019}{}%
-17. Mu Q, Wang J. CNAPE: A Machine Learning Method for Copy Number
+16. Mu Q, Wang J. CNAPE: A Machine Learning Method for Copy Number
 Alteration Prediction from Gene Expression. IEEE/ACM Transactions on
@@
 \leavevmode\hypertarget{ref-gengHiddenMarkovModel2008}{}%
-18. Geng H, Ali HH, Chan WC. A Hidden Markov Model Approach for
+17. Geng H, Ali HH, Chan WC. A Hidden Markov Model Approach for
 Prediction of Genomic Alterations from Gene Expression Profiling. In:
@@
 \leavevmode\hypertarget{ref-heGeneSignaturesAssociated2020}{}%
-19. He X, Qin C, Zhao Y, Zou L, Zhao H, Cheng C. Gene signatures
+18. He X, Qin C, Zhao Y, Zou L, Zhao H, Cheng C. Gene signatures
 associated with genomic aberrations predict prognosis in neuroblastoma.

### 2020-05-27 plosOne/plosOne.Rmd
first draft

@@
 ---
-title: Prediction of PIK3CA mutation with gene expression
+title: Prediction of PIK3CA mutation with gene expression in cancer
 author:
@@
     email: jkang.alien@gmail.com
-    affiliation: Department of Hospital Pathology, Seoul St. Mary’s Hospital, College of Medicine, The Catholic University of Korea, Seoul, South Korea
-
+    affiliation: Department of Hospital Pathology, Seoul St. Mary’s Hospital
   - name: Ahwon Lee
     email: klee@catholic.ac.kr
-    affiliation: Department of Hospital Pathology, Seoul St. Mary’s Hospital, College of Medicine, The Catholic University of Korea, Seoul, South Korea
-
+    affiliation: Department of Hospital Pathology, Seoul St. Mary’s Hospital
   - name: Youn Soo Lee
     email: lys9908@catholic.ac.kr
-    affiliation: Department of Hospital Pathology, Seoul St. Mary’s Hospital, College of Medicine, The Catholic University of Korea, Seoul, South Korea
+    affiliation: Department of Hospital Pathology, Seoul St. Mary’s Hospital
     corresponding: lys9908@catholic.ac.kr
+address:
+  - code: Department of Hospital Pathology, Seoul St. Mary’s Hospital
+    address: Department of Hospital Pathology, Seoul St. Mary’s Hospital, College of Medicine, The Catholic University of Korea, Seoul, South Korea
 
@@
 We apply a supervised elastic net penalized logistic regression model in prediction PIK3CA mutation from gene expression data.
-We used the TCGA pan-cancer dataset.
-Penalized logistic regression was applied to prediction modeling.
-10-fold cross-validation with targe variable stratification was done over the hyperparameter grid.
+Penalized logistic regression was applied to prediction modeling using the TCGA pan-cancer dataset.
 10845 cases were available for both PIK3CA mutation and mRNA expression data.
@@
 The AUPR of trainset was 0.66 and the testset AUPR was 0.39.
-Both IGFR1R and PTEN are key players in the tyrosine kinase pathway.
-The cancer type was an important predictor.
+The cancer types were the most important predictors.
+Both IGFR1R and PTEN were most important among gene expression predictors.
 Our study suggests that the prediction of genomic alteration with gene expression data is possible with good performance.
@@
 Targeted therapy becomes a standard treatment in many cancer patients.
-Much targeted therapy requires a test for a specific cancer genomic alteration, to treat the patients.
+The majority of targeted therapy requires a test for a specific cancer genomic alteration, to treat the patients.
 Many direct tests for the genomic alteration have been developed and prooved for their clinical utility to find which patients will be responding to the targeted therapy.
@@
 
-Prediction RAS pathway activation with gene expression data was done in the previous study.
+Prediction of RAS pathway activation with gene expression data was done in the previous study.
 @way_2018_machine_cellreports They trained pan-cancer The Cancer Genome Atlas (TCGA) data with a supervised elastic net penalized logistic regression classifier with stochastic gradient descent.
@@
 The authors suggested their approach can be applied to other genomic alterations.
+
+Breast cancer with PIK3CA mutation has been approved to use PIK3CA inhibitor in hormone receptor-positive HER2 negative subtype.
+@andre_2019_alpelisib_nengljmeda The PIK3CA mutation is the second most driver mutation after TP53.
+The PIK3CA mutation is most frequently founded in endometrial carcinoma (45%), and followed by breast invasive carcinoma (24%), cervical squamous cell carcinoma, and endocervical adenocarcinoma (20%) and colon adenocarcinoma (16%).
 
@@
 
-Breast cancer with PIK3CA mutation has been approved to use PIK3CA inhibitor in hormone receptor-positive HER2 negative subtype.
-@andre_2019_alpelisib_nengljmeda The PIK3CA mutation is the second most driver mutation after TP53.
-The PIK3CA mutation is most frequently founded in endometrial carcinoma (45%), and followed by breast invasive carcinoma (24%), cervical squamous cell carcinoma, and endocervical adenocarcinoma (20%) and colon adenocarcinoma (16%).
-
 We apply a supervised elastic net penalized logistic regression model in prediction PIK3CA mutation.
-The purpose of this study is to know this prediction model approach can be applied not only to RAS pathway activation but also PIK3CA mutation across many cancer types.
+The purpose of this study is to know whether this prediction model approach can be applied not only to RAS pathway activation but also PIK3CA mutation across many cancer types.
 
@@
 TCGA is a cancer genomic consortium that archives data of exome sequencing, gene expression, DNA methylation, protein expression, and clinical data of more than 10000 cancer samples across 33 common cancer types.
-The gene expression TCGA pan-cancer dataset is batch-corrected with normalization.
 The TCGA dataset is publically available.
 PIK3CA mutation data was got using cgdsr rpackage.[@jacobsenCgdsrRbasedAPI2019] Gene expression data was downloaded from the National Cancer Institute (NCI)'s Genomic Data Commons (GDC) website that archives data used for The Pan-Cancer Atlas initiative.
-(https://gdc.cancer.gov/about-data/publications/pancanatlas)
+(https://gdc.cancer.gov/about-data/publications/pancanatlas) The gene expression TCGA pan-cancer dataset is batch-corrected with normalization.
 
 The target variable was the PIK3CA mutation status.
-The status of  PIK3CA was considered as positive when the case has following PIK3CA variants which are the target variables of the therascreen PIK3CA RGQ PCR Kit; C420R, E542K, E545A, E545D, E545G, E545K, Q546E, Q546R, H1047L, H1047R, H1047Y.
+The status of  PIK3CA was considered as positive when the case has following PIK3CA variants which are the target variants of the therascreen PIK3CA RGQ PCR Kit; C420R, E542K, E545A, E545D, E545G, E545K, Q546E, Q546R, H1047L, H1047R, H1047Y.
 The therascreen PIK3CA RGQ PCR Kit was approved as a companion diagnostics to treat with PIK3CA inhibitor by the U.S.
@@
 
-Model performance was evaluated with the area under the receiver operating characteristic (ROC) and area under the ROC (AUROC) and precision-recall (PR) and area under the curve (AUPR).
+Model performance was evaluated with the AUROC curve and the AUPR curve.
 The AUPR is more informative than AUROC on imbalanced datasets.
@@
 Both IGFR1R and PTEN are key players in the tyrosine kinase pathway.
-The cancer type was an important predictor.
+The cancer types were important predictors.
 Some cancer types including uterine carcinosarcoma (UCS), bladder urothelial carcinoma (BLCA), pancreatic adenocarcinoma (PAAD), lymphoid neoplasm diffuse large B-cell lymphoma (DLBC) are the strongest predictors.
@@
 Prediction modeling from the TCGA pan-cancer dataset has limitations regarding data preprocessing.
-Methods for cancer gene expression have been developed for research.
 The gene expression data is processed between-sample normalization to remove the batch effect.
@@
 
-Our model includes cancer type predictor and they are stronger than gene expression data.
-The varying prevalence of PIC3CA mutation across cancer type might reason for the strong cancer type predictor.
+Our model includes cancer type as predictors and they are stronger predictors than gene expression.
+The varying prevalence of PIC3CA mutation across cancer type might reason for the strong predictive power of cancer type.
 If the cancer type was wrong or can not be determined, our model performance can be poor.
@@
 IGF1R is a tyrosine kinase receptor that activates PI3K.
-[@leroithInsulinlikeGrowthFactor2003] Insulin receptor substrate-2 (IRS2) is the adaptor protein of IGF1R.
-[@heInteractionInsulinReceptor1996] PTEN is an important regulator of PIP~3~ by dephosphorylating  PIP~3~ in contrast PI3K.[@cantleyPhosphoinositide3KinasePathway2002]
+[@leroithInsulinlikeGrowthFactor2003] PTEN is an important regulator of PIP~3~ by dephosphorylating  PIP~3~ in contrast PI3K.[@cantleyPhosphoinositide3KinasePathway2002]
 
@@
 
-- Figure 1. prevalence rate of PIK3CA across cancer types
+- Figure 1. Prevalence rate of PIK3CA across cancer types.
 The abbreviations of cancer types are explained in the S1 Appendix.
@@
 
-(A) Left: receiver operating characteristic (ROC) curve right: precision-recall (PR) curve of trainset and test set.
-The horizontal green line is the PIK3CA mutation rate (0.11)  (B) Correlation between trainset and test set of the area under the receiver operating characteristic curve (AUROC) and area under the precision-recall curve (AUPR) among cancer type.
+(A) Left: receiver operating characteristic (ROC) curve.
+right: precision-recall (PR) curve of trainset and test set.
+The horizontal green line is the PIK3CA mutation rate (0.11)  (B) Correlation between trainset and test set of the area under the receiver operating characteristic curve (AUROC) and the area under the precision-recall curve (AUPR) among cancer type.
 The gray band is 95% confidence interval.
 The abbreviations are explained in the S1 Appendix.
-(C) Correlation between the PIK3CA mutation rate of area under the receiver operating characteristic curve (AUROC) and area under the precision-recall curve (AUPR).
+(C) Correlation between the PIK3CA mutation rate of the AUROC and the AUPR.
 

### 2020-05-27 plosOne/plosOne.tex
first draft

@@
 \newcommand{\getAff}[1]{
-  \getIndex{#1}{}
+  \getIndex{#1}{Department of Hospital Pathology, Seoul St.~Mary's Hospital}
 }
@@
 {\Large
-\textbf\newline{Prediction of PIK3CA mutation with gene expression} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+\textbf\newline{Prediction of PIK3CA mutation with gene expression in cancer} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
 }
@@
 \\
-Jun Kang\textsuperscript{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital, College of
-Medicine, The Catholic University of Korea, Seoul, South Korea}},
-Ahwon Lee\textsuperscript{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital, College of
-Medicine, The Catholic University of Korea, Seoul, South Korea}},
-Youn Soo Lee\textsuperscript{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital, College of
-Medicine, The Catholic University of Korea, Seoul, South Korea}}\textsuperscript{*}\\
+Jun Kang\textsuperscript{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital}},
+Ahwon Lee\textsuperscript{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital}},
+Youn Soo Lee\textsuperscript{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital}}\textsuperscript{*}\\
 \bigskip
+\textbf{\getAff{Department of Hospital Pathology, Seoul St.~Mary's Hospital}}Department of Hospital Pathology, Seoul St.~Mary's Hospital, College of
+Medicine, The Catholic University of Korea, Seoul, South Korea\\
 \bigskip
@@
 supervised elastic net penalized logistic regression model in prediction
-PIK3CA mutation from gene expression data. We used the TCGA pan-cancer
-dataset. Penalized logistic regression was applied to prediction
-modeling. 10-fold cross-validation with targe variable stratification
-was done over the hyperparameter grid. 10845 cases were available for
-both PIK3CA mutation and mRNA expression data. In 10-fold
-cross-validation, the model with \(\lambda\) = 0.01 and \(\alpha\) = 1.0
-(Ridge regression) showed best performance in terms of AUROC. The final
-model was trained with the selected hyperparameters with the entire
-trainset. The trainset AUROC was 0.93 and the testset AUROC was 0.84.
-The AUPR of trainset was 0.66 and the testset AUPR was 0.39. Both IGFR1R
-and PTEN are key players in the tyrosine kinase pathway. The cancer type
-was an important predictor. Our study suggests that the prediction of
-genomic alteration with gene expression data is possible with good
-performance.
+PIK3CA mutation from gene expression data. Penalized logistic regression
+was applied to prediction modeling using the TCGA pan-cancer dataset.
+10845 cases were available for both PIK3CA mutation and mRNA expression
+data. In 10-fold cross-validation, the model with \(\lambda\) = 0.01 and
+\(\alpha\) = 1.0 (Ridge regression) showed best performance in terms of
+AUROC. The final model was trained with the selected hyperparameters
+with the entire trainset. The trainset AUROC was 0.93 and the testset
+AUROC was 0.84. The AUPR of trainset was 0.66 and the testset AUPR was
+0.39. The cancer types were the most important predictors. Both IGFR1R
+and PTEN were most important among gene expression predictors. Our study
+suggests that the prediction of genomic alteration with gene expression
+data is possible with good performance.
 
@@
 Targeted therapy becomes a standard treatment in many cancer patients.
-Much targeted therapy requires a test for a specific cancer genomic
-alteration, to treat the patients. Many direct tests for the genomic
-alteration have been developed and prooved for their clinical utility to
-find which patients will be responding to the targeted therapy.
+The majority of targeted therapy requires a test for a specific cancer
+genomic alteration, to treat the patients. Many direct tests for the
+genomic alteration have been developed and prooved for their clinical
+utility to find which patients will be responding to the targeted
+therapy.
 
@@
 
-Prediction RAS pathway activation with gene expression data was done in
-the previous study. {[}5{]} They trained pan-cancer The Cancer Genome
+Prediction of RAS pathway activation with gene expression data was done
+in the previous study. {[}5{]} They trained pan-cancer The Cancer Genome
 Atlas (TCGA) data with a supervised elastic net penalized logistic
@@
 be applied to other genomic alterations.
+
+Breast cancer with PIK3CA mutation has been approved to use PIK3CA
+inhibitor in hormone receptor-positive HER2 negative subtype. {[}6{]}
+The PIK3CA mutation is the second most driver mutation after TP53. The
+PIK3CA mutation is most frequently founded in endometrial carcinoma
+(45\%), and followed by breast invasive carcinoma (24\%), cervical
+squamous cell carcinoma, and endocervical adenocarcinoma (20\%) and
+colon adenocarcinoma (16\%).
 
@@
 (AKT) which is a serine/threonine-specific protein kinase. AKT inhibits
-apoptosis and promotes cell proliferation. {[}6{]}
-
-Breast cancer with PIK3CA mutation has been approved to use PIK3CA
-inhibitor in hormone receptor-positive HER2 negative subtype. {[}7{]}
-The PIK3CA mutation is the second most driver mutation after TP53. The
-PIK3CA mutation is most frequently founded in endometrial carcinoma
-(45\%), and followed by breast invasive carcinoma (24\%), cervical
-squamous cell carcinoma, and endocervical adenocarcinoma (20\%) and
-colon adenocarcinoma (16\%).
+apoptosis and promotes cell proliferation. {[}7{]}
 
 We apply a supervised elastic net penalized logistic regression model in
-prediction PIK3CA mutation. The purpose of this study is to know this
-prediction model approach can be applied not only to RAS pathway
+prediction PIK3CA mutation. The purpose of this study is to know whether
+this prediction model approach can be applied not only to RAS pathway
 activation but also PIK3CA mutation across many cancer types.
@@
 methylation, protein expression, and clinical data of more than 10000
-cancer samples across 33 common cancer types. The gene expression TCGA
-pan-cancer dataset is batch-corrected with normalization. The TCGA
-dataset is publically available. PIK3CA mutation data was got using
-cgdsr rpackage.{[}8{]} Gene expression data was downloaded from the
-National Cancer Institute (NCI)'s Genomic Data Commons (GDC) website
-that archives data used for The Pan-Cancer Atlas initiative.
-(https://gdc.cancer.gov/about-data/publications/pancanatlas)
+cancer samples across 33 common cancer types. The TCGA dataset is
+publically available. PIK3CA mutation data was got using cgdsr
+rpackage.{[}8{]} Gene expression data was downloaded from the National
+Cancer Institute (NCI)'s Genomic Data Commons (GDC) website that
+archives data used for The Pan-Cancer Atlas initiative.
+(https://gdc.cancer.gov/about-data/publications/pancanatlas) The gene
+expression TCGA pan-cancer dataset is batch-corrected with
+normalization.
 
@@
 was considered as positive when the case has following PIK3CA variants
-which are the target variables of the therascreen PIK3CA RGQ PCR Kit;
+which are the target variants of the therascreen PIK3CA RGQ PCR Kit;
 C420R, E542K, E545A, E545D, E545G, E545K, Q546E, Q546R, H1047L, H1047R,
@@
 
-Model performance was evaluated with the area under the receiver
-operating characteristic (ROC) and area under the ROC (AUROC) and
-precision-recall (PR) and area under the curve (AUPR). The AUPR is more
-informative than AUROC on imbalanced datasets. {[}11{]} The modeling
-process and accessing model performance were done with tidymodels
-rpackage. {[}12{]}
+Model performance was evaluated with the AUROC curve and the AUPR curve.
+The AUPR is more informative than AUROC on imbalanced datasets. {[}11{]}
+The modeling process and accessing model performance were done with
+tidymodels rpackage. {[}12{]}
 
@@
 predictor and PTEN was the strongest positive predictor. Both IGFR1R and
-PTEN are key players in the tyrosine kinase pathway. The cancer type was
-an important predictor. Some cancer types including uterine
+PTEN are key players in the tyrosine kinase pathway. The cancer types
+were important predictors. Some cancer types including uterine
 carcinosarcoma (UCS), bladder urothelial carcinoma (BLCA), pancreatic
@@
 Prediction modeling from the TCGA pan-cancer dataset has limitations
-regarding data preprocessing. Methods for cancer gene expression have
-been developed for research. The gene expression data is processed
+regarding data preprocessing. The gene expression data is processed
 between-sample normalization to remove the batch effect. If the model
@@
 
-Our model includes cancer type predictor and they are stronger than gene
-expression data. The varying prevalence of PIC3CA mutation across cancer
-type might reason for the strong cancer type predictor. If the cancer
-type was wrong or can not be determined, our model performance can be
-poor.
+Our model includes cancer type as predictors and they are stronger
+predictors than gene expression. The varying prevalence of PIC3CA
+mutation across cancer type might reason for the strong predictive power
+of cancer type. If the cancer type was wrong or can not be determined,
+our model performance can be poor.
 
@@
 predictor which have negative and positive predictive power. IGF1R is a
-tyrosine kinase receptor that activates PI3K. {[}13{]} Insulin receptor
-substrate-2 (IRS2) is the adaptor protein of IGF1R. {[}14{]} PTEN is an
+tyrosine kinase receptor that activates PI3K. {[}13{]} PTEN is an
 important regulator of PIP\textsubscript{3} by dephosphorylating
-PIP\textsubscript{3} in contrast PI3K.{[}6{]}
+PIP\textsubscript{3} in contrast PI3K.{[}7{]}
 
@@
 logarithmic gene expression. The model showed good performance AUROC
-0.71 in an independent test set. {[}4,15{]} Another study predicted copy
+0.71 in an independent test set. {[}4,14{]} Another study predicted copy
 number alterations with gene expression using a multinomial logistic
@@
 of 0.997. The gene-level prediction was good with an AUROC 0.75.
-{[}16{]} A Hidden Markov Model Approach for Prediction of Genomic
-Alterations from Gene Expression Profiling.{[}17{]} A logistic
+{[}15{]} A Hidden Markov Model Approach for Prediction of Genomic
+Alterations from Gene Expression Profiling.{[}16{]} A logistic
 regression model was used for MYCN gene amplification in neuroblastoma.
-{[}18{]}
+{[}17{]}
 
@@
 \item
-  Figure 1. prevalence rate of PIK3CA across cancer types The
+  Figure 1. Prevalence rate of PIK3CA across cancer types. The
   abbreviations of cancer types are explained in the S1 Appendix.
@@
 \item
-  Left: receiver operating characteristic (ROC) curve right:
+  Left: receiver operating characteristic (ROC) curve. right:
   precision-recall (PR) curve of trainset and test set. The horizontal
@@
   trainset and test set of the area under the receiver operating
-  characteristic curve (AUROC) and area under the precision-recall curve
-  (AUPR) among cancer type. The gray band is 95\% confidence interval.
-  The abbreviations are explained in the S1 Appendix. (C) Correlation
-  between the PIK3CA mutation rate of area under the receiver operating
-  characteristic curve (AUROC) and area under the precision-recall curve
-  (AUPR).
+  characteristic curve (AUROC) and the area under the precision-recall
+  curve (AUPR) among cancer type. The gray band is 95\% confidence
+  interval. The abbreviations are explained in the S1 Appendix. (C)
+  Correlation between the PIK3CA mutation rate of the AUROC and the
+  AUPR.
 \end{enumerate}
@@
 
-\leavevmode\hypertarget{ref-cantleyPhosphoinositide3KinasePathway2002}{}%
-6. Cantley LC. The Phosphoinositide 3-Kinase Pathway. Science. American
-Association for the Advancement of Science; 2002;296: 1655--1657.
-doi:\href{https://doi.org/10.1126/science.296.5573.1655}{10.1126/science.296.5573.1655}
-
 \leavevmode\hypertarget{ref-andre_2019_alpelisib_nengljmeda}{}%
-7. André F, Ciruelos E, Rubovszky G, Campone M, Loibl S, Rugo HS, et al.
+6. André F, Ciruelos E, Rubovszky G, Campone M, Loibl S, Rugo HS, et al.
 Alpelisib for PIK3CA-Mutated, Hormone ReceptorPositive Advanced Breast
@@
 doi:\href{https://doi.org/10.1056/NEJMoa1813904}{10.1056/NEJMoa1813904}
+
+\leavevmode\hypertarget{ref-cantleyPhosphoinositide3KinasePathway2002}{}%
+7. Cantley LC. The Phosphoinositide 3-Kinase Pathway. Science. American
+Association for the Advancement of Science; 2002;296: 1655--1657.
+doi:\href{https://doi.org/10.1126/science.296.5573.1655}{10.1126/science.296.5573.1655}
 
@@
 
-\leavevmode\hypertarget{ref-heInteractionInsulinReceptor1996}{}%
-14. He W, Craparo A, Zhu Y, O'Neill TJ, Wang L-M, Pierce JH, et al.
-Interaction of Insulin Receptor Substrate-2 (IRS-2) with the Insulin and
-Insulin-like Growth Factor I Receptors EVIDENCE FOR TWO DISTINCT
-PHOSPHOTYROSINE-DEPENDENT INTERACTION DOMAINS WITHIN IRS-2. J Biol Chem.
-American Society for Biochemistry and Molecular Biology; 1996;271:
-11641--11645.
-doi:\href{https://doi.org/10.1074/jbc.271.20.11641}{10.1074/jbc.271.20.11641}
-
 \leavevmode\hypertarget{ref-cizkovaGeneExpressionProfiling2010}{}%
-15. Cizkova M, Cizeron-Clairac G, Vacher S, Susini A, Andrieu C,
+14. Cizkova M, Cizeron-Clairac G, Vacher S, Susini A, Andrieu C,
 Lidereau R, et al. Gene Expression Profiling Reveals New Aspects of
@@
 \leavevmode\hypertarget{ref-muCNAPEMachineLearning2019}{}%
-16. Mu Q, Wang J. CNAPE: A Machine Learning Method for Copy Number
+15. Mu Q, Wang J. CNAPE: A Machine Learning Method for Copy Number
 Alteration Prediction from Gene Expression. IEEE/ACM Transactions on
@@
 \leavevmode\hypertarget{ref-gengHiddenMarkovModel2008}{}%
-17. Geng H, Ali HH, Chan WC. A Hidden Markov Model Approach for
+16. Geng H, Ali HH, Chan WC. A Hidden Markov Model Approach for
 Prediction of Genomic Alterations from Gene Expression Profiling. In:
@@
 \leavevmode\hypertarget{ref-heGeneSignaturesAssociated2020}{}%
-18. He X, Qin C, Zhao Y, Zou L, Zhao H, Cheng C. Gene signatures
+17. He X, Qin C, Zhao Y, Zou L, Zhao H, Cheng C. Gene signatures
 associated with genomic aberrations predict prognosis in neuroblastoma.

### 2020-06-11 plosOne/plosOne.Rmd
PRS

@@
 ---
-title: Prediction of PIK3CA mutation with gene expression in cancer
+title: Prediction of PIK3CA mutations from cancer gene expression data
+short-title: PIK3CA mutations prediction  
 author:
@@
 abstract: | 
-  Breast cancer with PIK3CA mutation has been approved to use PIK3CA inhibitor in hormone receptor-positive HER2 negative subtype.
-We apply a supervised elastic net penalized logistic regression model in prediction PIK3CA mutation from gene expression data.
-Penalized logistic regression was applied to prediction modeling using the TCGA pan-cancer dataset.
-10845 cases were available for both PIK3CA mutation and mRNA expression data.
-In 10-fold cross-validation, the model with $\lambda$ = 0.01 and $\alpha$ = 1.0 (Ridge regression) showed best performance in terms of AUROC.
-The final model was trained with the selected hyperparameters with the entire trainset.
-The trainset AUROC was 0.93 and the testset AUROC was 0.84.
-The AUPR of trainset was 0.66 and the testset AUPR was 0.39.
-The cancer types were the most important predictors.
-Both IGFR1R and PTEN were most important among gene expression predictors.
-Our study suggests that the prediction of genomic alteration with gene expression data is possible with good performance.
+  Breast cancers with PIK3CA mutations can be treated with PIK3CA inhibitors in hormone receptor-positive HER2 negative subtypes.
+We applied a supervised elastic net penalized logistic regression model to predict PIK3CA mutations from gene expression data.
+This regression approach was applied to predict modeling using the TCGA pan-cancer dataset.
+Approximately 10,845 cases were available for PIK3CA mutation and mRNA expression data.
+In 10-fold cross-validation, the model with $\lambda$ = 0.01 and $\alpha$ = 1.0 (Ridge regression) showed the best performance, in terms of area under the receiver operating characteristic (AUROC).
+The final model was developed with selected hyper-parameters using the entire train-set.
+The train-set AUROC was 0.93, and the test-set AUROC was 0.84.
+The area under the precision-recall (AUPR) of the train-set was 0.66, and the test-set AUPR was 0.39.
+Cancer types were the most important predictors.
+Both nsulin Like Growth Factor 1 Receptor (IGFR1R) and the phosphatase and tensin homolog (PTEN) were the most significant genes in gene expression predictors.
+Our study suggests that predicting genomic alterations using gene expression data is possible, with good outcomes.
 
@@
 
-Targeted therapy becomes a standard treatment in many cancer patients.
-The majority of targeted therapy requires a test for a specific cancer genomic alteration, to treat the patients.
-Many direct tests for the genomic alteration have been developed and prooved for their clinical utility to find which patients will be responding to the targeted therapy.
+Targeted therapy has become a standard treatment for many cancer patients, however the approach requires a test for a specific cancer genomic alteration, to treat patients.
+Several direct genomic alteration tests have been developed  and proven for their clinical utility to treat patients.
+Pyrosequencing is a standard clincial diagnostic test for EGFR mutation detection.
 
-It has been actively researched that the machine learnig approach can be applied to detect genomic alterations.
-Machine learning algorithm can build a prediction model from a large number of predictors such as radiolomic features @dercleIdentificationNonSmall2020, pathology image [@coudrayClassificationMutationPrediction2018] or gene expression data [@wayMachineLearningDetects2018 @loiPIK3CAMutationsAssociated2010].
-Because the most direct genomic test is more specific and sensitive than predictive models, the machine learning approach might have a limited role in clinical practice.
-However, the machine learning prediction can be a second-best when the direct test fails.
+Machine learning approaches can be applied to detect genomic alterations .
+Machine learning algorithms can build prediction models from a large number of predictors, such as radiomic  features @dercleIdentificationNonSmall2020, pathology image [@coudrayClassificationMutationPrediction2018] or gene expression data [@wayMachineLearningDetects2018 @loiPIK3CAMutationsAssociated2010].
+Because most direct genomic tests are more specific and sensitive than predictive models, machine learning approaches may have limited roles in clinical practice, however, machine learning approaches are ideal when direct tests are unavailable or fail.
 
-Prediction of RAS pathway activation with gene expression data was done in the previous study.
-@way_2018_machine_cellreports They trained pan-cancer The Cancer Genome Atlas (TCGA) data with a supervised elastic net penalized logistic regression classifier with stochastic gradient descent.
-The performance of their model was 84% with an area under the receiver operating characteristic (AUROC) curve and 63% with an area under the precision-recall (AUPR) curve.
-The authors suggested their approach can be applied to other genomic alterations.
+RAS pathway activation predictions have been performed using gene expression data @way_2018_machine_cellreports.
+Authors used data from The Cancer Genome Atlas (TCGA), with a supervised elastic net penalized logistic regression classifier, with stochastic gradient descent.
+Their model performance was 84% with an area under the receiver operating characteristic (AUROC) curve, and 63% with an area under the precision-recall (AUPR) curve.
+Importantly, these authors suggested their approach could be applied to other genomic alterations.
 
-Breast cancer with PIK3CA mutation has been approved to use PIK3CA inhibitor in hormone receptor-positive HER2 negative subtype.
-@andre_2019_alpelisib_nengljmeda The PIK3CA mutation is the second most driver mutation after TP53.
-The PIK3CA mutation is most frequently founded in endometrial carcinoma (45%), and followed by breast invasive carcinoma (24%), cervical squamous cell carcinoma, and endocervical adenocarcinoma (20%) and colon adenocarcinoma (16%).
+Breast cancer expressing PIK3CA mutations can be treated using PIK3CA inhibitors, in hormone receptor-positive HER2 negative subtypes  @andre_2019_alpelisib_nengljmeda.
+The PIK3CA mutation is the second most common driver mutation after TP53, and is most frequently detected in endometrial carcinoma (45%), followed by breast invasive carcinoma (24%), cervical squamous cell carcinoma, endo-cervical adenocarcinoma (20%) and colon adenocarcinoma (16%).
 
 PIK3CA encodes the p110$\alpha$ catalytic subunit of phosphatidylinositol 3′-kinase (PI3K).
-PI3K is a protein kinase that phosphorylates phosphatidylinositol 4,5-biphosphate (PIP~2~) to make phosphatidylinositol 3,4,5-triphosphate (PIP~3~).
-Phosphatase and tensin homolog (PTEN) changes PIP~2~ to PIP~3~ in contrast, PI3K.
-PIP~3~ is a second messenger to activate protein kinase B (AKT) which is a serine/threonine-specific protein kinase.
-AKT inhibits apoptosis and promotes cell proliferation.
-@cantleyPhosphoinositide3KinasePathway2002
+PI3K is a protein kinase that phosphorylates phosphatidylinositol 4,5-biphosphate (PIP~2~) to generate phosphatidylinositol 3,4,5-triphosphate (PIP~3~).
+The phosphatase and tensin homolog (PTEN) converts PIP~2~ to PIP~3~ in contrast to PI3K.
+PIP~3~ is a second messenger that activates protein kinase B (AKT), which is a serine/threonine-specific protein kinase.
+AKT inhibits apoptosis and promotes cell proliferation
+ @cantleyPhosphoinositide3KinasePathway2002.
 
-We apply a supervised elastic net penalized logistic regression model in prediction PIK3CA mutation.
-The purpose of this study is to know whether this prediction model approach can be applied not only to RAS pathway activation but also PIK3CA mutation across many cancer types.
+We applied a supervised elastic net penalized logistic regression model to predict PIK3CA mutations.
+We wanted to ascertain whether this prediction model approach could be applied not only to RAS pathway activation, but also to PIK3CA mutation predictions across several cancer types.
 
@@
 We used the TCGA pan-cancer dataset.
-TCGA is a cancer genomic consortium that archives data of exome sequencing, gene expression, DNA methylation, protein expression, and clinical data of more than 10000 cancer samples across 33 common cancer types.
+TCGA archives the following; exome sequencing, gene expression, DNA methylation, protein expression, and clinical data from > 10,000 cancer samples across 33 common cancer types.
 The TCGA dataset is publically available.
-PIK3CA mutation data was got using cgdsr rpackage.[@jacobsenCgdsrRbasedAPI2019] Gene expression data was downloaded from the National Cancer Institute (NCI)'s Genomic Data Commons (GDC) website that archives data used for The Pan-Cancer Atlas initiative.
-(https://gdc.cancer.gov/about-data/publications/pancanatlas) The gene expression TCGA pan-cancer dataset is batch-corrected with normalization.
-
-The target variable was the PIK3CA mutation status.
-The status of  PIK3CA was considered as positive when the case has following PIK3CA variants which are the target variants of the therascreen PIK3CA RGQ PCR Kit; C420R, E542K, E545A, E545D, E545G, E545K, Q546E, Q546R, H1047L, H1047R, H1047Y.
-The therascreen PIK3CA RGQ PCR Kit was approved as a companion diagnostics to treat with PIK3CA inhibitor by the U.S.
-Food and Drug Administration.
-We split the three-quarters of the dataset for the trainset and one quarter for the test set.
+PIK3CA mutation data was extracted using cgdsr rpackage [@jacobsenCgdsrRbasedAPI2019].
+Gene expression data was downloaded from the National Cancer Institute (NCI)’s Genomic Data Commons (GDC) website.
+This archives data for TCGA (https://gdc.cancer.gov/about-data/publications/pancanatlas).
+Gene expression in the TCGA pan-cancer dataset is batch-corrected with normalization.
+The target variable was PIK3CA mutation status.
+PIK3CA status was considered positive when the case had the following PIK3CA variants, which were the target variants of the Therascreen PIK3CA RGQ PCR Kit ; C420R, E542K, E545A, E545D, E545G, E545K, Q546E, Q546R, H1047L, H1047R, H1047Y.
+This kit  was approved as a companion diagnostics test to treat with PIK3CA inhibitor  by the United States Food and Drug Administration.
+We split three-quarters of the dataset into the train-set and one quarter into the test-set.
 
@@
 
-To narrow down potential predictors, genes with a large the median absolute deviation (more than third-quartiles) were selected.
-33 cancer type dummy variables were included in predictor variables.
-Yeo-Johnson transformation was done to correct skewness.
-Centering and scaling were done.
-All preprocessing was done using recipe r package.
-[@kuhnRecipesPreprocessingTools2020] Penalized logistic regression was applied to prediction modeling.
-10-fold cross-validation with targe variable stratification was done over the hyperparameter grid: $\lambda$ {$10^{-5}$, $10^{-4}$,$10^{-3}$,$10^{-2}$,$10^{-1}$, $10^{0}$}, $\alpha$ {0.0, 0.25, 0.5, 0.75}.
-Lambda is  penalty scaling parameter and alpha is mixing parameter of penalty function ($(1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1$).
-[@friedmanRegularizationPathsGeneralized2010]
+To narrow down potential predictors, genes with a large median absolute deviation (> third-quartiles) were selected.
+Thirty three cancer type dummy variables were included in predictor variables.
+Yeo-Johnson transformation was performed to correct skewness.
+Centering and scaling were also performed.
+All preprocessing was performed using the recipe r package [@kuhnRecipesPreprocessingTools2020].
+Penalized logistic regression was applied to prediction modeling.
+Ten-fold cross-validation with targe  variable stratification was performed over the hyper-parameter grid: $\lambda$ {$10^{-5}$, $10^{-4}$,$10^{-3}$,$10^{-2}$,$10^{-1}$, $10^{0}$}, $\alpha$ {0.0, 0.25, 0.5, 0.75}.
+Lambda ($\lambda$) is a penalty scaling parameter and alpha ($\alpha$) is a mixing parameter of penalty function ($(1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1$) [@friedmanRegularizationPathsGeneralized2010].
 
@@
 
-Model performance was evaluated with the AUROC curve and the AUPR curve.
-The AUPR is more informative than AUROC on imbalanced datasets.
-[@saitoPrecisionRecallPlotMore2015] The modeling process and accessing model performance were done with tidymodels rpackage.
-[@kuhnTidymodelsEasilyInstall2020]
+Model performance was evaluated using AUROC and AUPR curve approaches.
+The AUPR approach is more informative than AUROC for imbalanced datasets [@saitoPrecisionRecallPlotMore2015].
+The modeling process and accessing  model performance were performed with the tidymodels rpackage  [@kuhnTidymodelsEasilyInstall2020].
 
@@
 
-10845 cases were available for both PIK3CA mutation and mRNA expression data.
-5128 out of 20502 genes were included in the modeling process after filtering with median absolute deviation as described at the modeling process method.
-The prevalence rate of PIK3CA was 0.11 in all cases.
-The PIK3CA prevalence rate of each cancer type varied.
-The median prevalence rate of PIK3CA of each cancer type was 0.03 (range 0-0.33) (Figure 1).
+10,845 cases were available for both PIK3CA mutation and mRNA expression data.
+5,128 out of 20,502 genes were included in the modeling process, after filtering for median absolute deviation, as described in the modeling process method.
+The prevalence rate for PIK3CA was 0.11 in all cases.
+The PIK3CA prevalence rate in each cancer type varied.
+The median prevalence rate of PIK3CA for each cancer type was 0.03 (range 0–0.33) (Figure 1).
 
@@
 
-In 10-fold cross-validation, the model with $\lambda$ = 0.01 and $\alpha$ = 1.0 (Ridge regression) showed best performance in terms of AUROC.
-The final model was trained with the selected hyperparameters with the entire trainset.
-The trainset AUROC was 0.93 and the testset AUROC was 0.84.
-The AUPR of trainset was 0.66 and the testset AUPR was 0.39.
-(Figure 2A)
+For 10-fold cross-validation, the model with $\lambda$ = 0.01 and $\alpha$ = 1.0 (Ridge regression) showed the best performance in terms of AUROC.
+The final model was trained with the selected hyper-parameters with the entire train-set.
+The train-set AUROC was 0.93 and the test-set AUROC was 0.84.
+The AUPR of the train-set was 0.66 and the test-set AUPR was 0.39 (Figure 2A).
 
@@
 
-Because the prevalence of PIK3CA mutation varies across the cancer type, the performance of each cancer type was investigated.
-The AUROC and AUPR were positively correlated between the train set and test set in cancer type subanalysis.
-(Figure 2B) The AUPR was high in cancer type with high PIK3CA mutation rate such as colon, breast, uterus cancer types.
-The AUROC did not correlate with the PIK3CA mutation rate of each cancer type.
-(Figure 2C)
+Because PIK3CA mutation prevalence varied across cancer types, the performance of each cancer type was investigated.
+The AUROC and AUPR were positively correlated between the train-sets and test-sets in cancer type sub-analysis (Figure 2B).
+The AUPR was high in cancer types with high PIK3CA mutation rates such as colon, breast and uterus cancer types.
+The AUROC did not correlate with PIK3CA mutation rates of each cancer type (Figure 2C).
 
@@
 
-Figure 3 shows the top 30 important predictors.
+The top 30 important predictors are shown (Figure 3).
 The coefficient is the parameter of the predictor which represents the effect of the predictor on prediction.
-IGF1R mRNA expression was the strongest negative predictor and PTEN was the strongest positive predictor.
-Both IGFR1R and PTEN are key players in the tyrosine kinase pathway.
+Insulin Like Growth Factor 1 Receptor (IGF1R)  mRNA expression was the strongest negative predictor, and PTEN was the strongest positive predictor.
+Both IGFR1R  and PTEN are key players in the tyrosine kinase pathway .
 The cancer types were important predictors.
-Some cancer types including uterine carcinosarcoma (UCS), bladder urothelial carcinoma (BLCA), pancreatic adenocarcinoma (PAAD), lymphoid neoplasm diffuse large B-cell lymphoma (DLBC) are the strongest predictors.
+Some cancer types including uterine carcinosarcoma (UCS), bladder urothelial carcinoma (BLCA), pancreatic adenocarcinoma (PAAD), lymphoid neoplasm diffuse large B-cell lymphoma (DLBC) were the strongest predictors.
 
@@
 
-Our model showed good performance to predict the PIK3CA mutation of various cancer types.
-This result suggests that the supervised elastic net penalized logistic regression model can be applied not only to the RAS activation pathway but also other genomic alterations.
-Both the RAS activation pathway and PIK3CA mutation are important and common cancer genomic alterations.
-They have a significant effect on gene expression in cancer cells.
-It can not be generalized that the supervised elastic net penalized logistic regression model can be applied to other genomic alteration.
+Our model showed good performance in predicting PIK3CA mutations in various cancer types.
+Our data suggested that the supervised elastic net penalized logistic regression model could be applied not only to the RAS activation pathway, but also to other genomic alterations.
+Both the RAS activation pathway and PIK3CA mutations are key, common cancer genomic alterations.
+When mutated, they exert significant effect on gene expression in cells.
+However, the supervised elastic net penalized logistic regression model cannot be generalized or applied to other genomic alterations.
 It might be a challenging prediction of genomic alterations that are infrequent or have a weak effect on gene expression.
 
-Prediction modeling from the TCGA pan-cancer dataset has limitations regarding data preprocessing.
-The gene expression data is processed between-sample normalization to remove the batch effect.
-If the model has been trained from the between-sample normalization, a new sample can not be exactly preprocessed with normalization which was done on trainset.
-A model based on gene expression TCGA pan-cancer dataset has a limitation on data preprocessing.
-It is necessary for developing a preprocessing method that is independent with a dataset to apply the gene expression data to the prediction model.
+Prediction modeling from the TCGA pan-cancer dataset can be limiting in terms of data preprocessing.
+The gene expression data is processed by between-sample normalization to remove batch effects.
+If the model has been trained from between-sample normalization, a new sample cannot be exactly preprocessed with normalization which was done on trainset.
+A model based on gene expression from the TCGA pan-cancer dataset is limited in terms of data preprocessing.
+It is necessary to develop a preprocessing method that is independent of a dataset, to apply gene expression data to the prediction model.
 
-Our PIK3CA prediction model performed similar with the RAS activation prediction model of the previous study in terms of AUROC (0.84).
-However AUPR of our model for PIK3CA is lower than model for RAS activation (0.39 vs 0.63).
-The reason of lower AUPR of our model can be explained by low prevalence rate of PIK3CA mutation and imbalanced dataset.
+Our PIK3CA prediction model was similar to the RAS activation prediction model in terms of AUROC (0.84).
+However the AUPR of our model was lower than the RAS activation model (0.39 versus 0.63).
+The reason for our lower AUPR may be explained by the low prevalence rate of PIK3CA mutations, and an imbalanced dataset.
 The model for RAS activation trained with cancer types with more than 0.05 prevalence of RAS activation to avoid imbalance classification problem.
-We included all cancer types into modeling process.
-The lower prevalance rate of target variable mean our dataset has a lower baseline of AUPR.
-In the subanalysis for performance of each cancer types, the cancer types with higher PIK3CA mutation rate showe better AUPR.
+We included all cancer types in our modeling process.
+The lower prevalence rate of target variables meant our dataset had a lower AUPR baseline.
+In the sub-analysis performance of each cancer type, the cancer types with higher PIK3CA mutation rates showed better AUPRs.
 
-Our model includes cancer type as predictors and they are stronger predictors than gene expression.
-The varying prevalence of PIC3CA mutation across cancer type might reason for the strong predictive power of cancer type.
-If the cancer type was wrong or can not be determined, our model performance can be poor.
+Our model included cancer types as predictors, and they were stronger predictors than gene expression.
+The varying prevalence of PIC3CA mutations across cancer types may be a reason for the strong predictive power of cancer types.
+If the cancer type was wrong or could not be determined, our model performance was poor.
 
 Some significant gene expression predictors were closely related to PTEN and the PI3K pathway.
-PTEN and IGFR1R are the strongest gene expression predictor which have negative and positive predictive power.
-IGF1R is a tyrosine kinase receptor that activates PI3K.
-[@leroithInsulinlikeGrowthFactor2003] PTEN is an important regulator of PIP~3~ by dephosphorylating  PIP~3~ in contrast PI3K.[@cantleyPhosphoinositide3KinasePathway2002]
+PTEN and IGFR1R were the strongest gene expression predictors, which has negative and positive predictive powers.
+IGF1R is a tyrosine kinase receptor that activates PI3K  [@leroithInsulinlikeGrowthFactor2003].
+], and PTEN is an important regulator of PIP~3~ by dephosphorylating PIP~3~, in contrast to PI3K [@cantleyPhosphoinositide3KinasePathway2002].
 
-There is some study trying to predict genomic alterations from gene expression data.
-A study performed PIK3CA mutation prediction by a gene-expression signature which is a sum of the average of the logarithmic gene expression.
-The model showed good performance AUROC 0.71 in an independent test set.
-[@cizkovaGeneExpressionProfiling2010; @loiPIK3CAMutationsAssociated2010] Another study predicted copy number alterations with gene expression using a multinomial logistic regression model with least absolute shrinkage and selection operator (LASSO).
-The prediction of the 1p/19q codel was very good with an AUROC of 0.997.
-The gene-level prediction was good with an AUROC 0.75.
-[@muCNAPEMachineLearning2019] A Hidden Markov Model Approach for Prediction of Genomic Alterations from Gene Expression Profiling.[@gengHiddenMarkovModel2008] A logistic regression model was used for MYCN gene amplification in neuroblastoma.
-[@heGeneSignaturesAssociated2020]
+Several studies have attempted to predict genomic alterations from gene expression data [@cizkovaGeneExpressionProfiling2010; @loiPIK3CAMutationsAssociated2010].
+A study investigated PIK3CA mutation predictions using gene-expression signatures which is a sum of the average of the logarithmic gene expression .
+The model showed good performance AUROC 0.71 in an independent test set  [@cizkovaGeneExpressionProfiling2010; @loiPIK3CAMutationsAssociated2010].Another study predicted copy number alterations with gene expression, using a multinomial logistic regression model with least absolute shrinkage and selection operator (LASSO) parameters [@muCNAPEMachineLearning2019].
+The prediction of the 1p/19q codel  was very good, with an AUROC of 0.997, and gene-level predictions were good, with an AUROC of 0.75 [@muCNAPEMachineLearning2019].
+A Hidden Markov Model Approach for Prediction of Genomic Alterations from Gene Expression Profiling [@gengHiddenMarkovModel2008].
+A logistic regression model was used for MYCN  gene amplification in neuroblastoma [@heGeneSignaturesAssociated2020].
 
-Our study suggests that the prediction of genomic alteration with gene expression data is possible with good performance.
-However, improvement of performance is required for applying to clinical test and the standardization of generation processing of gene expression data is also needed.
+Our study suggested that the prediction of genomic alterations using gene expression data was possible, with good performance.
+However, improved performances are required for clinical tests, and the standardization of generation processing of gene expression data is also needed .
 
@@
 
-- Figure 1. Prevalence rate of PIK3CA across cancer types.
-The abbreviations of cancer types are explained in the S1 Appendix.
+- Figure 1. Prevalence rate of PIK3CA mutations across cancer types. Cancer type abbreviations are explained in the S1 Appendix.
 
-- Figure 2. Summary of modeling results
+- Figure 2.
+Summary of modeling results.
+(A) Left: receiver operating characteristic (ROC) curve.
+Right: precision-recall (PR) curve of train-set and test-set.
+The horizontal green line is the PIK3CA mutation rate (0.11) (B) Correlation between train-set and test-set of the area under the receiver operating characteristic curve (AUROC), and the area under the precision-recall curve (AUPR) among cancer types.
+The gray band is the 95% confidence interval.
+Abbreviations are explained in the S1 Appendix.
+(C) Correlations between the PIK3CA mutation rate of the AUROC, and the AUPR.
 
-(A) Left: receiver operating characteristic (ROC) curve.
-right: precision-recall (PR) curve of trainset and test set.
-The horizontal green line is the PIK3CA mutation rate (0.11)  (B) Correlation between trainset and test set of the area under the receiver operating characteristic curve (AUROC) and the area under the precision-recall curve (AUPR) among cancer type.
-The gray band is 95% confidence interval.
-The abbreviations are explained in the S1 Appendix.
-(C) Correlation between the PIK3CA mutation rate of the AUROC and the AUPR.
-
-- Figure 3. Coefficients of model
-
-(A) Top 30 high coefficients of mRNA. (B) Coefficients of cancer types. The abbreviations of cancer types are explained in the S1 Appendix.
+- Figure 3. Coefficient model. (A) Top 30 high mRNA coefficients. (B) Cancer type coefficients. Cancer types abbreviations are explained in the S1 Appendix.
 
@@
 
-# References {#references .unnumbered}
+# References {#references}

### 2020-06-11 plosOne/plosOne.tex
PRS

@@
 \section*{Abstract}
-Breast cancer with PIK3CA mutation has been approved to use PIK3CA
-inhibitor in hormone receptor-positive HER2 negative subtype. We apply a
-supervised elastic net penalized logistic regression model in prediction
-PIK3CA mutation from gene expression data. Penalized logistic regression
-was applied to prediction modeling using the TCGA pan-cancer dataset.
-10845 cases were available for both PIK3CA mutation and mRNA expression
-data. In 10-fold cross-validation, the model with \(\lambda\) = 0.01 and
-\(\alpha\) = 1.0 (Ridge regression) showed best performance in terms of
-AUROC. The final model was trained with the selected hyperparameters
-with the entire trainset. The trainset AUROC was 0.93 and the testset
-AUROC was 0.84. The AUPR of trainset was 0.66 and the testset AUPR was
-0.39. The cancer types were the most important predictors. Both IGFR1R
-and PTEN were most important among gene expression predictors. Our study
-suggests that the prediction of genomic alteration with gene expression
-data is possible with good performance.
+Breast cancers with PIK3CA mutations can be treated with PIK3CA
+inhibitors in hormone receptor-positive HER2 negative subtypes. We
+applied a supervised elastic net penalized logistic regression model to
+predict PIK3CA mutations from gene expression data. This regression
+approach was applied to predict modeling using the TCGA pan-cancer
+dataset. Approximately 10,845 cases were available for PIK3CA mutation
+and mRNA expression data. In 10-fold cross-validation, the model with
+\(\lambda\) = 0.01 and \(\alpha\) = 1.0 (Ridge regression) showed the
+best performance, in terms of area under the receiver operating
+characteristic (AUROC). The final model was developed with selected
+hyper-parameters using the entire train-set. The train-set AUROC was
+0.93, and the test-set AUROC was 0.84. The area under the
+precision-recall (AUPR) of the train-set was 0.66, and the test-set AUPR
+was 0.39. Cancer types were the most important predictors. Both IGFR1R
+and the phosphatase and tensin homolog (PTEN) were the most significant
+genes in gene expression predictors. Our study suggests that predicting
+genomic alterations using gene expression data is possible, with good
+outcomes.
 
@@
 
-Targeted therapy becomes a standard treatment in many cancer patients.
-The majority of targeted therapy requires a test for a specific cancer
-genomic alteration, to treat the patients. Many direct tests for the
-genomic alteration have been developed and prooved for their clinical
-utility to find which patients will be responding to the targeted
-therapy.
-
-It has been actively researched that the machine learnig approach can be
-applied to detect genomic alterations. Machine learning algorithm can
-build a prediction model from a large number of predictors such as
-radiolomic features {[}1{]}, pathology image {[}2{]} or gene expression
-data {[}3{]}. Because the most direct genomic test is more specific and
-sensitive than predictive models, the machine learning approach might
-have a limited role in clinical practice. However, the machine learning
-prediction can be a second-best when the direct test fails.
-
-Prediction of RAS pathway activation with gene expression data was done
-in the previous study. {[}5{]} They trained pan-cancer The Cancer Genome
-Atlas (TCGA) data with a supervised elastic net penalized logistic
-regression classifier with stochastic gradient descent. The performance
-of their model was 84\% with an area under the receiver operating
-characteristic (AUROC) curve and 63\% with an area under the
-precision-recall (AUPR) curve. The authors suggested their approach can
-be applied to other genomic alterations.
-
-Breast cancer with PIK3CA mutation has been approved to use PIK3CA
-inhibitor in hormone receptor-positive HER2 negative subtype. {[}6{]}
-The PIK3CA mutation is the second most driver mutation after TP53. The
-PIK3CA mutation is most frequently founded in endometrial carcinoma
-(45\%), and followed by breast invasive carcinoma (24\%), cervical
-squamous cell carcinoma, and endocervical adenocarcinoma (20\%) and
-colon adenocarcinoma (16\%).
-
-PIK3CA encodes the p110\(\alpha\) catalytic subunit of
+Targeted therapy has become a standard treatment for many cancer
+patients, however the approach requires a test for a specific cancer
+genomic alteration, to treat patients. Several direct genomic alteration
+tests have been developed and proven for their clinical utility to treat
+patients. Machine learning approaches can be applied to detect genomic
+alterations . Machine learning algorithms can build prediction models
+from a large number of predictors, such as radiomic features {[}1{]},
+pathology image {[}2{]} or gene expression data {[}3{]}. Because most
+direct genomic tests are more specific and sensitive than predictive
+models, machine learning approaches may have limited roles in clinical
+practice, however, machine learning approaches are ideal when direct
+tests are unavailable or fail. RAS pathway activation predictions have
+been performed using gene expression data {[}5{]}. Authors used data
+from The Cancer Genome Atlas (TCGA), with a supervised elastic net
+penalized logistic regression classifier, with stochastic gradient
+descent. Their model performance was 84\% with an area under the
+receiver operating characteristic (AUROC) curve, and 63\% with an area
+under the precision-recall (AUPR) curve. Importantly, these authors
+suggested their approach could be applied to other genomic alterations.
+Breast cancer expressing PIK3CA mutations can be treated using PIK3CA
+inhibitors, in hormone receptor-positive HER2 negative subtypes {[}6{]}.
+The PIK3CA mutation is the second most common driver mutation after
+TP53, and is most frequently detected in endometrial carcinoma (45\%),
+followed by breast invasive carcinoma (24\%), cervical squamous cell
+carcinoma, endo-cervical adenocarcinoma (20\%) and colon adenocarcinoma
+(16\%). PIK3CA encodes the p110\(\alpha\) catalytic subunit of
 phosphatidylinositol 3′-kinase (PI3K). PI3K is a protein kinase that
 phosphorylates phosphatidylinositol 4,5-biphosphate
-(PIP\textsubscript{2}) to make phosphatidylinositol 3,4,5-triphosphate
-(PIP\textsubscript{3}). Phosphatase and tensin homolog (PTEN) changes
-PIP\textsubscript{2} to PIP\textsubscript{3} in contrast, PI3K.
-PIP\textsubscript{3} is a second messenger to activate protein kinase B
-(AKT) which is a serine/threonine-specific protein kinase. AKT inhibits
-apoptosis and promotes cell proliferation. {[}7{]}
+(PIP\textsubscript{2}) to generate phosphatidylinositol
+3,4,5-triphosphate (PIP\textsubscript{3}). The phosphatase and tensin
+homolog (PTEN) converts PIP\textsubscript{2} to PIP\textsubscript{3} in
+contrast to PI3K. PIP\textsubscript{3} is a second messenger that
+activates protein kinase B (AKT), which is a serine/threonine-specific
+protein kinase. AKT inhibits apoptosis and promotes cell proliferation
+{[}7{]}.
 
-We apply a supervised elastic net penalized logistic regression model in
-prediction PIK3CA mutation. The purpose of this study is to know whether
-this prediction model approach can be applied not only to RAS pathway
-activation but also PIK3CA mutation across many cancer types.
+We applied a supervised elastic net penalized logistic regression model
+to predict PIK3CA mutations. We wanted to ascertain whether this
+prediction model approach could be applied not only to RAS pathway
+activation, but also to PIK3CA mutation predictions across several
+cancer types.
 
@@
 
-We used the TCGA pan-cancer dataset. TCGA is a cancer genomic consortium
-that archives data of exome sequencing, gene expression, DNA
-methylation, protein expression, and clinical data of more than 10000
-cancer samples across 33 common cancer types. The TCGA dataset is
-publically available. PIK3CA mutation data was got using cgdsr
-rpackage.{[}8{]} Gene expression data was downloaded from the National
-Cancer Institute (NCI)'s Genomic Data Commons (GDC) website that
-archives data used for The Pan-Cancer Atlas initiative.
-(https://gdc.cancer.gov/about-data/publications/pancanatlas) The gene
-expression TCGA pan-cancer dataset is batch-corrected with
-normalization.
-
-The target variable was the PIK3CA mutation status. The status of PIK3CA
-was considered as positive when the case has following PIK3CA variants
-which are the target variants of the therascreen PIK3CA RGQ PCR Kit;
-C420R, E542K, E545A, E545D, E545G, E545K, Q546E, Q546R, H1047L, H1047R,
-H1047Y. The therascreen PIK3CA RGQ PCR Kit was approved as a companion
-diagnostics to treat with PIK3CA inhibitor by the U.S. Food and Drug
-Administration. We split the three-quarters of the dataset for the
-trainset and one quarter for the test set.
+We used the TCGA pan-cancer dataset. TCGA archives the following; exome
+sequencing, gene expression, DNA methylation, protein expression, and
+clinical data from \textgreater{} 10,000 cancer samples across 33 common
+cancer types. The TCGA dataset is publically available. PIK3CA mutation
+data was extracted using cgdsr rpackage {[}8{]}. Gene expression data
+was downloaded from the National Cancer Institute (NCI)'s Genomic Data
+Commons (GDC) website. This archives data for TCGA
+(https://gdc.cancer.gov/about-data/publications/pancanatlas). Gene
+expression in the TCGA pan-cancer dataset is batch-corrected with
+normalization. The target variable was PIK3CA mutation status. PIK3CA
+status was considered positive when the case had the following PIK3CA
+variants, which were the target variants of the Therascreen PIK3CA RGQ
+PCR Kit ; C420R, E542K, E545A, E545D, E545G, E545K, Q546E, Q546R,
+H1047L, H1047R, H1047Y. This kit was approved as a companion diagnostics
+test to treat with PIK3CA inhibitor by the United States Food and Drug
+Administration. We split three-quarters of the dataset into the
+train-set and one quarter into the test-set.
 
@@
 
-To narrow down potential predictors, genes with a large the median
-absolute deviation (more than third-quartiles) were selected. 33 cancer
-type dummy variables were included in predictor variables. Yeo-Johnson
-transformation was done to correct skewness. Centering and scaling were
-done. All preprocessing was done using recipe r package. {[}9{]}
-Penalized logistic regression was applied to prediction modeling.
-10-fold cross-validation with targe variable stratification was done
-over the hyperparameter grid: \(\lambda\) \{\(10^{-5}\),
-\(10^{-4}\),\(10^{-3}\),\(10^{-2}\),\(10^{-1}\), \(10^{0}\)\},
-\(\alpha\) \{0.0, 0.25, 0.5, 0.75\}. Lambda is penalty scaling parameter
-and alpha is mixing parameter of penalty function
-(\((1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1\)).
-{[}10{]}
+To narrow down potential predictors, genes with a large median absolute
+deviation (\textgreater{} third-quartiles) were selected. Thirty three
+cancer type dummy variables were included in predictor variables.
+Yeo-Johnson transformation was performed to correct skewness. Centering
+and scaling were also performed. All preprocessing was performed using
+the recipe r package {[}9{]}. Penalized logistic regression was applied
+to prediction modeling. Ten-fold cross-validation with targe variable
+stratification was performed over the hyper-parameter grid: \(\lambda\)
+\{\(10^{-5}\), \(10^{-4}\),\(10^{-3}\),\(10^{-2}\),\(10^{-1}\),
+\(10^{0}\)\}, \(\alpha\) \{0.0, 0.25, 0.5, 0.75\}. Lambda (\(\lambda\))
+is a penalty scaling parameter and alpha (\(\alpha\)) is a mixing
+parameter of penalty function
+(\((1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1\))
+{[}10{]}.
 
@@
 
-Model performance was evaluated with the AUROC curve and the AUPR curve.
-The AUPR is more informative than AUROC on imbalanced datasets. {[}11{]}
-The modeling process and accessing model performance were done with
-tidymodels rpackage. {[}12{]}
+Model performance was evaluated using AUROC and AUPR curve approaches.
+The AUPR approach is more informative than AUROC for imbalanced datasets
+{[}11{]}. The modeling process and accessing model performance were
+performed with the tidymodels rpackage {[}12{]}.
 
@@
 
-10845 cases were available for both PIK3CA mutation and mRNA expression
-data. 5128 out of 20502 genes were included in the modeling process
-after filtering with median absolute deviation as described at the
-modeling process method. The prevalence rate of PIK3CA was 0.11 in all
-cases. The PIK3CA prevalence rate of each cancer type varied. The median
-prevalence rate of PIK3CA of each cancer type was 0.03 (range 0-0.33)
+10,845 cases were available for both PIK3CA mutation and mRNA expression
+data. 5,128 out of 20,502 genes were included in the modeling process,
+after filtering for median absolute deviation, as described in the
+modeling process method. The prevalence rate for PIK3CA was 0.11 in all
+cases. The PIK3CA prevalence rate in each cancer type varied. The median
+prevalence rate of PIK3CA for each cancer type was 0.03 (range 0--0.33)
 (Figure 1).
@@
 
-In 10-fold cross-validation, the model with \(\lambda\) = 0.01 and
-\(\alpha\) = 1.0 (Ridge regression) showed best performance in terms of
-AUROC. The final model was trained with the selected hyperparameters
-with the entire trainset. The trainset AUROC was 0.93 and the testset
-AUROC was 0.84. The AUPR of trainset was 0.66 and the testset AUPR was
-0.39. (Figure 2A)
+For 10-fold cross-validation, the model with \(\lambda\) = 0.01 and
+\(\alpha\) = 1.0 (Ridge regression) showed the best performance in terms
+of AUROC. The final model was trained with the selected hyper-parameters
+with the entire train-set. The train-set AUROC was 0.93 and the test-set
+AUROC was 0.84. The AUPR of the train-set was 0.66 and the test-set AUPR
+was 0.39 (Figure 2A).
 
@@
 
-Because the prevalence of PIK3CA mutation varies across the cancer type,
-the performance of each cancer type was investigated. The AUROC and AUPR
-were positively correlated between the train set and test set in cancer
-type subanalysis. (Figure 2B) The AUPR was high in cancer type with high
-PIK3CA mutation rate such as colon, breast, uterus cancer types. The
-AUROC did not correlate with the PIK3CA mutation rate of each cancer
-type. (Figure 2C)
+Because PIK3CA mutation prevalence varied across cancer types, the
+performance of each cancer type was investigated. The AUROC and AUPR
+were positively correlated between the train-sets and test-sets in
+cancer type sub-analysis (Figure 2B). The AUPR was high in cancer types
+with high PIK3CA mutation rates such as colon, breast and uterus cancer
+types. The AUROC did not correlate with PIK3CA mutation rates of each
+cancer type (Figure 2C).
 
@@
 
-Figure 3 shows the top 30 important predictors. The coefficient is the
-parameter of the predictor which represents the effect of the predictor
-on prediction. IGF1R mRNA expression was the strongest negative
-predictor and PTEN was the strongest positive predictor. Both IGFR1R and
-PTEN are key players in the tyrosine kinase pathway. The cancer types
-were important predictors. Some cancer types including uterine
-carcinosarcoma (UCS), bladder urothelial carcinoma (BLCA), pancreatic
-adenocarcinoma (PAAD), lymphoid neoplasm diffuse large B-cell lymphoma
-(DLBC) are the strongest predictors.
+The top 30 important predictors are shown (Figure 3). The coefficient is
+the parameter of the predictor which represents the effect of the
+predictor on prediction . IGF1R mRNA expression was the strongest
+negative predictor, and PTEN was the strongest positive predictor. Both
+IGFR1R and PTEN are key players in the tyrosine kinase pathway . The
+cancer types were important predictors. Some cancer types including
+uterine carcinosarcoma (UCS), bladder urothelial carcinoma (BLCA),
+pancreatic adenocarcinoma (PAAD), lymphoid neoplasm diffuse large B-cell
+lymphoma (DLBC) were the strongest predictors.
 
@@
 
-Our model showed good performance to predict the PIK3CA mutation of
-various cancer types. This result suggests that the supervised elastic
-net penalized logistic regression model can be applied not only to the
-RAS activation pathway but also other genomic alterations. Both the RAS
-activation pathway and PIK3CA mutation are important and common cancer
-genomic alterations. They have a significant effect on gene expression
-in cancer cells. It can not be generalized that the supervised elastic
-net penalized logistic regression model can be applied to other genomic
-alteration. It might be a challenging prediction of genomic alterations
-that are infrequent or have a weak effect on gene expression.
+Our model showed good performance in predicting PIK3CA mutations in
+various cancer types. Our data suggested that the supervised elastic net
+penalized logistic regression model could be applied not only to the RAS
+activation pathway, but also to other genomic alterations. Both the RAS
+activation pathway and PIK3CA mutations are key, common cancer genomic
+alterations. When mutated, they exert significant effect on gene
+expression in cells. However, the supervised elastic net penalized
+logistic regression model cannot be generalized or applied to other
+genomic alterations. It might be a challenging prediction of genomic
+alterations that are infrequent or have a weak effect on gene
+expression.
 
-Prediction modeling from the TCGA pan-cancer dataset has limitations
-regarding data preprocessing. The gene expression data is processed
-between-sample normalization to remove the batch effect. If the model
-has been trained from the between-sample normalization, a new sample can
-not be exactly preprocessed with normalization which was done on
-trainset. A model based on gene expression TCGA pan-cancer dataset has a
-limitation on data preprocessing. It is necessary for developing a
-preprocessing method that is independent with a dataset to apply the
-gene expression data to the prediction model.
+Prediction modeling from the TCGA pan-cancer dataset can be limiting in
+terms of data preprocessing. The gene expression data is processed by
+between-sample normalization to remove batch effects. If the model has
+been trained from between-sample normalization, a new sample cannot be
+exactly preprocessed with normalization which was done on trainset. A
+model based on gene expression from the TCGA pan-cancer dataset is
+limited in terms of data preprocessing. It is necessary to develop a
+preprocessing method that is independent of a dataset, to apply gene
+expression data to the prediction model.
 
-Our PIK3CA prediction model performed similar with the RAS activation
-prediction model of the previous study in terms of AUROC (0.84). However
-AUPR of our model for PIK3CA is lower than model for RAS activation
-(0.39 vs 0.63). The reason of lower AUPR of our model can be explained
-by low prevalence rate of PIK3CA mutation and imbalanced dataset. The
-model for RAS activation trained with cancer types with more than 0.05
-prevalence of RAS activation to avoid imbalance classification problem.
-We included all cancer types into modeling process. The lower prevalance
-rate of target variable mean our dataset has a lower baseline of AUPR.
-In the subanalysis for performance of each cancer types, the cancer
-types with higher PIK3CA mutation rate showe better AUPR.
+Our PIK3CA prediction model was similar to the RAS activation prediction
+model in terms of AUROC (0.84). However the AUPR of our model was lower
+than the RAS activation model (0.39 versus 0.63). The reason for our
+lower AUPR may be explained by the low prevalence rate of PIK3CA
+mutations, and an imbalanced dataset. The model for RAS activation
+trained with cancer types with more than 0.05 prevalence of RAS
+activation to avoid imbalance classification problem. We included all
+cancer types in our modeling process. The lower prevalence rate of
+target variables meant our dataset had a lower AUPR baseline. In the
+sub-analysis performance of each cancer type, the cancer types with
+higher PIK3CA mutation rates showed better AUPRs.
 
-Our model includes cancer type as predictors and they are stronger
+Our model included cancer types as predictors, and they were stronger
 predictors than gene expression. The varying prevalence of PIC3CA
-mutation across cancer type might reason for the strong predictive power
-of cancer type. If the cancer type was wrong or can not be determined,
-our model performance can be poor.
+mutations across cancer types may be a reason for the strong predictive
+power of cancer types. If the cancer type was wrong or could not be
+determined, our model performance was poor.
 
 Some significant gene expression predictors were closely related to PTEN
-and the PI3K pathway. PTEN and IGFR1R are the strongest gene expression
-predictor which have negative and positive predictive power. IGF1R is a
-tyrosine kinase receptor that activates PI3K. {[}13{]} PTEN is an
-important regulator of PIP\textsubscript{3} by dephosphorylating
-PIP\textsubscript{3} in contrast PI3K.{[}7{]}
+and the PI3K pathway. PTEN and IGFR1R were the strongest gene expression
+predictors, which has negative and positive predictive powers. IGF1R is
+a tyrosine kinase receptor that activates PI3K {[}13{]}. {]}, and PTEN
+is an important regulator of PIP\textsubscript{3} by dephosphorylating
+PIP\textsubscript{3}, in contrast to PI3K {[}7{]}.
 
-There is some study trying to predict genomic alterations from gene
-expression data. A study performed PIK3CA mutation prediction by a
-gene-expression signature which is a sum of the average of the
-logarithmic gene expression. The model showed good performance AUROC
-0.71 in an independent test set. {[}4,14{]} Another study predicted copy
-number alterations with gene expression using a multinomial logistic
-regression model with least absolute shrinkage and selection operator
-(LASSO). The prediction of the 1p/19q codel was very good with an AUROC
-of 0.997. The gene-level prediction was good with an AUROC 0.75.
-{[}15{]} A Hidden Markov Model Approach for Prediction of Genomic
-Alterations from Gene Expression Profiling.{[}16{]} A logistic
-regression model was used for MYCN gene amplification in neuroblastoma.
-{[}17{]}
+Several studies have attempted to predict genomic alterations from gene
+expression data {[}4,14{]}. A study investigated PIK3CA mutation
+predictions using gene-expression signatures which is a sum of the
+average of the logarithmic gene expression . The model showed good
+performance AUROC 0.71 in an independent test set {[}4,14{]}.Another
+study predicted copy number alterations with gene expression, using a
+multinomial logistic regression model with least absolute shrinkage and
+selection operator (LASSO) parameters {[}15{]}. The prediction of the
+1p/19q codel was very good, with an AUROC of 0.997, and gene-level
+predictions were good, with an AUROC of 0.75 {[}15{]}. A Hidden Markov
+Model Approach for Prediction of Genomic Alterations from Gene
+Expression Profiling {[}16{]}. A logistic regression model was used for
+MYCN gene amplification in neuroblastoma {[}17{]}.
 
-Our study suggests that the prediction of genomic alteration with gene
-expression data is possible with good performance. However, improvement
-of performance is required for applying to clinical test and the
+Our study suggested that the prediction of genomic alterations using
+gene expression data was possible, with good performance. However,
+improved performances are required for clinical tests, and the
 standardization of generation processing of gene expression data is also
-needed.
+needed .
 
@@
 \item
-  Figure 1. Prevalence rate of PIK3CA across cancer types. The
-  abbreviations of cancer types are explained in the S1 Appendix.
+  Figure 1. Prevalence rate of PIK3CA mutations across cancer types.
+  Cancer type abbreviations are explained in the S1 Appendix.
 \item
-  Figure 2. Summary of modeling results
+  Figure 2. Summary of modeling results. (A) Left: receiver operating
+  characteristic (ROC) curve. Right: precision-recall (PR) curve of
+  train-set and test-set. The horizontal green line is the PIK3CA
+  mutation rate (0.11) (B) Correlation between train-set and test-set of
+  the area under the receiver operating characteristic curve (AUROC),
+  and the area under the precision-recall curve (AUPR) among cancer
+  types. The gray band is the 95\% confidence interval. Abbreviations
+  are explained in the S1 Appendix. (C) Correlations between the PIK3CA
+  mutation rate of the AUROC, and the AUPR.
+\item
+  Figure 3. Coefficient model. (A) Top 30 high mRNA coefficients. (B)
+  Cancer type coefficients. Cancer types abbreviations are explained in
+  the S1 Appendix.
 \end{itemize}
-
-\begin{enumerate}
-\def\labelenumi{(\Alph{enumi})}
-\tightlist
-\item
-  Left: receiver operating characteristic (ROC) curve. right:
-  precision-recall (PR) curve of trainset and test set. The horizontal
-  green line is the PIK3CA mutation rate (0.11) (B) Correlation between
-  trainset and test set of the area under the receiver operating
-  characteristic curve (AUROC) and the area under the precision-recall
-  curve (AUPR) among cancer type. The gray band is 95\% confidence
-  interval. The abbreviations are explained in the S1 Appendix. (C)
-  Correlation between the PIK3CA mutation rate of the AUROC and the
-  AUPR.
-\end{enumerate}
-
-\begin{itemize}
-\tightlist
-\item
-  Figure 3. Coefficients of model
-\end{itemize}
-
-\begin{enumerate}
-\def\labelenumi{(\Alph{enumi})}
-\tightlist
-\item
-  Top 30 high coefficients of mRNA. (B) Coefficients of cancer types.
-  The abbreviations of cancer types are explained in the S1 Appendix.
-\end{enumerate}
 

### 2020-06-12 plosOne/plosOne.Rmd
itelic

@@
 ---
-title: Prediction of PIK3CA mutations from cancer gene expression data
-short-title: PIK3CA mutations prediction  
+title: Prediction of *PIK3CA* mutations from cancer gene expression data
+short-title: *PIK3CA* mutations prediction  
 author:
@@
 abstract: | 
-  Breast cancers with PIK3CA mutations can be treated with PIK3CA inhibitors in hormone receptor-positive HER2 negative subtypes.
-We applied a supervised elastic net penalized logistic regression model to predict PIK3CA mutations from gene expression data.
+  Breast cancers with *PIK3CA* mutations can be treated with *PIK3CA* inhibitors in hormone receptor-positive HER2 negative subtypes.
+We applied a supervised elastic net penalized logistic regression model to predict *PIK3CA* mutations from gene expression data.
 This regression approach was applied to predict modeling using the TCGA pan-cancer dataset.
-Approximately 10,845 cases were available for PIK3CA mutation and mRNA expression data.
+Approximately 10,000 cases were available for *PIK3CA* mutation and mRNA expression data.
 In 10-fold cross-validation, the model with $\lambda$ = 0.01 and $\alpha$ = 1.0 (Ridge regression) showed the best performance, in terms of area under the receiver operating characteristic (AUROC).
-The final model was developed with selected hyper-parameters using the entire train-set.
-The train-set AUROC was 0.93, and the test-set AUROC was 0.84.
-The area under the precision-recall (AUPR) of the train-set was 0.66, and the test-set AUPR was 0.39.
+The final model was developed with selected hyper-parameters using the entire training set.
+The training set AUROC was 0.93, and the test set AUROC was 0.84.
+The area under the precision-recall (AUPR) of the training set was 0.66, and the test set AUPR was 0.39.
 Cancer types were the most important predictors.
-Both nsulin Like Growth Factor 1 Receptor (IGFR1R) and the phosphatase and tensin homolog (PTEN) were the most significant genes in gene expression predictors.
+Both *insulin like growth factor 1 Receptor* (*IGF1R*) and the *phosphatase and tensin homolog* (*PTEN*) were the most significant genes in gene expression predictors.
 Our study suggests that predicting genomic alterations using gene expression data is possible, with good outcomes.
 
-bibliography: [mybibfile.bib, pik3ca.bib, rpackage.bib, genes.bib] 
+bibliography: [mybibfile.bib, PIK3CA.bib, rpackage.bib, genes.bib, PRS.bib] 
 indent: true
@@
 Targeted therapy has become a standard treatment for many cancer patients, however the approach requires a test for a specific cancer genomic alteration, to treat patients.
-Several direct genomic alteration tests have been developed  and proven for their clinical utility to treat patients.
-Pyrosequencing is a standard clincial diagnostic test for EGFR mutation detection.
+Several direct genomic alteration tests have been developed  and proven for their clinical utility to treat patients [@healthNucleicAcidBased2020, @sahnanePyrosequencingEGFRMutation2013].
 
 Machine learning approaches can be applied to detect genomic alterations .
-Machine learning algorithms can build prediction models from a large number of predictors, such as radiomic  features @dercleIdentificationNonSmall2020, pathology image [@coudrayClassificationMutationPrediction2018] or gene expression data [@wayMachineLearningDetects2018 @loiPIK3CAMutationsAssociated2010].
+Machine learning algorithms can build prediction models from a large number of predictors, such as radiomic  features @dercleIdentificationNonSmall2020, pathology image [@coudrayClassificationMutationPrediction2018] or gene expression data [@wayMachineLearningDetects2018 @loi*PIK3CA*MutationsAssociated2010].
 Because most direct genomic tests are more specific and sensitive than predictive models, machine learning approaches may have limited roles in clinical practice, however, machine learning approaches are ideal when direct tests are unavailable or fail.
@@
 
-Breast cancer expressing PIK3CA mutations can be treated using PIK3CA inhibitors, in hormone receptor-positive HER2 negative subtypes  @andre_2019_alpelisib_nengljmeda.
-The PIK3CA mutation is the second most common driver mutation after TP53, and is most frequently detected in endometrial carcinoma (45%), followed by breast invasive carcinoma (24%), cervical squamous cell carcinoma, endo-cervical adenocarcinoma (20%) and colon adenocarcinoma (16%).
+Breast cancer having *PIK3CA* mutations can be treated using *PIK3CA* inhibitors, in hormone receptor-positive HER2 negative subtypes  @andre_2019_alpelisib_nengljmeda.
+The *PIK3CA* mutation is the second most common driver mutation after TP53, and is most frequently detected in endometrial carcinoma (45%), followed by breast invasive carcinoma (24%), cervical squamous cell carcinoma, endo-cervical adenocarcinoma (20%) and colon adenocarcinoma (16%) @baileyComprehensiveCharacterizationCancer2018.
 
-PIK3CA encodes the p110$\alpha$ catalytic subunit of phosphatidylinositol 3′-kinase (PI3K).
+*PIK3CA* encodes the p110$\alpha$ catalytic subunit of phosphatidylinositol 3′-kinase (PI3K).
 PI3K is a protein kinase that phosphorylates phosphatidylinositol 4,5-biphosphate (PIP~2~) to generate phosphatidylinositol 3,4,5-triphosphate (PIP~3~).
-The phosphatase and tensin homolog (PTEN) converts PIP~2~ to PIP~3~ in contrast to PI3K.
+The phosphatase and tensin homolog (*PTEN*) converts PIP~2~ to PIP~3~ in contrast to PI3K.
 PIP~3~ is a second messenger that activates protein kinase B (AKT), which is a serine/threonine-specific protein kinase.
@@
 
-We applied a supervised elastic net penalized logistic regression model to predict PIK3CA mutations.
-We wanted to ascertain whether this prediction model approach could be applied not only to RAS pathway activation, but also to PIK3CA mutation predictions across several cancer types.
+We applied a supervised elastic net penalized logistic regression model to predict *PIK3CA* mutations.
+We wanted to ascertain whether this prediction model approach could be applied not only to RAS pathway activation, but also to *PIK3CA* mutation predictions.
 
@@
 The TCGA dataset is publically available.
-PIK3CA mutation data was extracted using cgdsr rpackage [@jacobsenCgdsrRbasedAPI2019].
+*PIK3CA* mutation data was extracted using cgdsr rpackage [@jacobsenCgdsrRbasedAPI2019].
 Gene expression data was downloaded from the National Cancer Institute (NCI)’s Genomic Data Commons (GDC) website.
@@
 Gene expression in the TCGA pan-cancer dataset is batch-corrected with normalization.
-The target variable was PIK3CA mutation status.
-PIK3CA status was considered positive when the case had the following PIK3CA variants, which were the target variants of the Therascreen PIK3CA RGQ PCR Kit ; C420R, E542K, E545A, E545D, E545G, E545K, Q546E, Q546R, H1047L, H1047R, H1047Y.
-This kit  was approved as a companion diagnostics test to treat with PIK3CA inhibitor  by the United States Food and Drug Administration.
-We split three-quarters of the dataset into the train-set and one quarter into the test-set.
+
+The target variable was *PIK3CA* mutation status.
+*PIK3CA* status was considered positive when the case had the following *PIK3CA* variants (C420R, E542K, E545A, E545D, E545G, E545K, Q546E, Q546R, H1047L, H1047R, H1047Y) which were the target variants of the Therascreen *PIK3CA* RGQ PCR Kit, Qiagen, Hilden, Germany.
+This kit  was approved as a companion diagnostics test to treat with *PIK3CA* inhibitor by the United States Food and Drug Administration.
 
@@
 Thirty three cancer type dummy variables were included in predictor variables.
+We split three-quarters of the dataset into the training set and one quarter into the test set.
 Yeo-Johnson transformation was performed to correct skewness.
@@
 Penalized logistic regression was applied to prediction modeling.
-Ten-fold cross-validation with targe  variable stratification was performed over the hyper-parameter grid: $\lambda$ {$10^{-5}$, $10^{-4}$,$10^{-3}$,$10^{-2}$,$10^{-1}$, $10^{0}$}, $\alpha$ {0.0, 0.25, 0.5, 0.75}.
+Ten-fold cross-validation with target variable stratification was performed over the hyper-parameter grid: $\lambda$ {$10^{-5}$, $10^{-4}$,$10^{-3}$,$10^{-2}$,$10^{-1}$, $10^{0}$}, $\alpha$ {0.0, 0.25, 0.5, 0.75}.
 Lambda ($\lambda$) is a penalty scaling parameter and alpha ($\alpha$) is a mixing parameter of penalty function ($(1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1$) [@friedmanRegularizationPathsGeneralized2010].
@@
 
-10,845 cases were available for both PIK3CA mutation and mRNA expression data.
+10,845 cases were available for both *PIK3CA* mutation and mRNA expression data.
 5,128 out of 20,502 genes were included in the modeling process, after filtering for median absolute deviation, as described in the modeling process method.
-The prevalence rate for PIK3CA was 0.11 in all cases.
-The PIK3CA prevalence rate in each cancer type varied.
-The median prevalence rate of PIK3CA for each cancer type was 0.03 (range 0–0.33) (Figure 1).
+The prevalence rate for *PIK3CA* was 0.11 in all cases.
+The *PIK3CA* prevalence rate in each cancer type varied.
+The median prevalence rate of *PIK3CA* for each cancer type was 0.03 (range 0–0.33) (Figure 1).
 
@@
 For 10-fold cross-validation, the model with $\lambda$ = 0.01 and $\alpha$ = 1.0 (Ridge regression) showed the best performance in terms of AUROC.
-The final model was trained with the selected hyper-parameters with the entire train-set.
-The train-set AUROC was 0.93 and the test-set AUROC was 0.84.
-The AUPR of the train-set was 0.66 and the test-set AUPR was 0.39 (Figure 2A).
+The final model was trained with the selected hyper-parameters with the entire training set.
+The training set AUROC was 0.93 and the test set AUROC was 0.84.
+The AUPR of the training set was 0.66 and the test set AUPR was 0.39 (Figure 2A).
 
@@
 
-Because PIK3CA mutation prevalence varied across cancer types, the performance of each cancer type was investigated.
-The AUROC and AUPR were positively correlated between the train-sets and test-sets in cancer type sub-analysis (Figure 2B).
-The AUPR was high in cancer types with high PIK3CA mutation rates such as colon, breast and uterus cancer types.
-The AUROC did not correlate with PIK3CA mutation rates of each cancer type (Figure 2C).
+Because *PIK3CA* mutation prevalence varied across cancer types, the performance of each cancer type was investigated.
+The AUROC and AUPR were positively correlated between the training sets and test sets in cancer type sub-analysis (Figure 2B).
+The AUPR was high in cancer types with high *PIK3CA* mutation rates such as colon, breast and uterus cancer types.
+The AUROC did not correlate with *PIK3CA* mutation rates of each cancer type (Figure 2C).
 
@@
 The coefficient is the parameter of the predictor which represents the effect of the predictor on prediction.
-Insulin Like Growth Factor 1 Receptor (IGF1R)  mRNA expression was the strongest negative predictor, and PTEN was the strongest positive predictor.
-Both IGFR1R  and PTEN are key players in the tyrosine kinase pathway .
+*insulin like growth factor 1 Receptor* (*IGF1R*)  mRNA expression was the strongest negative predictor, and *PTEN* was the strongest positive predictor.
+Both *IGF1R*  and *PTEN* are key players in the tyrosine kinase pathway .
 The cancer types were important predictors.
@@
 
-Our model showed good performance in predicting PIK3CA mutations in various cancer types.
+Our model showed good performance in predicting *PIK3CA* mutations in various cancer types.
 Our data suggested that the supervised elastic net penalized logistic regression model could be applied not only to the RAS activation pathway, but also to other genomic alterations.
-Both the RAS activation pathway and PIK3CA mutations are key, common cancer genomic alterations.
+Both the RAS activation pathway and *PIK3CA* mutations are key, common cancer genomic alterations.
 When mutated, they exert significant effect on gene expression in cells.
@@
 
-Our PIK3CA prediction model was similar to the RAS activation prediction model in terms of AUROC (0.84).
+Our *PIK3CA* prediction model was similar to the RAS activation prediction model in terms of AUROC (0.84).
 However the AUPR of our model was lower than the RAS activation model (0.39 versus 0.63).
-The reason for our lower AUPR may be explained by the low prevalence rate of PIK3CA mutations, and an imbalanced dataset.
+The reason for our lower AUPR may be explained by the low prevalence rate of *PIK3CA* mutations, and an imbalanced dataset.
 The model for RAS activation trained with cancer types with more than 0.05 prevalence of RAS activation to avoid imbalance classification problem.
@@
 The lower prevalence rate of target variables meant our dataset had a lower AUPR baseline.
-In the sub-analysis performance of each cancer type, the cancer types with higher PIK3CA mutation rates showed better AUPRs.
+In the sub-analysis performance of each cancer type, the cancer types with higher *PIK3CA* mutation rates showed better AUPRs.
 
@@
 
-Some significant gene expression predictors were closely related to PTEN and the PI3K pathway.
-PTEN and IGFR1R were the strongest gene expression predictors, which has negative and positive predictive powers.
-IGF1R is a tyrosine kinase receptor that activates PI3K  [@leroithInsulinlikeGrowthFactor2003].
-], and PTEN is an important regulator of PIP~3~ by dephosphorylating PIP~3~, in contrast to PI3K [@cantleyPhosphoinositide3KinasePathway2002].
+Some significant gene expression predictors were closely related to *PTEN* and the PI3K pathway.
+*PTEN* and IGFR1R were the strongest gene expression predictors, which has negative and positive predictive powers.
+*IGF1R* is a tyrosine kinase receptor that activates PI3K  [@leroithInsulinlikeGrowthFactor2003].
+], and *PTEN* is an important regulator of PIP~3~ by dephosphorylating PIP~3~, in contrast to PI3K [@cantleyPhosphoinositide3KinasePathway2002].
 
-Several studies have attempted to predict genomic alterations from gene expression data [@cizkovaGeneExpressionProfiling2010; @loiPIK3CAMutationsAssociated2010].
-A study investigated PIK3CA mutation predictions using gene-expression signatures which is a sum of the average of the logarithmic gene expression .
-The model showed good performance AUROC 0.71 in an independent test set  [@cizkovaGeneExpressionProfiling2010; @loiPIK3CAMutationsAssociated2010].Another study predicted copy number alterations with gene expression, using a multinomial logistic regression model with least absolute shrinkage and selection operator (LASSO) parameters [@muCNAPEMachineLearning2019].
+Several studies have attempted to predict genomic alterations from gene expression data [@cizkovaGeneExpressionProfiling2010; @loi*PIK3CA*MutationsAssociated2010].
+A study investigated *PIK3CA* mutation predictions using gene-expression signatures which is a sum of the average of the logarithmic gene expression .
+The model showed good performance AUROC 0.71 in an independent test set  [@cizkovaGeneExpressionProfiling2010; @loi*PIK3CA*MutationsAssociated2010].Another study predicted copy number alterations with gene expression, using a multinomial logistic regression model with least absolute shrinkage and selection operator (LASSO) parameters [@muCNAPEMachineLearning2019].
 The prediction of the 1p/19q codel  was very good, with an AUROC of 0.997, and gene-level predictions were good, with an AUROC of 0.75 [@muCNAPEMachineLearning2019].
@@
 
-- Figure 1. Prevalence rate of PIK3CA mutations across cancer types. Cancer type abbreviations are explained in the S1 Appendix.
+- Figure 1. Prevalence rate of *PIK3CA* mutations across cancer types. Cancer type abbreviations are explained in the S1 Appendix.
 
@@
 (A) Left: receiver operating characteristic (ROC) curve.
-Right: precision-recall (PR) curve of train-set and test-set.
-The horizontal green line is the PIK3CA mutation rate (0.11) (B) Correlation between train-set and test-set of the area under the receiver operating characteristic curve (AUROC), and the area under the precision-recall curve (AUPR) among cancer types.
+Right: precision-recall (PR) curve of training set and test set.
+The horizontal green line is the *PIK3CA* mutation rate (0.11) (B) Correlation between training set and test set of the area under the receiver operating characteristic curve (AUROC), and the area under the precision-recall curve (AUPR) among cancer types.
 The gray band is the 95% confidence interval.
 Abbreviations are explained in the S1 Appendix.
-(C) Correlations between the PIK3CA mutation rate of the AUROC, and the AUPR.
+(C) Correlations between the *PIK3CA* mutation rate of the AUROC, and the AUPR.
 

### 2020-06-12 plosOne/plosOne.tex
itelic

@@
 {\Large
-\textbf\newline{Prediction of PIK3CA mutation with gene expression in cancer} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+\textbf\newline{Prediction of PIK3CA mutations from cancer gene expression data} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
 }
@@
 characteristic (AUROC). The final model was developed with selected
-hyper-parameters using the entire train-set. The train-set AUROC was
-0.93, and the test-set AUROC was 0.84. The area under the
-precision-recall (AUPR) of the train-set was 0.66, and the test-set AUPR
-was 0.39. Cancer types were the most important predictors. Both IGFR1R
-and the phosphatase and tensin homolog (PTEN) were the most significant
-genes in gene expression predictors. Our study suggests that predicting
-genomic alterations using gene expression data is possible, with good
-outcomes.
+hyper-parameters using the entire training set. The training set AUROC
+was 0.93, and the test set AUROC was 0.84. The area under the
+precision-recall (AUPR) of the training set was 0.66, and the test set
+AUPR was 0.39. Cancer types were the most important predictors. Both
+nsulin Like Growth Factor 1 Receptor (IGF1R) and the phosphatase and
+tensin homolog (PTEN) were the most significant genes in gene expression
+predictors. Our study suggests that predicting genomic alterations using
+gene expression data is possible, with good outcomes.
 
@@
 tests have been developed and proven for their clinical utility to treat
-patients. Machine learning approaches can be applied to detect genomic
-alterations . Machine learning algorithms can build prediction models
-from a large number of predictors, such as radiomic features {[}1{]},
-pathology image {[}2{]} or gene expression data {[}3{]}. Because most
-direct genomic tests are more specific and sensitive than predictive
-models, machine learning approaches may have limited roles in clinical
-practice, however, machine learning approaches are ideal when direct
-tests are unavailable or fail. RAS pathway activation predictions have
-been performed using gene expression data {[}5{]}. Authors used data
-from The Cancer Genome Atlas (TCGA), with a supervised elastic net
-penalized logistic regression classifier, with stochastic gradient
-descent. Their model performance was 84\% with an area under the
-receiver operating characteristic (AUROC) curve, and 63\% with an area
-under the precision-recall (AUPR) curve. Importantly, these authors
-suggested their approach could be applied to other genomic alterations.
+patients {[}1{]}.
+
+Machine learning approaches can be applied to detect genomic alterations
+. Machine learning algorithms can build prediction models from a large
+number of predictors, such as radiomic features {[}3{]}, pathology image
+{[}4{]} or gene expression data {[}5{]}. Because most direct genomic
+tests are more specific and sensitive than predictive models, machine
+learning approaches may have limited roles in clinical practice,
+however, machine learning approaches are ideal when direct tests are
+unavailable or fail.
+
+RAS pathway activation predictions have been performed using gene
+expression data {[}7{]}. Authors used data from The Cancer Genome Atlas
+(TCGA), with a supervised elastic net penalized logistic regression
+classifier, with stochastic gradient descent. Their model performance
+was 84\% with an area under the receiver operating characteristic
+(AUROC) curve, and 63\% with an area under the precision-recall (AUPR)
+curve. Importantly, these authors suggested their approach could be
+applied to other genomic alterations.
+
 Breast cancer expressing PIK3CA mutations can be treated using PIK3CA
-inhibitors, in hormone receptor-positive HER2 negative subtypes {[}6{]}.
+inhibitors, in hormone receptor-positive HER2 negative subtypes {[}8{]}.
 The PIK3CA mutation is the second most common driver mutation after
@@
 carcinoma, endo-cervical adenocarcinoma (20\%) and colon adenocarcinoma
-(16\%). PIK3CA encodes the p110\(\alpha\) catalytic subunit of
+(16\%) {[}9{]}.
+
+PIK3CA encodes the p110\(\alpha\) catalytic subunit of
 phosphatidylinositol 3′-kinase (PI3K). PI3K is a protein kinase that
@@
 protein kinase. AKT inhibits apoptosis and promotes cell proliferation
-{[}7{]}.
+{[}10{]}.
 
@@
 cancer types. The TCGA dataset is publically available. PIK3CA mutation
-data was extracted using cgdsr rpackage {[}8{]}. Gene expression data
+data was extracted using cgdsr rpackage {[}11{]}. Gene expression data
 was downloaded from the National Cancer Institute (NCI)'s Genomic Data
@@
 expression in the TCGA pan-cancer dataset is batch-corrected with
-normalization. The target variable was PIK3CA mutation status. PIK3CA
-status was considered positive when the case had the following PIK3CA
-variants, which were the target variants of the Therascreen PIK3CA RGQ
-PCR Kit ; C420R, E542K, E545A, E545D, E545G, E545K, Q546E, Q546R,
-H1047L, H1047R, H1047Y. This kit was approved as a companion diagnostics
-test to treat with PIK3CA inhibitor by the United States Food and Drug
-Administration. We split three-quarters of the dataset into the
-train-set and one quarter into the test-set.
+normalization.
+
+The target variable was PIK3CA mutation status. PIK3CA status was
+considered positive when the case had the following PIK3CA variants
+(C420R, E542K, E545A, E545D, E545G, E545K, Q546E, Q546R, H1047L, H1047R,
+H1047Y) which were the target variants of the Therascreen PIK3CA RGQ PCR
+Kit, Qiagen, Hilden, Germany. This kit was approved as a companion
+diagnostics test to treat with PIK3CA inhibitor by the United States
+Food and Drug Administration.
 
@@
 deviation (\textgreater{} third-quartiles) were selected. Thirty three
-cancer type dummy variables were included in predictor variables.
-Yeo-Johnson transformation was performed to correct skewness. Centering
-and scaling were also performed. All preprocessing was performed using
-the recipe r package {[}9{]}. Penalized logistic regression was applied
-to prediction modeling. Ten-fold cross-validation with targe variable
-stratification was performed over the hyper-parameter grid: \(\lambda\)
-\{\(10^{-5}\), \(10^{-4}\),\(10^{-3}\),\(10^{-2}\),\(10^{-1}\),
-\(10^{0}\)\}, \(\alpha\) \{0.0, 0.25, 0.5, 0.75\}. Lambda (\(\lambda\))
-is a penalty scaling parameter and alpha (\(\alpha\)) is a mixing
-parameter of penalty function
+cancer type dummy variables were included in predictor variables. We
+split three-quarters of the dataset into the training set and one
+quarter into the test set. Yeo-Johnson transformation was performed to
+correct skewness. Centering and scaling were also performed. All
+preprocessing was performed using the recipe r package {[}12{]}.
+Penalized logistic regression was applied to prediction modeling.
+Ten-fold cross-validation with target variable stratification was
+performed over the hyper-parameter grid: \(\lambda\) \{\(10^{-5}\),
+\(10^{-4}\),\(10^{-3}\),\(10^{-2}\),\(10^{-1}\), \(10^{0}\)\},
+\(\alpha\) \{0.0, 0.25, 0.5, 0.75\}. Lambda (\(\lambda\)) is a penalty
+scaling parameter and alpha (\(\alpha\)) is a mixing parameter of
+penalty function
 (\((1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1\))
-{[}10{]}.
+{[}13{]}.
 
@@
 The AUPR approach is more informative than AUROC for imbalanced datasets
-{[}11{]}. The modeling process and accessing model performance were
-performed with the tidymodels rpackage {[}12{]}.
+{[}14{]}. The modeling process and accessing model performance were
+performed with the tidymodels rpackage {[}15{]}.
 
@@
 of AUROC. The final model was trained with the selected hyper-parameters
-with the entire train-set. The train-set AUROC was 0.93 and the test-set
-AUROC was 0.84. The AUPR of the train-set was 0.66 and the test-set AUPR
-was 0.39 (Figure 2A).
+with the entire training set. The training set AUROC was 0.93 and the
+test set AUROC was 0.84. The AUPR of the training set was 0.66 and the
+test set AUPR was 0.39 (Figure 2A).
 
@@
 performance of each cancer type was investigated. The AUROC and AUPR
-were positively correlated between the train-sets and test-sets in
+were positively correlated between the training sets and test sets in
 cancer type sub-analysis (Figure 2B). The AUPR was high in cancer types
@@
 the parameter of the predictor which represents the effect of the
-predictor on prediction . IGF1R mRNA expression was the strongest
-negative predictor, and PTEN was the strongest positive predictor. Both
-IGFR1R and PTEN are key players in the tyrosine kinase pathway . The
-cancer types were important predictors. Some cancer types including
-uterine carcinosarcoma (UCS), bladder urothelial carcinoma (BLCA),
-pancreatic adenocarcinoma (PAAD), lymphoid neoplasm diffuse large B-cell
-lymphoma (DLBC) were the strongest predictors.
+predictor on prediction. Insulin Like Growth Factor 1 Receptor (IGF1R)
+mRNA expression was the strongest negative predictor, and PTEN was the
+strongest positive predictor. Both IGF1R and PTEN are key players in the
+tyrosine kinase pathway . The cancer types were important predictors.
+Some cancer types including uterine carcinosarcoma (UCS), bladder
+urothelial carcinoma (BLCA), pancreatic adenocarcinoma (PAAD), lymphoid
+neoplasm diffuse large B-cell lymphoma (DLBC) were the strongest
+predictors.
 
@@
 predictors, which has negative and positive predictive powers. IGF1R is
-a tyrosine kinase receptor that activates PI3K {[}13{]}. {]}, and PTEN
+a tyrosine kinase receptor that activates PI3K {[}16{]}. {]}, and PTEN
 is an important regulator of PIP\textsubscript{3} by dephosphorylating
-PIP\textsubscript{3}, in contrast to PI3K {[}7{]}.
+PIP\textsubscript{3}, in contrast to PI3K {[}10{]}.
 
 Several studies have attempted to predict genomic alterations from gene
-expression data {[}4,14{]}. A study investigated PIK3CA mutation
+expression data {[}6,17{]}. A study investigated PIK3CA mutation
 predictions using gene-expression signatures which is a sum of the
 average of the logarithmic gene expression . The model showed good
-performance AUROC 0.71 in an independent test set {[}4,14{]}.Another
+performance AUROC 0.71 in an independent test set {[}6,17{]}.Another
 study predicted copy number alterations with gene expression, using a
 multinomial logistic regression model with least absolute shrinkage and
-selection operator (LASSO) parameters {[}15{]}. The prediction of the
+selection operator (LASSO) parameters {[}18{]}. The prediction of the
 1p/19q codel was very good, with an AUROC of 0.997, and gene-level
-predictions were good, with an AUROC of 0.75 {[}15{]}. A Hidden Markov
+predictions were good, with an AUROC of 0.75 {[}18{]}. A Hidden Markov
 Model Approach for Prediction of Genomic Alterations from Gene
-Expression Profiling {[}16{]}. A logistic regression model was used for
-MYCN gene amplification in neuroblastoma {[}17{]}.
+Expression Profiling {[}19{]}. A logistic regression model was used for
+MYCN gene amplification in neuroblastoma {[}20{]}.
 
@@
   characteristic (ROC) curve. Right: precision-recall (PR) curve of
-  train-set and test-set. The horizontal green line is the PIK3CA
-  mutation rate (0.11) (B) Correlation between train-set and test-set of
-  the area under the receiver operating characteristic curve (AUROC),
+  training set and test set. The horizontal green line is the PIK3CA
+  mutation rate (0.11) (B) Correlation between training set and test set
+  of the area under the receiver operating characteristic curve (AUROC),
   and the area under the precision-recall curve (AUPR) among cancer
@@
 \hypertarget{refs}{}
+\leavevmode\hypertarget{ref-healthNucleicAcidBased2020}{}%
+1. Health C for D and R. Nucleic Acid Based Tests. FDA. FDA;
+
+\leavevmode\hypertarget{ref-sahnanePyrosequencingEGFRMutation2013}{}%
+2. Sahnane N, Gueli R, Tibiletti M, Bernasconi B, Stefanoli M, Franzi F,
+et al. Pyrosequencing for EGFR Mutation Detection: Diagnostic Accuracy
+and Clinical Implications. Diagnostic Molecular Pathology. 2013;22:
+196--203.
+doi:\href{https://doi.org/10.1097/PDM.0b013e3182893f55}{10.1097/PDM.0b013e3182893f55}
+
 \leavevmode\hypertarget{ref-dercleIdentificationNonSmall2020}{}%
-1. Dercle L, Fronheiser M, Lu L, Du S, Hayes W, Leung DK, et al.
+3. Dercle L, Fronheiser M, Lu L, Du S, Hayes W, Leung DK, et al.
 Identification of NonSmall Cell Lung Cancer Sensitive to Systemic Cancer
@@
 \leavevmode\hypertarget{ref-coudrayClassificationMutationPrediction2018}{}%
-2. Coudray N, Ocampo PS, Sakellaropoulos T, Narula N, Snuderl M, Fenyö
+4. Coudray N, Ocampo PS, Sakellaropoulos T, Narula N, Snuderl M, Fenyö
 D, et al. Classification and mutation prediction from nonSmall cell lung
@@
 \leavevmode\hypertarget{ref-wayMachineLearningDetects2018}{}%
-3. Way GP, Sanchez-Vega F, La K, Armenia J, Chatila WK, Luna A, et al.
+5. Way GP, Sanchez-Vega F, La K, Armenia J, Chatila WK, Luna A, et al.
 Machine Learning Detects Pan-cancer Ras Pathway Activation in The Cancer
@@
 \leavevmode\hypertarget{ref-loiPIK3CAMutationsAssociated2010}{}%
-4. Loi S, Haibe-Kains B, Majjaj S, Lallemand F, Durbecq V, Larsimont D,
+6. Loi S, Haibe-Kains B, Majjaj S, Lallemand F, Durbecq V, Larsimont D,
 et al. PIK3CA mutations associated with gene signature of low mTORC1
@@
 \leavevmode\hypertarget{ref-way_2018_machine_cellreports}{}%
-5. Way GP, Sanchez-Vega F, La K, Armenia J, Chatila WK, Luna A, et al.
+7. Way GP, Sanchez-Vega F, La K, Armenia J, Chatila WK, Luna A, et al.
 Machine Learning Detects Pan-cancer Ras Pathway Activation in The Cancer
@@
 \leavevmode\hypertarget{ref-andre_2019_alpelisib_nengljmeda}{}%
-6. André F, Ciruelos E, Rubovszky G, Campone M, Loibl S, Rugo HS, et al.
+8. André F, Ciruelos E, Rubovszky G, Campone M, Loibl S, Rugo HS, et al.
 Alpelisib for PIK3CA-Mutated, Hormone ReceptorPositive Advanced Breast
@@
 
+\leavevmode\hypertarget{ref-baileyComprehensiveCharacterizationCancer2018}{}%
+9. Bailey MH, Tokheim C, Porta-Pardo E, Sengupta S, Bertrand.
+Comprehensive Characterization of Cancer Driver Genes and Mutations.
+Cell. 2018;173: 371--385.e18.
+doi:\href{https://doi.org/10.1016/j.cell.2018.02.060}{10.1016/j.cell.2018.02.060}
+
 \leavevmode\hypertarget{ref-cantleyPhosphoinositide3KinasePathway2002}{}%
-7. Cantley LC. The Phosphoinositide 3-Kinase Pathway. Science. American
+10. Cantley LC. The Phosphoinositide 3-Kinase Pathway. Science. American
 Association for the Advancement of Science; 2002;296: 1655--1657.
@@
 \leavevmode\hypertarget{ref-jacobsenCgdsrRbasedAPI2019}{}%
-8. Jacobsen A, Luna A. Cgdsr: R-based API for accessing the MSKCC cancer
-genomics data server (CGDS). 2019.
+11. Jacobsen A, Luna A. Cgdsr: R-based API for accessing the MSKCC
+cancer genomics data server (CGDS). 2019.
 
 \leavevmode\hypertarget{ref-kuhnRecipesPreprocessingTools2020}{}%
-9. Kuhn M, Wickham H. Recipes: Preprocessing tools to create design
+12. Kuhn M, Wickham H. Recipes: Preprocessing tools to create design
 matrices. 2020.
@@
 \leavevmode\hypertarget{ref-friedmanRegularizationPathsGeneralized2010}{}%
-10. Friedman J, Hastie T, Tibshirani R. Regularization Paths for
+13. Friedman J, Hastie T, Tibshirani R. Regularization Paths for
 Generalized Linear Models via Coordinate Descent. J Stat Softw. 2010;33:
@@
 \leavevmode\hypertarget{ref-saitoPrecisionRecallPlotMore2015}{}%
-11. Saito T, Rehmsmeier M. The Precision-Recall Plot Is More Informative
+14. Saito T, Rehmsmeier M. The Precision-Recall Plot Is More Informative
 than the ROC Plot When Evaluating Binary Classifiers on Imbalanced
@@
 \leavevmode\hypertarget{ref-kuhnTidymodelsEasilyInstall2020}{}%
-12. Kuhn M, Wickham H. Tidymodels: Easily install and load the
+15. Kuhn M, Wickham H. Tidymodels: Easily install and load the
 'tidymodels' packages. 2020.
@@
 \leavevmode\hypertarget{ref-leroithInsulinlikeGrowthFactor2003}{}%
-13. LeRoith D, Roberts CT. The insulin-like growth factor system and
+16. LeRoith D, Roberts CT. The insulin-like growth factor system and
 cancer. Cancer Letters. 2003;195: 127--137.
@@
 \leavevmode\hypertarget{ref-cizkovaGeneExpressionProfiling2010}{}%
-14. Cizkova M, Cizeron-Clairac G, Vacher S, Susini A, Andrieu C,
+17. Cizkova M, Cizeron-Clairac G, Vacher S, Susini A, Andrieu C,
 Lidereau R, et al. Gene Expression Profiling Reveals New Aspects of
@@
 \leavevmode\hypertarget{ref-muCNAPEMachineLearning2019}{}%
-15. Mu Q, Wang J. CNAPE: A Machine Learning Method for Copy Number
+18. Mu Q, Wang J. CNAPE: A Machine Learning Method for Copy Number
 Alteration Prediction from Gene Expression. IEEE/ACM Transactions on
@@
 \leavevmode\hypertarget{ref-gengHiddenMarkovModel2008}{}%
-16. Geng H, Ali HH, Chan WC. A Hidden Markov Model Approach for
+19. Geng H, Ali HH, Chan WC. A Hidden Markov Model Approach for
 Prediction of Genomic Alterations from Gene Expression Profiling. In:
@@
 \leavevmode\hypertarget{ref-heGeneSignaturesAssociated2020}{}%
-17. He X, Qin C, Zhao Y, Zou L, Zhao H, Cheng C. Gene signatures
+20. He X, Qin C, Zhao Y, Zou L, Zhao H, Cheng C. Gene signatures
 associated with genomic aberrations predict prognosis in neuroblastoma.

### 2020-07-16 plosOne/plosOne.Rmd
proof reading

@@
 title: Prediction of *PIK3CA* mutations from cancer gene expression data
-short-title: *PIK3CA* mutations prediction  
+short-title: PIK3CA mutations prediction
 author:
@@
 
-Machine learning approaches can be applied to detect genomic alterations .
-Machine learning algorithms can build prediction models from a large number of predictors, such as radiomic  features @dercleIdentificationNonSmall2020, pathology image [@coudrayClassificationMutationPrediction2018] or gene expression data [@wayMachineLearningDetects2018 @loi*PIK3CA*MutationsAssociated2010].
+Machine learning approaches can be applied to detect genomic alterations.
+Machine learning algorithms can build prediction models from a large number of predictors, such as radiomic  features @dercleIdentificationNonSmall2020, pathology image [@coudrayClassificationMutationPrediction2018] or gene expression data [@wayMachineLearningDetects2018 @loiPIK3CAMutationsAssociated2010].
 Because most direct genomic tests are more specific and sensitive than predictive models, machine learning approaches may have limited roles in clinical practice, however, machine learning approaches are ideal when direct tests are unavailable or fail.
 
-RAS pathway activation predictions have been performed using gene expression data @way_2018_machine_cellreports.
+RAS pathway activation predictions have been performed using gene expression data @wayMachineLearningDetects2018.
 Authors used data from The Cancer Genome Atlas (TCGA), with a supervised elastic net penalized logistic regression classifier, with stochastic gradient descent.
@@
 
-Breast cancer having *PIK3CA* mutations can be treated using *PIK3CA* inhibitors, in hormone receptor-positive HER2 negative subtypes  @andre_2019_alpelisib_nengljmeda.
-The *PIK3CA* mutation is the second most common driver mutation after TP53, and is most frequently detected in endometrial carcinoma (45%), followed by breast invasive carcinoma (24%), cervical squamous cell carcinoma, endo-cervical adenocarcinoma (20%) and colon adenocarcinoma (16%) @baileyComprehensiveCharacterizationCancer2018.
+Breast cancer having *PIK3CA* mutations can be treated using PIK3CA inhibitors, in hormone receptor-positive HER2 negative subtypes  @andre_2019_alpelisib_nengljmeda.
+The *PIK3CA* mutation is the second most common driver mutation after *TP53*, and is most frequently detected in endometrial carcinoma (45%), followed by breast invasive carcinoma (24%), cervical squamous cell carcinoma, endo-cervical adenocarcinoma (20%) and colon adenocarcinoma (16%) @baileyComprehensiveCharacterizationCancer2018.
 
@@
 *PIK3CA* status was considered positive when the case had the following *PIK3CA* variants (C420R, E542K, E545A, E545D, E545G, E545K, Q546E, Q546R, H1047L, H1047R, H1047Y) which were the target variants of the Therascreen *PIK3CA* RGQ PCR Kit, Qiagen, Hilden, Germany.
-This kit  was approved as a companion diagnostics test to treat with *PIK3CA* inhibitor by the United States Food and Drug Administration.
+This kit  was approved as a companion diagnostics test to treat with PIK3CA inhibitor by the United States Food and Drug Administration.
 
@@
 Penalized logistic regression was applied to prediction modeling.
-Ten-fold cross-validation with target variable stratification was performed over the hyper-parameter grid: $\lambda$ {$10^{-5}$, $10^{-4}$,$10^{-3}$,$10^{-2}$,$10^{-1}$, $10^{0}$}, $\alpha$ {0.0, 0.25, 0.5, 0.75}.
+Ten-fold cross-validation with target variable stratification was performed over the hyper-parameter grid: $\lambda$ {$10^{-5}$, $10^{-4}$,$10^{-3}$,$10^{-2}$,$10^{-1}$, $10^{0}$}, $\alpha$ {0.0, 0.25, 0.5, 0.75, 1.0}.
 Lambda ($\lambda$) is a penalty scaling parameter and alpha ($\alpha$) is a mixing parameter of penalty function ($(1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1$) [@friedmanRegularizationPathsGeneralized2010].
 
-## Accessing model performance
+## Assessing model performance
 
@@
 The AUPR approach is more informative than AUROC for imbalanced datasets [@saitoPrecisionRecallPlotMore2015].
-The modeling process and accessing  model performance were performed with the tidymodels rpackage  [@kuhnTidymodelsEasilyInstall2020].
+The modeling process and assessing  model performance were performed with the tidymodels rpackage  [@kuhnTidymodelsEasilyInstall2020].
 
@@
 5,128 out of 20,502 genes were included in the modeling process, after filtering for median absolute deviation, as described in the modeling process method.
-The prevalence rate for *PIK3CA* was 0.11 in all cases.
-The *PIK3CA* prevalence rate in each cancer type varied.
-The median prevalence rate of *PIK3CA* for each cancer type was 0.03 (range 0–0.33) (Figure 1).
+The prevalence rate for *PIK3CA* mutation was 0.11 in all cases.
+The *PIK3CA* mutation prevalence rate in each cancer type varied.
+The median prevalence rate of *PIK3CA* mutation for each cancer type was 0.03 (range 0–0.33) (Figure 1).
 
@@
 The coefficient is the parameter of the predictor which represents the effect of the predictor on prediction.
-*insulin like growth factor 1 Receptor* (*IGF1R*)  mRNA expression was the strongest negative predictor, and *PTEN* was the strongest positive predictor.
-Both *IGF1R*  and *PTEN* are key players in the tyrosine kinase pathway .
+*Insulin like growth factor 1 Receptor* (*IGF1R*)  mRNA expression was the strongest negative predictor, and *PTEN* was the strongest positive predictor.
+Both *IGF1R*  and *PTEN* are key players in the tyrosine kinase pathway [@leroithInsulinlikeGrowthFactor2003; @cantleyPhosphoinositide3KinasePathway2002].
 The cancer types were important predictors.
@@
 Both the RAS activation pathway and *PIK3CA* mutations are key, common cancer genomic alterations.
-When mutated, they exert significant effect on gene expression in cells.
-However, the supervised elastic net penalized logistic regression model cannot be generalized or applied to other genomic alterations.
-It might be a challenging prediction of genomic alterations that are infrequent or have a weak effect on gene expression.
+Because they exert significant effect on gene expression in cancer cells, prediction from gene expression data can be good.
+However, the supervised elastic net penalized logistic regression model cannot be generalized or applied to other genomic alterations which have have a weak effect on gene expression.
 
@@
 The gene expression data is processed by between-sample normalization to remove batch effects.
-If the model has been trained from between-sample normalization, a new sample cannot be exactly preprocessed with normalization which was done on trainset.
-A model based on gene expression from the TCGA pan-cancer dataset is limited in terms of data preprocessing.
-It is necessary to develop a preprocessing method that is independent of a dataset, to apply gene expression data to the prediction model.
+If the model has been trained from between-sample normalization, a new sample cannot be exactly processed with normalization which was done on trainset.
+A model based on gene expression from the TCGA pan-cancer dataset has limitation in terms of data preprocessing.
+It is necessary to develop a processing method that is independent of a dataset, to apply gene expression data to the prediction model.
 
@@
 However the AUPR of our model was lower than the RAS activation model (0.39 versus 0.63).
-The reason for our lower AUPR may be explained by the low prevalence rate of *PIK3CA* mutations, and an imbalanced dataset.
+The reason for our lower AUPR may be explained by an imbalanced dataset that has the low prevalence rate of *PIK3CA* mutations [@wayMachineLearningDetects2018].
 The model for RAS activation trained with cancer types with more than 0.05 prevalence of RAS activation to avoid imbalance classification problem.
@@
 The varying prevalence of PIC3CA mutations across cancer types may be a reason for the strong predictive power of cancer types.
-If the cancer type was wrong or could not be determined, our model performance was poor.
 
-Some significant gene expression predictors were closely related to *PTEN* and the PI3K pathway.
-*PTEN* and IGFR1R were the strongest gene expression predictors, which has negative and positive predictive powers.
-*IGF1R* is a tyrosine kinase receptor that activates PI3K  [@leroithInsulinlikeGrowthFactor2003].
-], and *PTEN* is an important regulator of PIP~3~ by dephosphorylating PIP~3~, in contrast to PI3K [@cantleyPhosphoinositide3KinasePathway2002].
+Some significant gene expression predictors were closely related to the PTEN-PI3K pathway.
+*PTEN* and *IGFR1R* were the strongest gene expression predictors, which has negative and positive predictive powers.
+*IGF1R* is a tyrosine kinase receptor that activates PI3K  [@leroithInsulinlikeGrowthFactor2003], and *PTEN* is an important regulator of PIP~3~ by dephosphorylating PIP~3~ [@cantleyPhosphoinositide3KinasePathway2002].
 
-Several studies have attempted to predict genomic alterations from gene expression data [@cizkovaGeneExpressionProfiling2010; @loi*PIK3CA*MutationsAssociated2010].
-A study investigated *PIK3CA* mutation predictions using gene-expression signatures which is a sum of the average of the logarithmic gene expression .
-The model showed good performance AUROC 0.71 in an independent test set  [@cizkovaGeneExpressionProfiling2010; @loi*PIK3CA*MutationsAssociated2010].Another study predicted copy number alterations with gene expression, using a multinomial logistic regression model with least absolute shrinkage and selection operator (LASSO) parameters [@muCNAPEMachineLearning2019].
-The prediction of the 1p/19q codel  was very good, with an AUROC of 0.997, and gene-level predictions were good, with an AUROC of 0.75 [@muCNAPEMachineLearning2019].
-A Hidden Markov Model Approach for Prediction of Genomic Alterations from Gene Expression Profiling [@gengHiddenMarkovModel2008].
-A logistic regression model was used for MYCN  gene amplification in neuroblastoma [@heGeneSignaturesAssociated2020].
+Several studies have attempted to predict genomic alterations from gene expression data [@cizkovaGeneExpressionProfiling2010; @loiPIK3CAMutationsAssociated2010].
+A study investigated *PIK3CA* mutation predictions using gene-expression signatures which is a sum of the average of the logarithmic gene expression.
+The model showed good performance AUROC 0.71 in an independent test set  [@cizkovaGeneExpressionProfiling2010; @loiPIK3CAMutationsAssociated2010].
+Another study predicted copy number alterations with gene expression, using a multinomial logistic regression model with least absolute shrinkage and selection operator (LASSO) parameters [@muCNAPEMachineLearning2019].
+The prediction of the 1p/19q codeletion  was very good, with an AUROC of 0.997, and gene-level predictions were good, with an AUROC of 0.75 [@muCNAPEMachineLearning2019].
+A logistic regression model was used for *MYCN Proto-Oncogene, BHLH Transcription Factor* (*MYCN*) gene amplification in neuroblastoma [@heGeneSignaturesAssociated2020].
 
 Our study suggested that the prediction of genomic alterations using gene expression data was possible, with good performance.
-However, improved performances are required for clinical tests, and the standardization of generation processing of gene expression data is also needed .
+However, improved performances are required for clinical tests, and the standardization of generation processing of gene expression data is also needed.
 

### 2020-07-16 plosOne/plosOne.tex
proof reading

@@
 {\Large
-\textbf\newline{Prediction of PIK3CA mutations from cancer gene expression data} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
+\textbf\newline{Prediction of \emph{PIK3CA} mutations from cancer gene expression data} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
 }
@@
 \section*{Abstract}
-Breast cancers with PIK3CA mutations can be treated with PIK3CA
-inhibitors in hormone receptor-positive HER2 negative subtypes. We
-applied a supervised elastic net penalized logistic regression model to
-predict PIK3CA mutations from gene expression data. This regression
-approach was applied to predict modeling using the TCGA pan-cancer
-dataset. Approximately 10,845 cases were available for PIK3CA mutation
-and mRNA expression data. In 10-fold cross-validation, the model with
-\(\lambda\) = 0.01 and \(\alpha\) = 1.0 (Ridge regression) showed the
-best performance, in terms of area under the receiver operating
-characteristic (AUROC). The final model was developed with selected
-hyper-parameters using the entire training set. The training set AUROC
-was 0.93, and the test set AUROC was 0.84. The area under the
-precision-recall (AUPR) of the training set was 0.66, and the test set
-AUPR was 0.39. Cancer types were the most important predictors. Both
-nsulin Like Growth Factor 1 Receptor (IGF1R) and the phosphatase and
-tensin homolog (PTEN) were the most significant genes in gene expression
+Breast cancers with \emph{PIK3CA} mutations can be treated with
+\emph{PIK3CA} inhibitors in hormone receptor-positive HER2 negative
+subtypes. We applied a supervised elastic net penalized logistic
+regression model to predict \emph{PIK3CA} mutations from gene expression
+data. This regression approach was applied to predict modeling using the
+TCGA pan-cancer dataset. Approximately 10,000 cases were available for
+\emph{PIK3CA} mutation and mRNA expression data. In 10-fold
+cross-validation, the model with \(\lambda\) = 0.01 and \(\alpha\) = 1.0
+(Ridge regression) showed the best performance, in terms of area under
+the receiver operating characteristic (AUROC). The final model was
+developed with selected hyper-parameters using the entire training set.
+The training set AUROC was 0.93, and the test set AUROC was 0.84. The
+area under the precision-recall (AUPR) of the training set was 0.66, and
+the test set AUPR was 0.39. Cancer types were the most important
+predictors. Both \emph{insulin like growth factor 1 Receptor}
+(\emph{IGF1R}) and the \emph{phosphatase and tensin homolog}
+(\emph{PTEN}) were the most significant genes in gene expression
 predictors. Our study suggests that predicting genomic alterations using
@@
 
-Machine learning approaches can be applied to detect genomic alterations
-. Machine learning algorithms can build prediction models from a large
-number of predictors, such as radiomic features {[}3{]}, pathology image
-{[}4{]} or gene expression data {[}5{]}. Because most direct genomic
-tests are more specific and sensitive than predictive models, machine
-learning approaches may have limited roles in clinical practice,
-however, machine learning approaches are ideal when direct tests are
-unavailable or fail.
+Machine learning approaches can be applied to detect genomic
+alterations. Machine learning algorithms can build prediction models
+from a large number of predictors, such as radiomic features {[}3{]},
+pathology image {[}4{]} or gene expression data {[}5{]}. Because most
+direct genomic tests are more specific and sensitive than predictive
+models, machine learning approaches may have limited roles in clinical
+practice, however, machine learning approaches are ideal when direct
+tests are unavailable or fail.
 
 RAS pathway activation predictions have been performed using gene
-expression data {[}7{]}. Authors used data from The Cancer Genome Atlas
+expression data {[}5{]}. Authors used data from The Cancer Genome Atlas
 (TCGA), with a supervised elastic net penalized logistic regression
@@
 
-Breast cancer expressing PIK3CA mutations can be treated using PIK3CA
-inhibitors, in hormone receptor-positive HER2 negative subtypes {[}8{]}.
-The PIK3CA mutation is the second most common driver mutation after
-TP53, and is most frequently detected in endometrial carcinoma (45\%),
-followed by breast invasive carcinoma (24\%), cervical squamous cell
-carcinoma, endo-cervical adenocarcinoma (20\%) and colon adenocarcinoma
-(16\%) {[}9{]}.
+Breast cancer having \emph{PIK3CA} mutations can be treated using PIK3CA
+inhibitors, in hormone receptor-positive HER2 negative subtypes {[}7{]}.
+The \emph{PIK3CA} mutation is the second most common driver mutation
+after \emph{TP53}, and is most frequently detected in endometrial
+carcinoma (45\%), followed by breast invasive carcinoma (24\%), cervical
+squamous cell carcinoma, endo-cervical adenocarcinoma (20\%) and colon
+adenocarcinoma (16\%) {[}8{]}.
 
-PIK3CA encodes the p110\(\alpha\) catalytic subunit of
+\emph{PIK3CA} encodes the p110\(\alpha\) catalytic subunit of
 phosphatidylinositol 3′-kinase (PI3K). PI3K is a protein kinase that
@@
 3,4,5-triphosphate (PIP\textsubscript{3}). The phosphatase and tensin
-homolog (PTEN) converts PIP\textsubscript{2} to PIP\textsubscript{3} in
-contrast to PI3K. PIP\textsubscript{3} is a second messenger that
-activates protein kinase B (AKT), which is a serine/threonine-specific
-protein kinase. AKT inhibits apoptosis and promotes cell proliferation
-{[}10{]}.
+homolog (\emph{PTEN}) converts PIP\textsubscript{2} to
+PIP\textsubscript{3} in contrast to PI3K. PIP\textsubscript{3} is a
+second messenger that activates protein kinase B (AKT), which is a
+serine/threonine-specific protein kinase. AKT inhibits apoptosis and
+promotes cell proliferation {[}9{]}.
 
 We applied a supervised elastic net penalized logistic regression model
-to predict PIK3CA mutations. We wanted to ascertain whether this
+to predict \emph{PIK3CA} mutations. We wanted to ascertain whether this
 prediction model approach could be applied not only to RAS pathway
-activation, but also to PIK3CA mutation predictions across several
-cancer types.
+activation, but also to \emph{PIK3CA} mutation predictions.
 
@@
 clinical data from \textgreater{} 10,000 cancer samples across 33 common
-cancer types. The TCGA dataset is publically available. PIK3CA mutation
-data was extracted using cgdsr rpackage {[}11{]}. Gene expression data
-was downloaded from the National Cancer Institute (NCI)'s Genomic Data
-Commons (GDC) website. This archives data for TCGA
+cancer types. The TCGA dataset is publically available. \emph{PIK3CA}
+mutation data was extracted using cgdsr rpackage {[}10{]}. Gene
+expression data was downloaded from the National Cancer Institute
+(NCI)'s Genomic Data Commons (GDC) website. This archives data for TCGA
 (https://gdc.cancer.gov/about-data/publications/pancanatlas). Gene
@@
 
-The target variable was PIK3CA mutation status. PIK3CA status was
-considered positive when the case had the following PIK3CA variants
-(C420R, E542K, E545A, E545D, E545G, E545K, Q546E, Q546R, H1047L, H1047R,
-H1047Y) which were the target variants of the Therascreen PIK3CA RGQ PCR
-Kit, Qiagen, Hilden, Germany. This kit was approved as a companion
-diagnostics test to treat with PIK3CA inhibitor by the United States
-Food and Drug Administration.
+The target variable was \emph{PIK3CA} mutation status. \emph{PIK3CA}
+status was considered positive when the case had the following
+\emph{PIK3CA} variants (C420R, E542K, E545A, E545D, E545G, E545K, Q546E,
+Q546R, H1047L, H1047R, H1047Y) which were the target variants of the
+Therascreen \emph{PIK3CA} RGQ PCR Kit, Qiagen, Hilden, Germany. This kit
+was approved as a companion diagnostics test to treat with PIK3CA
+inhibitor by the United States Food and Drug Administration.
 
@@
 correct skewness. Centering and scaling were also performed. All
-preprocessing was performed using the recipe r package {[}12{]}.
+preprocessing was performed using the recipe r package {[}11{]}.
 Penalized logistic regression was applied to prediction modeling.
@@
 \(10^{-4}\),\(10^{-3}\),\(10^{-2}\),\(10^{-1}\), \(10^{0}\)\},
-\(\alpha\) \{0.0, 0.25, 0.5, 0.75\}. Lambda (\(\lambda\)) is a penalty
-scaling parameter and alpha (\(\alpha\)) is a mixing parameter of
-penalty function
+\(\alpha\) \{0.0, 0.25, 0.5, 0.75, 1.0\}. Lambda (\(\lambda\)) is a
+penalty scaling parameter and alpha (\(\alpha\)) is a mixing parameter
+of penalty function
 (\((1-\alpha)/2 \lVert\beta\rVert_2^2+\alpha\lVert \beta \rVert_1\))
-{[}13{]}.
+{[}12{]}.
 
-\hypertarget{accessing-model-performance}{%
-\subsection{Accessing model
-performance}\label{accessing-model-performance}}
+\hypertarget{assessing-model-performance}{%
+\subsection{Assessing model
+performance}\label{assessing-model-performance}}
 
@@
 The AUPR approach is more informative than AUROC for imbalanced datasets
-{[}14{]}. The modeling process and accessing model performance were
-performed with the tidymodels rpackage {[}15{]}.
+{[}13{]}. The modeling process and assessing model performance were
+performed with the tidymodels rpackage {[}14{]}.
 
@@
 
-10,845 cases were available for both PIK3CA mutation and mRNA expression
-data. 5,128 out of 20,502 genes were included in the modeling process,
-after filtering for median absolute deviation, as described in the
-modeling process method. The prevalence rate for PIK3CA was 0.11 in all
-cases. The PIK3CA prevalence rate in each cancer type varied. The median
-prevalence rate of PIK3CA for each cancer type was 0.03 (range 0--0.33)
+10,845 cases were available for both \emph{PIK3CA} mutation and mRNA
+expression data. 5,128 out of 20,502 genes were included in the modeling
+process, after filtering for median absolute deviation, as described in
+the modeling process method. The prevalence rate for \emph{PIK3CA}
+mutation was 0.11 in all cases. The \emph{PIK3CA} mutation prevalence
+rate in each cancer type varied. The median prevalence rate of
+\emph{PIK3CA} mutation for each cancer type was 0.03 (range 0--0.33)
 (Figure 1).
@@
 
-Because PIK3CA mutation prevalence varied across cancer types, the
-performance of each cancer type was investigated. The AUROC and AUPR
+Because \emph{PIK3CA} mutation prevalence varied across cancer types,
+the performance of each cancer type was investigated. The AUROC and AUPR
 were positively correlated between the training sets and test sets in
 cancer type sub-analysis (Figure 2B). The AUPR was high in cancer types
-with high PIK3CA mutation rates such as colon, breast and uterus cancer
-types. The AUROC did not correlate with PIK3CA mutation rates of each
-cancer type (Figure 2C).
+with high \emph{PIK3CA} mutation rates such as colon, breast and uterus
+cancer types. The AUROC did not correlate with \emph{PIK3CA} mutation
+rates of each cancer type (Figure 2C).
 
@@
 the parameter of the predictor which represents the effect of the
-predictor on prediction. Insulin Like Growth Factor 1 Receptor (IGF1R)
-mRNA expression was the strongest negative predictor, and PTEN was the
-strongest positive predictor. Both IGF1R and PTEN are key players in the
-tyrosine kinase pathway . The cancer types were important predictors.
-Some cancer types including uterine carcinosarcoma (UCS), bladder
-urothelial carcinoma (BLCA), pancreatic adenocarcinoma (PAAD), lymphoid
-neoplasm diffuse large B-cell lymphoma (DLBC) were the strongest
-predictors.
+predictor on prediction. \emph{Insulin like growth factor 1 Receptor}
+(\emph{IGF1R}) mRNA expression was the strongest negative predictor, and
+\emph{PTEN} was the strongest positive predictor. Both \emph{IGF1R} and
+\emph{PTEN} are key players in the tyrosine kinase pathway {[}9,15{]}.
+The cancer types were important predictors. Some cancer types including
+uterine carcinosarcoma (UCS), bladder urothelial carcinoma (BLCA),
+pancreatic adenocarcinoma (PAAD), lymphoid neoplasm diffuse large B-cell
+lymphoma (DLBC) were the strongest predictors.
 
@@
 
-Our model showed good performance in predicting PIK3CA mutations in
-various cancer types. Our data suggested that the supervised elastic net
-penalized logistic regression model could be applied not only to the RAS
-activation pathway, but also to other genomic alterations. Both the RAS
-activation pathway and PIK3CA mutations are key, common cancer genomic
-alterations. When mutated, they exert significant effect on gene
-expression in cells. However, the supervised elastic net penalized
-logistic regression model cannot be generalized or applied to other
-genomic alterations. It might be a challenging prediction of genomic
-alterations that are infrequent or have a weak effect on gene
-expression.
+Our model showed good performance in predicting \emph{PIK3CA} mutations
+in various cancer types. Our data suggested that the supervised elastic
+net penalized logistic regression model could be applied not only to the
+RAS activation pathway, but also to other genomic alterations. Both the
+RAS activation pathway and \emph{PIK3CA} mutations are key, common
+cancer genomic alterations. Because they exert significant effect on
+gene expression in cancer cells, prediction from gene expression data
+can be good. However, the supervised elastic net penalized logistic
+regression model cannot be generalized or applied to other genomic
+alterations which have have a weak effect on gene expression.
 
@@
 been trained from between-sample normalization, a new sample cannot be
-exactly preprocessed with normalization which was done on trainset. A
-model based on gene expression from the TCGA pan-cancer dataset is
-limited in terms of data preprocessing. It is necessary to develop a
-preprocessing method that is independent of a dataset, to apply gene
-expression data to the prediction model.
+exactly processed with normalization which was done on trainset. A model
+based on gene expression from the TCGA pan-cancer dataset has limitation
+in terms of data preprocessing. It is necessary to develop a processing
+method that is independent of a dataset, to apply gene expression data
+to the prediction model.
 
-Our PIK3CA prediction model was similar to the RAS activation prediction
-model in terms of AUROC (0.84). However the AUPR of our model was lower
-than the RAS activation model (0.39 versus 0.63). The reason for our
-lower AUPR may be explained by the low prevalence rate of PIK3CA
-mutations, and an imbalanced dataset. The model for RAS activation
-trained with cancer types with more than 0.05 prevalence of RAS
-activation to avoid imbalance classification problem. We included all
-cancer types in our modeling process. The lower prevalence rate of
-target variables meant our dataset had a lower AUPR baseline. In the
-sub-analysis performance of each cancer type, the cancer types with
-higher PIK3CA mutation rates showed better AUPRs.
+Our \emph{PIK3CA} prediction model was similar to the RAS activation
+prediction model in terms of AUROC (0.84). However the AUPR of our model
+was lower than the RAS activation model (0.39 versus 0.63). The reason
+for our lower AUPR may be explained by an imbalanced dataset that has
+the low prevalence rate of \emph{PIK3CA} mutations {[}5{]}. The model
+for RAS activation trained with cancer types with more than 0.05
+prevalence of RAS activation to avoid imbalance classification problem.
+We included all cancer types in our modeling process. The lower
+prevalence rate of target variables meant our dataset had a lower AUPR
+baseline. In the sub-analysis performance of each cancer type, the
+cancer types with higher \emph{PIK3CA} mutation rates showed better
+AUPRs.
 
@@
 mutations across cancer types may be a reason for the strong predictive
-power of cancer types. If the cancer type was wrong or could not be
-determined, our model performance was poor.
+power of cancer types.
 
-Some significant gene expression predictors were closely related to PTEN
-and the PI3K pathway. PTEN and IGFR1R were the strongest gene expression
-predictors, which has negative and positive predictive powers. IGF1R is
-a tyrosine kinase receptor that activates PI3K {[}16{]}. {]}, and PTEN
-is an important regulator of PIP\textsubscript{3} by dephosphorylating
-PIP\textsubscript{3}, in contrast to PI3K {[}10{]}.
+Some significant gene expression predictors were closely related to the
+PTEN-PI3K pathway. \emph{PTEN} and \emph{IGFR1R} were the strongest gene
+expression predictors, which has negative and positive predictive
+powers. \emph{IGF1R} is a tyrosine kinase receptor that activates PI3K
+{[}15{]}, and \emph{PTEN} is an important regulator of
+PIP\textsubscript{3} by dephosphorylating PIP\textsubscript{3} {[}9{]}.
 
 Several studies have attempted to predict genomic alterations from gene
-expression data {[}6,17{]}. A study investigated PIK3CA mutation
+expression data {[}6,16{]}. A study investigated \emph{PIK3CA} mutation
 predictions using gene-expression signatures which is a sum of the
-average of the logarithmic gene expression . The model showed good
-performance AUROC 0.71 in an independent test set {[}6,17{]}.Another
+average of the logarithmic gene expression. The model showed good
+performance AUROC 0.71 in an independent test set {[}6,16{]}. Another
 study predicted copy number alterations with gene expression, using a
 multinomial logistic regression model with least absolute shrinkage and
-selection operator (LASSO) parameters {[}18{]}. The prediction of the
-1p/19q codel was very good, with an AUROC of 0.997, and gene-level
-predictions were good, with an AUROC of 0.75 {[}18{]}. A Hidden Markov
-Model Approach for Prediction of Genomic Alterations from Gene
-Expression Profiling {[}19{]}. A logistic regression model was used for
-MYCN gene amplification in neuroblastoma {[}20{]}.
+selection operator (LASSO) parameters {[}17{]}. The prediction of the
+1p/19q codeletion was very good, with an AUROC of 0.997, and gene-level
+predictions were good, with an AUROC of 0.75 {[}17{]}. A logistic
+regression model was used for \emph{MYCN Proto-Oncogene, BHLH
+Transcription Factor} (\emph{MYCN}) gene amplification in neuroblastoma
+{[}18{]}.
 
@@
 standardization of generation processing of gene expression data is also
-needed .
+needed.
 
@@
 \item
-  Figure 1. Prevalence rate of PIK3CA mutations across cancer types.
-  Cancer type abbreviations are explained in the S1 Appendix.
+  Figure 1. Prevalence rate of \emph{PIK3CA} mutations across cancer
+  types. Cancer type abbreviations are explained in the S1 Appendix.
 \item
@@
   characteristic (ROC) curve. Right: precision-recall (PR) curve of
-  training set and test set. The horizontal green line is the PIK3CA
-  mutation rate (0.11) (B) Correlation between training set and test set
-  of the area under the receiver operating characteristic curve (AUROC),
-  and the area under the precision-recall curve (AUPR) among cancer
-  types. The gray band is the 95\% confidence interval. Abbreviations
-  are explained in the S1 Appendix. (C) Correlations between the PIK3CA
-  mutation rate of the AUROC, and the AUPR.
+  training set and test set. The horizontal green line is the
+  \emph{PIK3CA} mutation rate (0.11) (B) Correlation between training
+  set and test set of the area under the receiver operating
+  characteristic curve (AUROC), and the area under the precision-recall
+  curve (AUPR) among cancer types. The gray band is the 95\% confidence
+  interval. Abbreviations are explained in the S1 Appendix. (C)
+  Correlations between the \emph{PIK3CA} mutation rate of the AUROC, and
+  the AUPR.
 \item
@@
 
-\leavevmode\hypertarget{ref-way_2018_machine_cellreports}{}%
-7. Way GP, Sanchez-Vega F, La K, Armenia J, Chatila WK, Luna A, et al.
-Machine Learning Detects Pan-cancer Ras Pathway Activation in The Cancer
-Genome Atlas. Cell Reports. Elsevier; 2018;23: 172--180.e3.
-doi:\href{https://doi.org/10.1016/j.celrep.2018.03.046}{10.1016/j.celrep.2018.03.046}
-
 \leavevmode\hypertarget{ref-andre_2019_alpelisib_nengljmeda}{}%
-8. André F, Ciruelos E, Rubovszky G, Campone M, Loibl S, Rugo HS, et al.
+7. André F, Ciruelos E, Rubovszky G, Campone M, Loibl S, Rugo HS, et al.
 Alpelisib for PIK3CA-Mutated, Hormone ReceptorPositive Advanced Breast
@@
 \leavevmode\hypertarget{ref-baileyComprehensiveCharacterizationCancer2018}{}%
-9. Bailey MH, Tokheim C, Porta-Pardo E, Sengupta S, Bertrand.
+8. Bailey MH, Tokheim C, Porta-Pardo E, Sengupta S, Bertrand.
 Comprehensive Characterization of Cancer Driver Genes and Mutations.
@@
 \leavevmode\hypertarget{ref-cantleyPhosphoinositide3KinasePathway2002}{}%
-10. Cantley LC. The Phosphoinositide 3-Kinase Pathway. Science. American
+9. Cantley LC. The Phosphoinositide 3-Kinase Pathway. Science. American
 Association for the Advancement of Science; 2002;296: 1655--1657.
@@
 \leavevmode\hypertarget{ref-jacobsenCgdsrRbasedAPI2019}{}%
-11. Jacobsen A, Luna A. Cgdsr: R-based API for accessing the MSKCC
+10. Jacobsen A, Luna A. Cgdsr: R-based API for accessing the MSKCC
 cancer genomics data server (CGDS). 2019.
@@
 \leavevmode\hypertarget{ref-kuhnRecipesPreprocessingTools2020}{}%
-12. Kuhn M, Wickham H. Recipes: Preprocessing tools to create design
+11. Kuhn M, Wickham H. Recipes: Preprocessing tools to create design
 matrices. 2020.
@@
 \leavevmode\hypertarget{ref-friedmanRegularizationPathsGeneralized2010}{}%
-13. Friedman J, Hastie T, Tibshirani R. Regularization Paths for
+12. Friedman J, Hastie T, Tibshirani R. Regularization Paths for
 Generalized Linear Models via Coordinate Descent. J Stat Softw. 2010;33:
@@
 \leavevmode\hypertarget{ref-saitoPrecisionRecallPlotMore2015}{}%
-14. Saito T, Rehmsmeier M. The Precision-Recall Plot Is More Informative
+13. Saito T, Rehmsmeier M. The Precision-Recall Plot Is More Informative
 than the ROC Plot When Evaluating Binary Classifiers on Imbalanced
@@
 \leavevmode\hypertarget{ref-kuhnTidymodelsEasilyInstall2020}{}%
-15. Kuhn M, Wickham H. Tidymodels: Easily install and load the
+14. Kuhn M, Wickham H. Tidymodels: Easily install and load the
 'tidymodels' packages. 2020.
@@
 \leavevmode\hypertarget{ref-leroithInsulinlikeGrowthFactor2003}{}%
-16. LeRoith D, Roberts CT. The insulin-like growth factor system and
+15. LeRoith D, Roberts CT. The insulin-like growth factor system and
 cancer. Cancer Letters. 2003;195: 127--137.
@@
 \leavevmode\hypertarget{ref-cizkovaGeneExpressionProfiling2010}{}%
-17. Cizkova M, Cizeron-Clairac G, Vacher S, Susini A, Andrieu C,
+16. Cizkova M, Cizeron-Clairac G, Vacher S, Susini A, Andrieu C,
 Lidereau R, et al. Gene Expression Profiling Reveals New Aspects of
@@
 \leavevmode\hypertarget{ref-muCNAPEMachineLearning2019}{}%
-18. Mu Q, Wang J. CNAPE: A Machine Learning Method for Copy Number
+17. Mu Q, Wang J. CNAPE: A Machine Learning Method for Copy Number
 Alteration Prediction from Gene Expression. IEEE/ACM Transactions on
@@
 
-\leavevmode\hypertarget{ref-gengHiddenMarkovModel2008}{}%
-19. Geng H, Ali HH, Chan WC. A Hidden Markov Model Approach for
-Prediction of Genomic Alterations from Gene Expression Profiling. In:
-Măndoiu I, Sunderraman R, Zelikovsky A, editors. Bioinformatics Research
-and Applications. Berlin, Heidelberg: Springer; 2008. pp. 414--425.
-doi:\href{https://doi.org/10.1007/978-3-540-79450-9_38}{10.1007/978-3-540-79450-9\_38}
-
 \leavevmode\hypertarget{ref-heGeneSignaturesAssociated2020}{}%
-20. He X, Qin C, Zhao Y, Zou L, Zhao H, Cheng C. Gene signatures
+18. He X, Qin C, Zhao Y, Zou L, Zhao H, Cheng C. Gene signatures
 associated with genomic aberrations predict prognosis in neuroblastoma.

### 2020-07-22 plosOne/plosOne.Rmd
submition

@@
 Approximately 10,000 cases were available for *PIK3CA* mutation and mRNA expression data.
-In 10-fold cross-validation, the model with $\lambda$ = 0.01 and $\alpha$ = 1.0 (Ridge regression) showed the best performance, in terms of area under the receiver operating characteristic (AUROC).
+In 10-fold cross-validation, the model with $\lambda$ = 0.01 and $\alpha$ = 1.0 (ridge regression) showed the best performance, in terms of area under the receiver operating characteristic (AUROC).
 The final model was developed with selected hyper-parameters using the entire training set.
@@
 Cancer types were the most important predictors.
-Both *insulin like growth factor 1 Receptor* (*IGF1R*) and the *phosphatase and tensin homolog* (*PTEN*) were the most significant genes in gene expression predictors.
+Both *insulin like growth factor 1 receptor* (*IGF1R*) and the *phosphatase and tensin homolog* (*PTEN*) were the most significant genes in gene expression predictors.
 Our study suggests that predicting genomic alterations using gene expression data is possible, with good outcomes.
@@
 
-For 10-fold cross-validation, the model with $\lambda$ = 0.01 and $\alpha$ = 1.0 (Ridge regression) showed the best performance in terms of AUROC.
+For 10-fold cross-validation, the model with $\lambda$ = 0.01 and $\alpha$ = 1.0 (ridge regression) showed the best performance in terms of AUROC.
 The final model was trained with the selected hyper-parameters with the entire training set.

### 2020-07-22 plosOne/plosOne.tex
submition

@@
 cross-validation, the model with \(\lambda\) = 0.01 and \(\alpha\) = 1.0
-(Ridge regression) showed the best performance, in terms of area under
+(ridge regression) showed the best performance, in terms of area under
 the receiver operating characteristic (AUROC). The final model was
@@
 the test set AUPR was 0.39. Cancer types were the most important
-predictors. Both \emph{insulin like growth factor 1 Receptor}
+predictors. Both \emph{insulin like growth factor 1 receptor}
 (\emph{IGF1R}) and the \emph{phosphatase and tensin homolog}
@@
 For 10-fold cross-validation, the model with \(\lambda\) = 0.01 and
-\(\alpha\) = 1.0 (Ridge regression) showed the best performance in terms
+\(\alpha\) = 1.0 (ridge regression) showed the best performance in terms
 of AUROC. The final model was trained with the selected hyper-parameters

### 2020-09-09 Response to Reviewers.Rmd
revision

---
title: "Response to Reviewers"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Dear, Editor-in-Chief, PLOS ONE

Thank you for considering our manuscript entitled "Prediction of PIK3CA mutations from cancer gene expression data (PONE-D-20-22669)." We are really happy to hear that the manuscript is considering for publication with a revision. We considered all suggestions that were raised by the reviewers and tried to answer to all queries to the best of our capability. In our annotated version, we indicated what we responded to reviewers’ suggestions by using Track Change function with marginal notes and symbols (e.g., R1-1 indicates response to comment #1 of Reviewer #1). Our specific responses are as follows:

Comments to the Author

1. Is the manuscript technically sound, and do the data support the conclusions?

The manuscript must describe a technically sound piece of scientific research with data that supports the conclusions. Experiments must have been conducted rigorously, with appropriate controls, replication, and sample sizes. The conclusions must be drawn appropriately based on the data presented.

Reviewer #1: Partly

Reviewer #2: Yes

2. Has the statistical analysis been performed appropriately and rigorously?

Reviewer #1: I Don't Know

Reviewer #2: Yes

3. Have the authors made all data underlying the findings in their manuscript fully available?

The PLOS Data policy requires authors to make all data underlying the findings described in their manuscript fully available without restriction, with rare exception (please refer to the Data Availability Statement in the manuscript PDF file). The data should be provided as part of the manuscript or its supporting information, or deposited to a public repository. For example, in addition to summary statistics, the data points behind means, medians and variance measures should be available. If there are restrictions on publicly sharing data—e.g. participant privacy or use of data from a third party—those must be specified.

Reviewer #1: Yes

Reviewer #2: Yes

4. Is the manuscript presented in an intelligible fashion and written in standard English?

PLOS ONE does not copyedit accepted manuscripts, so the language in submitted articles must be clear, correct, and unambiguous. Any typographical or grammatical errors should be corrected at revision, so please note any specific errors here.

Reviewer #1: Yes

Reviewer #2: Yes

5. Review Comments to the Author

Please use the space provided to explain your answers to the questions above. You may also include additional comments for the author, including concerns about dual publication, research ethics, or publication ethics. (Please upload your review as an attachment if it exceeds 20,000 characters)

Reviewer #1:

Comment 1

This is a regulated regression analysis to find out the most significant variable(s) and that is kept in the final model. This is a very concise article and lack of detail methodology.

Response

Thank you for your constructive feedback considering the lack of detailed methodology. TCGA is a widely used public data of cancer genomics. The detail of the TCGA pan-cancer data is described in the reference. For the method of prediction modeling, we tried to follow guidelines for developing and reporting machine learning predictive models in biomedical research. We add a supplementary figure to help understand hyperparameter tuning.

Luo, Wei, Dinh Phung, Truyen Tran, Sunil Gupta, Santu Rana, Chandan Karmakar, Alistair Shilton, et al. “Guidelines for Developing and Reporting Machine Learning Predictive Models in Biomedical Research: A Multidisciplinary View.” Journal of Medical Internet Research 18, no. 12 (2016): e323. https://doi.org/10.2196/jmir.5870.

Comment 2

It has been known for quite some times that PIK3CA GOF mutation is very common (in fact next to TP53) in solid tumors. It is also known that PIK3CA is very much related to PTEN and IGF1R signaling.
The finding is not new and the rationale for this article is not very clear.

Response

Thank you for your opinion. Our study aims to build the PIK3CA mutation prediction model not to search for important variables. Because the penalized logistic regression model is highly interpretable, we were able to find significant variables like IGF1R and PTEN. But this findings of significant variables is not the primary purpose of this study.  The purpose of this study is to investigate the *PIK3CA* mutation prediction performance of machine learning models. The purpose of the study is further described in the manuscript.

Comment 3

More importantly, this type of article is not suitable for PLOS ONE audience. Authors may consider to some bio-informatics or bio-statistics journal.

Response

Thank you for your suggestion. Since machine learning modeling is complex and has begun to be widely used relatively recently, the audience may lack an understanding of detailed methods. However, our study used a widely used data set (TCGA) and modeling framework (R tidymodels package). We believe our research will benefit audiences interested in applying machine learning to patient care. We also believe that publishers targeting a broad audience are publishing predictive model studies using machine learning.

Reviewer #2:

Comment 1

The authors present a succinct study on the prediction of PIK3CA mutations from gene expression data. This study applies an elastic net penalized logistic regression classifier to the cancer genome atlas (TCGA) pan-cancer gene expression dataset, a method that was previously established for detecting RAS pathway activation. The methods used and the results presented in the figures appear to be appropriate for the work performed. Both the AUROC and AUPRC demonstrate predictive performance well above baseline. Limitations of the approach used were also appropriately discussed.

Response

Thank you for your opinion.

Comment 2

It may be questionable why PIK3CA mutation prediction from mRNA expression is useful when targeted sequencing panels can assay these mutations directly, but it has been proposed elsewhere that clinical transcriptomics may add important functional or phenotypic information.

Response

As you pointed out, the clinical utility of PIK3CA mutation prediction from mRNA expression is unclear because most direct genomic tests are more specific and sensitive than predictive models. Our prediction model is not an application that is immediately applicable to a cancer patient for the detection of PIK3CA mutation. It is not known how it will be used, but finding out the mutation prediction performance using gene expression data could play a role in advancing machine learning to be helpful in patient treatment. We discussed further the limitations of this study in the manuscript.

Comment 3

Overall this work demonstrates that machine learning approaches can predict PIK3CA mutation status from gene expression data with a reasonably good level of performance.

Response

Thank you for your opinion.

Thank you again for reconsidering our manuscript. We think we did our best to revise faithfully our previous manuscript in line with the indications that were raised by the reviewer. We hope the revised version would meet with your, and the reviewer’s approval and be finally accepted by the PLOS ONE.

With best wishes and respectfulness,
Youn Soo Lee, MD

### 2020-09-09 Decision_letter.Rmd
revision

---
title: "decision_letter"
author: "Jun Kang"
date: '2020 8 31 '
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

PONE-D-20-22669
Prediction of PIK3CA mutations from cancer gene expression data
PLOS ONE

Dear Dr. Lee,

Thank you for submitting your manuscript to PLOS ONE. After careful consideration, we feel that it has merit but does not fully meet PLOS ONE’s publication criteria as it currently stands. Therefore, we invite you to submit a revised version of the manuscript that addresses the points raised during the review process.

Please respond to the comments item-by-item to satisfy the reviewers' concerns. Please carefully edit the MS for English corrections and typos.

Please submit your revised manuscript by Oct 10 2020 11:59PM. If you will need more time than this to complete your revisions, please reply to this message or contact the journal office at plosone@plos.org. When you're ready to submit your revision, log on to https://www.editorialmanager.com/pone/ and select the 'Submissions Needing Revision' folder to locate your manuscript file.

Please include the following items when submitting your revised manuscript:

A rebuttal letter that responds to each point raised by the academic editor and reviewer(s). You should upload this letter as a separate file labeled 'Response to Reviewers'.
A marked-up copy of your manuscript that highlights changes made to the original version. You should upload this as a separate file labeled 'Revised Manuscript with Track Changes'.
An unmarked version of your revised paper without tracked changes. You should upload this as a separate file labeled 'Manuscript'.

If you would like to make changes to your financial disclosure, please include your updated statement in your cover letter. Guidelines for resubmitting your figure files are available below the reviewer comments at the end of this letter.

If applicable, we recommend that you deposit your laboratory protocols in protocols.io to enhance the reproducibility of your results. Protocols.io assigns your protocol its own identifier (DOI) so that it can be cited independently in the future. For instructions see: http://journals.plos.org/plosone/s/submission-guidelines#loc-laboratory-protocols

We look forward to receiving your revised manuscript.

Kind regards,

Nandini Dey, MS., Ph.D
Academic Editor
PLOS ONE

Journal Requirements:

When submitting your revision, we need you to address these additional requirements.

1. Please ensure that your manuscript meets PLOS ONE's style requirements, including those for file naming. The PLOS ONE style templates can be found at

https://journals.plos.org/plosone/s/file?id=wjVg/PLOSOne_formatting_sample_main_body.pdf and

https://journals.plos.org/plosone/s/file?id=ba62/PLOSOne_formatting_sample_title_authors_affiliations.pdf

2. PLOS requires an ORCID iD for the corresponding author in Editorial Manager on papers submitted after December 6th, 2016. Please ensure that you have an ORCID iD and that it is validated in Editorial Manager. To do this, go to ‘Update my Information’ (in the upper left-hand corner of the main menu), and click on the Fetch/Validate link next to the ORCID field. This will take you to the ORCID site and allow you to create a new iD or authenticate a pre-existing iD in Editorial Manager. Please see the following video for instructions on linking an ORCID iD to your Editorial Manager account: https://www.youtube.com/watch?v=_xcclfuvtxQ

[Note: HTML markup is below. Please do not edit.]

Reviewers' comments:

Reviewer's Responses to Questions

Comments to the Author

1. Is the manuscript technically sound, and do the data support the conclusions?

The manuscript must describe a technically sound piece of scientific research with data that supports the conclusions. Experiments must have been conducted rigorously, with appropriate controls, replication, and sample sizes. The conclusions must be drawn appropriately based on the data presented.

Reviewer #1: Partly

Reviewer #2: Yes

2. Has the statistical analysis been performed appropriately and rigorously?

Reviewer #1: I Don't Know

Reviewer #2: Yes

3. Have the authors made all data underlying the findings in their manuscript fully available?

The PLOS Data policy requires authors to make all data underlying the findings described in their manuscript fully available without restriction, with rare exception (please refer to the Data Availability Statement in the manuscript PDF file). The data should be provided as part of the manuscript or its supporting information, or deposited to a public repository. For example, in addition to summary statistics, the data points behind means, medians and variance measures should be available. If there are restrictions on publicly sharing data—e.g. participant privacy or use of data from a third party—those must be specified.

Reviewer #1: Yes

Reviewer #2: Yes

4. Is the manuscript presented in an intelligible fashion and written in standard English?

PLOS ONE does not copyedit accepted manuscripts, so the language in submitted articles must be clear, correct, and unambiguous. Any typographical or grammatical errors should be corrected at revision, so please note any specific errors here.

Reviewer #1: Yes

Reviewer #2: Yes

5. Review Comments to the Author

Please use the space provided to explain your answers to the questions above. You may also include additional comments for the author, including concerns about dual publication, research ethics, or publication ethics. (Please upload your review as an attachment if it exceeds 20,000 characters)

Reviewer #1: This is a regulated regression analysis to find out the most significant variable(s) and that is kept in the final model. This is a very concise article and lack of detail methodology. It has been known for quite some times that PIK3CA GOF mutation is very common (in fact next to TP53) in solid tumors. It is also known that PIK3CA is very much related to PTEN and IGF1R signaling.
The finding is not new and the rationale for this article is not very clear.

More importantly, this type of article is not suitable for PLOS ONE audience.
Authors may consider to some bio-informatics or bio-statistics journal.

Reviewer #2: The authors present a succinct study on the prediction of PIK3CA mutations from gene expression data. This study applies an elastic net penalized logistic regression classifier to the cancer genome atlas (TCGA) pan-cancer gene expression dataset, a method that was previously established for detecting RAS pathway activation. The methods used and the results presented in the figures appear to be appropriate for the work performed. Both the AUROC and AUPRC demonstrate predictive performance well above baseline. Limitations of the approach used were also appropriately discussed. It may be questionable why PIK3CA mutation prediction from mRNA expression is useful when targeted sequencing panels can assay these mutations directly, but it has been proposed elsewhere that clinical transcriptomics may add important functional or phenotypic information. Overall this work demonstrates that machine learning approaches can predict PIK3CA mutation status from gene expression data with a reasonably good level of performance.

6. PLOS authors have the option to publish the peer review history of their article (what does this mean?). If published, this will include your full peer review and any attached files.

If you choose “no”, your identity will remain anonymous but your review may still be made public.

Do you want your identity to be public for this peer review? For information about this choice, including consent withdrawal, please see our Privacy Policy.

Reviewer #1: No

Reviewer #2: No

[NOTE: If reviewer comments were submitted as an attachment file, they will be attached to this email and accessible via the submission site. Please log into your account, locate the manuscript record, and check for the action link "View Attachments". If this link does not appear, there are no attachment files.]

While revising your submission, please upload your figure files to the Preflight Analysis and Conversion Engine (PACE) digital diagnostic tool, https://pacev2.apexcovantage.com/. PACE helps ensure that figures meet PLOS requirements. To use PACE, you must first register as a user. Registration is free. Then, login and navigate to the UPLOAD tab, where you will find detailed instructions on how to use the tool. If you encounter any issues or have any questions when using PACE, please email PLOS at figures@plos.org. Please note that Supporting Information files do not need this step.

In compliance with data protection regulations, you may request that we remove your personal registration details at any time. (Remove my information/details). Please contact the publication office if you have any questions.

### 2020-09-09 S1.Table.Rmd
revision

---
title: "S1.Table"
author: "Jun Kang"
date: '2020 9 3 '
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r hyperparameter}
library(tidymodels)

load(here::here("glmn_tune_pancancer.RData"))

tune::show_best(glmn_tune, metric = "roc_auc")
```

### 2020-09-09 plosOne/S1Appendix.Rmd
revision

@@
 
-read_delim("../TCGAAbbreviations.txt", delim = "\t") %>%
+read_delim(here::here("TCGAAbbreviations.txt"), delim = "\t") %>%
   knitr::kable()
@@
 ```
+
+```{r select model}
+
+library(tidymodels)
+library(latex2exp)
+
+load(here::here("glmn_tune_pancancer.RData"))
+
+roc_vals <- 
+  collect_metrics(glmn_tune) %>% 
+  filter(.metric == "roc_auc")
+roc_vals %>% 
+  mutate(mixture = format(mixture)) %>% 
+  ggplot(aes(x = penalty, y = mean, col = mixture)) + 
+  geom_line() + 
+  geom_point() + 
+  geom_errorbar(aes(ymin = mean - std_err, ymax = mean + std_err, width = 0.05)) +
+  scale_x_log10() +
+  xlab(TeX("Penalty ($\\lambda$)")) +
+  ylab("Area under the receiver operating characteristic (AUROC)") +
+  labs(color = TeX("Mixture ($\\alpha$)"))
+
+```

### 2020-09-09 plosOne/plosOne_revision.Rmd
revision

--- 2020-07-22 plosOne/plosOne.Rmd
@@
 We wanted to ascertain whether this prediction model approach could be applied not only to RAS pathway activation, but also to *PIK3CA* mutation predictions.
+The purpose of this study is to investigate the *PIK3CA* mutation prediction performance of a supervised elastic net penalized logistic regression model.
 
@@
 
-For 10-fold cross-validation, the model with $\lambda$ = 0.01 and $\alpha$ = 1.0 (ridge regression) showed the best performance in terms of AUROC.
+For 10-fold cross-validation, the model with $\lambda$ = 0.01 and $\alpha$ = 1.0 (ridge regression) showed the best performance in terms of AUROC (S1 Figure).
 The final model was trained with the selected hyper-parameters with the entire training set.
@@
 
+The clinical utility of PIK3CA mutation prediction from mRNA expression is unclear because most direct genomic tests are more specific and sensitive than predictive models.
+Our prediction model is not an application that is immediately applicable to a cancer patient for the detection of PIK3CA mutation.
+It is not known how it will be used, but finding out the mutation prediction performance using gene expression data could play a role in advancing machine learning to be helpful in patient treatment.
+
 Our study suggested that the prediction of genomic alterations using gene expression data was possible, with good performance.
@@
 
-- S1 Appendix.
+- S1 Appendix. Abbreviations of cancer type.  
+- S1 Figure.
+Hyperparameter tuning and performance assessment in 10-fold cross-validation resampling.
+The x-axis is a penalty scaling parameter: $\lambda$ {$10^{-5}$, $10^{-4}$,$10^{-3}$,$10^{-2}$,$10^{-1}$, $10^{0}$}, color is mixture hyperparameter of penalty function: $\alpha$ {0.0, 0.25, 0.5, 0.75, 1.0}.
+y-axis is estimates of area under the receiver operating characteristic (AUROC) using 10-fold cross-validation resampling.
 
 # References {#references}
+
+<script src="https://hypothes.is/embed.js" async></script>

### 2020-09-09 plosOne/plosOne_revision.tex
revision

--- 2020-07-22 plosOne/plosOne.tex
@@
 prediction model approach could be applied not only to RAS pathway
-activation, but also to \emph{PIK3CA} mutation predictions.
+activation, but also to \emph{PIK3CA} mutation predictions. The purpose
+of this study is to investigate the \emph{PIK3CA} mutation prediction
+performance of machine learning models.
 
@@
 \(\alpha\) = 1.0 (ridge regression) showed the best performance in terms
-of AUROC. The final model was trained with the selected hyper-parameters
-with the entire training set. The training set AUROC was 0.93 and the
-test set AUROC was 0.84. The AUPR of the training set was 0.66 and the
-test set AUPR was 0.39 (Figure 2A).
+of AUROC (S1 Figure). The final model was trained with the selected
+hyper-parameters with the entire training set. The training set AUROC
+was 0.93 and the test set AUROC was 0.84. The AUPR of the training set
+was 0.66 and the test set AUPR was 0.39 (Figure 2A).
 
@@
 
+The clinical utility of PIK3CA mutation prediction from mRNA expression
+is unclear because most direct genomic tests are more specific and
+sensitive than predictive models. Our prediction model is not an
+application that is immediately applicable to a cancer patient for
+detection of PIK3CA mutation. It is not known how it will be used, but
+finding out the mutation prediction performance using gene expression
+data could play a role in advancing machine learning to be helpful in
+patient treatment.
+
 Our study suggested that the prediction of genomic alterations using
@@
 \item
-  S1 Appendix.
+  S1 Appendix. Abbreviations of cancer type.\\
+\item
+  S1 Figure. Hyperparameter tuning and performance assessment in 10-fold
+  cross-validation resampling. The x-axis is a penalty scaling
+  parameter: \(\lambda\) \{\(10^{-5}\),
+  \(10^{-4}\),\(10^{-3}\),\(10^{-2}\),\(10^{-1}\), \(10^{0}\)\}, color
+  is mixture hyperparameter of penalty function: \(\alpha\) \{0.0, 0.25,
+  0.5, 0.75, 1.0\}. y-axis is estimates of area under the receiver
+  operating characteristic (AUROC) using 10-fold cross-validation
+  resampling.
 \end{itemize}
@@
 \hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
+\section{References}\label{references}}
 

### 2020-09-09 figure.R

@@ -135,3 +135,27 @@ typePlot +
        y = "Cancer type") +
   scale_colour_Publication() +
   theme_Publication()
+library(tidyverse)
+library(tidymodels)
+library(latex2exp)
+library(Cairo)
+
+load(here::here("glmn_tune_pancancer.RData"))
+
+roc_vals <- 
+  collect_metrics(glmn_tune) %>% 
+  filter(.metric == "roc_auc")
+
+CairoPDF(file = 'S1Figure.pdf',
+         width = 5, height = 5, pointsize=10)
+roc_vals %>% 
+  mutate(mixture = format(mixture)) %>% 
+  ggplot(aes(x = penalty, y = mean, col = mixture)) + 
+  geom_line() + 
+  geom_point() + 
+  geom_errorbar(aes(ymin = mean - std_err, ymax = mean + std_err, width = 0.05)) +
+  scale_x_log10() +
+  xlab(TeX("Penalty ($\\lambda$)")) +
+  ylab("Area under the receiver operating characteristic (AUROC)") +
+  labs(color = TeX("Mixture ($\\alpha$)"))
+dev.off()
